---
title: "Plaque Assay + Violin Plots — Viral Titer Analysis"
author: "Alejandro Ponce-Flores"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: cerulean
    toc: true
    number_sections: true
    code_folding: hide
  pdf_document:
    toc: true
    number_sections: true
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(
  fig.width  = 6,
  fig.height = 4,
  dpi        = 600,
  fig.align  = "center"
)
```

```{r user-config, message=FALSE, warning=FALSE}
## ── USER CONFIGURATION ──────────────────────────────────────────────────────
#
# DATA_FILE: Path to your CSV file with plaque assay data.
#   Expected columns (rename yours to match, or update the rename() call below):
#   - "Mouse ID"              : animal/sample identifier
#   - "Group"                 : group label (e.g., "Group 1", "Group 2")
#   - "Virus"                 : virus or mock label used in sample
#   - "# plaques counted R1"  : plaque count replicate 1
#   - "# plaques counted R2"  : plaque count replicate 2
#   - "Average plaques from duplicate wells in a 12-well plate"
#   - "Plated 10e-1 to 10e-6 dilution"
#   - "Total Volume (ml)"
#   - "pfu/mL right-side lung homogenate"  : calculated viral titer
#
DATA_FILE <- "data/titer_data.csv"

# EXCLUDE_PATTERN: Regex to drop rows you do not want (e.g., backtitrations).
#   Set to NULL to keep all rows.
EXCLUDE_PATTERN <- NULL   # e.g., "backtitration"

# GROUP_MAP: Named character vector mapping your "Group N" labels to
#   human-readable short names used in all plots and tables.
#   Keys  = values in the "Group" column of your CSV.
#   Values = display names (shown on plot axes).
#
#   NOTE: for the shipped demo data this mapping is asserted against the
#   simulator (data/simulate_data.R defines the same Group -> label pairs in
#   GROUP_DEFS). Every group present in the CSV must appear as a key here —
#   the template stops with an error otherwise, rather than silently dropping
#   unmapped animals from the plots and the ANOVA.
GROUP_MAP <- c(
  "Group 1" = "Mock + Compound A",
  "Group 2" = "Mock + Compound B",
  "Group 3" = "Compound A",
  "Group 4" = "Compound B",
  "Group 5" = "Drug X",
  "Group 6" = "Compound A + Drug X",
  "Group 7" = "Compound B + Drug X",
  "Group 8" = "Vehicle"
)

# GROUP_ORDER: Display order on the x-axis. Must contain the same strings
#   as the values of GROUP_MAP.
GROUP_ORDER <- c(
  "Mock + Compound A",
  "Mock + Compound B",
  "Vehicle",
  "Drug X",
  "Compound A",
  "Compound A + Drug X",
  "Compound B",
  "Compound B + Drug X"
)

# PALETTE_GROUPS: Fill color for each group in the full violin plot.
#   Names must match the values of GROUP_MAP / GROUP_ORDER.
PALETTE_GROUPS <- c(
  "Mock + Compound A"    = "grey85",
  "Mock + Compound B"    = "grey70",
  "Vehicle"              = "grey40",
  "Drug X"               = "#984ea3",
  "Compound A"           = "#74add1",
  "Compound B"           = "#4575b4",
  "Compound A + Drug X"  = "#66c2a5",
  "Compound B + Drug X"  = "#238b45"
)

# MERGE_MAP: Optionally collapse related groups for the combined-group plot.
#   Keys = ShortGroup values; Values = merged display name.
#   Groups not listed here are kept as-is.
MERGE_MAP <- c(
  "Mock + Compound A"    = "Mock",
  "Mock + Compound B"    = "Mock",
  "Compound A"           = "Inhibitor",
  "Compound B"           = "Inhibitor",
  "Compound A + Drug X"  = "Drug + Inhibitor",
  "Compound B + Drug X"  = "Drug + Inhibitor",
  "Vehicle"              = "Vehicle",
  "Drug X"               = "Drug X"
)

# MERGE_ORDER: Display order for the combined-group violin plot.
MERGE_ORDER <- c("Mock", "Vehicle", "Drug X", "Inhibitor", "Drug + Inhibitor")

# PALETTE_MERGED: Fill colors for the combined-group violin plot.
PALETTE_MERGED <- c(
  "Mock"            = "grey80",
  "Vehicle"         = "grey40",
  "Drug X"          = "#984ea3",
  "Inhibitor"       = "#4575b4",
  "Drug + Inhibitor"= "#238b45"
)

# COMPARISONS: Pairs of merged groups to test with stat_compare_means().
#   Each element is a length-2 character vector of MERGE_ORDER values.
COMPARISONS <- list(
  c("Drug X", "Inhibitor"),
  c("Drug X", "Drug + Inhibitor"),
  c("Inhibitor", "Drug + Inhibitor")
)

# TITER_COL: Column name in your CSV that holds the computed titer value.
TITER_COL <- "pfu/mL right-side lung homogenate"

# PSEUDOCOUNT: Added to the titer before log10 so that true zeros (below the
#   limit of detection) survive both the statistics AND the log-scaled figures.
#   Use the same value everywhere — never log10() a raw titer.
PSEUDOCOUNT <- 1

# ALPHA: Significance threshold used for assumption checks and for deciding
#   which Tukey comparisons get a bracket on the final figure.
ALPHA <- 0.05

# LOA_MULTIPLIER: Width of the replicate limits of agreement, in SDs of the
#   replicate difference (1.96 = 95% LoA).
LOA_MULTIPLIER <- 1.96

# OUTPUT_DIR: Where figures are written. Created if it does not exist.
OUTPUT_DIR <- "Plots"

# Figure export settings (applied to every ggsave() call below).
FIG_WIDTH  <- 8      # inches
FIG_HEIGHT <- 5      # inches
FIG_DPI    <- 600

# VIOLIN_BW: Kernel bandwidth for geom_violin, on the log10 scale.
#   Increase for smoother violins, decrease for more detail.
VIOLIN_BW <- 0.15

## ────────────────────────────────────────────────────────────────────────────
```

# Setup

## Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(janitor)
library(stringr)
library(ggplot2)
library(scales)
library(rstatix)
library(car)
library(ggpubr)
library(gt)
library(broom)
```

## Load Data

```{r load-data, message=FALSE, warning=FALSE}
if (!file.exists(DATA_FILE)) {
  stop("DATA_FILE not found: '", DATA_FILE, "'. ",
       "Run `Rscript data/simulate_data.R` from the template directory to ",
       "regenerate the demo data, or point DATA_FILE at your own CSV.")
}

df <- read_csv(DATA_FILE)

# ── Input validation ─────────────────────────────────────────────────────────
# Fail loudly and early rather than producing plots from a partially parsed
# file. Every column renamed in the next chunk must exist.
REQUIRED_COLS <- c(
  "Mouse ID", "Group",
  "# plaques counted R1", "# plaques counted R2",
  "Average plaques from duplicate wells in a 12-well plate",
  "Plated 10e-1 to 10e-6 dilution", "Total Volume (ml)",
  TITER_COL
)
missing_cols <- setdiff(REQUIRED_COLS, names(df))
if (length(missing_cols) > 0) {
  stop("DATA_FILE is missing required column(s): ",
       paste(sQuote(missing_cols), collapse = ", "),
       ". Rename your columns to match, or edit REQUIRED_COLS and the ",
       "rename() call in the next chunk.")
}

# Every group in the data must be mapped, or it silently disappears from the
# figures and the ANOVA when factor(levels = GROUP_ORDER) turns it into NA.
unmapped_groups <- setdiff(unique(df$Group), names(GROUP_MAP))
if (length(unmapped_groups) > 0) {
  stop("Group value(s) present in the data but absent from GROUP_MAP: ",
       paste(sQuote(unmapped_groups), collapse = ", "),
       ". Add them to GROUP_MAP (and to GROUP_ORDER / PALETTE_GROUPS / MERGE_MAP).")
}
stopifnot(all(unique(df$Group) %in% names(GROUP_MAP)))

# The display vocabularies must agree with each other too.
stopifnot(
  all(GROUP_MAP   %in% GROUP_ORDER),
  all(GROUP_ORDER %in% GROUP_MAP),
  all(GROUP_ORDER %in% names(PALETTE_GROUPS)),
  all(GROUP_ORDER %in% names(MERGE_MAP)),
  all(MERGE_MAP   %in% MERGE_ORDER),
  all(MERGE_ORDER %in% names(PALETTE_MERGED)),
  all(unlist(COMPARISONS) %in% MERGE_ORDER)
)

dir.create(OUTPUT_DIR, showWarnings = FALSE, recursive = TRUE)
```

## Clean and Organize

```{r clean}
df_clean <- df %>%
  { if (!is.null(EXCLUDE_PATTERN)) filter(., !str_detect(Group, EXCLUDE_PATTERN)) else . } %>%
  rename(
    MouseID    = `Mouse ID`,
    Plaques_R1 = `# plaques counted R1`,
    Plaques_R2 = `# plaques counted R2`,
    AvgPlaques = `Average plaques from duplicate wells in a 12-well plate`,
    Dilution   = `Plated 10e-1 to 10e-6 dilution`,
    Volume_mL  = `Total Volume (ml)`,
    Titer      = all_of(TITER_COL)
  ) %>%
  mutate(
    ShortGroup  = coalesce(unname(GROUP_MAP[Group]), Group),
    MergedGroup = coalesce(unname(MERGE_MAP[ShortGroup]), ShortGroup),
    ShortGroup  = factor(ShortGroup, levels = GROUP_ORDER),
    MergedGroup = factor(MergedGroup, levels = MERGE_ORDER),
    mean_plaques  = (Plaques_R1 + Plaques_R2) / 2,
    diff_plaques  = Plaques_R1 - Plaques_R2,
    # Titer_pseudo is what every log-scaled FIGURE plots, and logTiter is what
    # every TEST uses. Both carry the same pseudocount, so figure n == test n
    # and animals at the limit of detection (Titer == 0) are never dropped.
    Titer_pseudo  = Titer + PSEUDOCOUNT,
    logTiter      = log10(Titer + PSEUDOCOUNT)
  )

# The GROUP_MAP / MERGE_MAP checks above guarantee every row is mapped; assert
# the factor conversion did not introduce NAs anyway (e.g. from a stray level).
stopifnot(
  !any(is.na(df_clean$ShortGroup)),
  !any(is.na(df_clean$MergedGroup))
)

# ── Limit of Agreement (LoA) ─────────────────────────────────────────────────
loa_stats <- df_clean %>%
  summarise(
    bias    = mean(diff_plaques, na.rm = TRUE),
    sd_diff = sd(diff_plaques, na.rm = TRUE)
  )
loa_upper <- loa_stats$bias + LOA_MULTIPLIER * loa_stats$sd_diff
loa_lower <- loa_stats$bias - LOA_MULTIPLIER * loa_stats$sd_diff

df_clean <- df_clean %>%
  mutate(
    flag_bad_replicate = diff_plaques < loa_lower | diff_plaques > loa_upper
  )
```

# Check Replicates

```{r replicate-check}
labeled_points <- filter(df_clean, flag_bad_replicate)

print(
  ggplot(df_clean, aes(x = Plaques_R1, y = Plaques_R2, color = flag_bad_replicate)) +
    geom_point(size = 2) +
    geom_abline(slope = 1, intercept = 0, color = "grey40") +
    geom_text(
      data    = labeled_points,
      mapping = aes(
        x = Plaques_R1,
        y = Plaques_R2,
        label = as.character(MouseID)
      ),
      nudge_y = 1, size = 3, color = "orange", inherit.aes = FALSE
    ) +
    scale_color_manual(values = c("black", "orange"), labels = c("OK", "Outside LoA")) +
    labs(
      x     = "Plaques counted (Replicate 1)",
      y     = "Plaques counted (Replicate 2)",
      color = "Flagged",
      title = paste0("Replicate Agreement (", LOA_MULTIPLIER,
                     " × SD Limits of Agreement)")
    ) +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
)
```

Data points outside the LoA are flagged in orange. Review these samples before proceeding.

# Plots

## Full Group Violin Plot

```{r violin-full, message=FALSE, warning=FALSE}
p_full <- ggplot(df_clean, aes(x = ShortGroup, y = Titer_pseudo, fill = ShortGroup)) +
  geom_violin(trim = FALSE, alpha = 0.7, color = "black", bw = VIOLIN_BW) +
  stat_summary(
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "pointrange", color = "black", size = 0.8, alpha = 0.9, shape = 22
  ) +
  geom_jitter(width = 0.18, size = 1, alpha = 0.60) +
  scale_fill_manual(values = PALETTE_GROUPS, guide = "none") +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(
    x       = "Group",
    y       = paste0("Viral Titer + ", PSEUDOCOUNT, " (pfu/mL, log scale)"),
    title   = "Viral Titers by Group",
    caption = paste0("Black bars indicate mean \u00b1 1 SD. Titers are plotted as ",
                     "Titer + ", PSEUDOCOUNT,
                     " so that animals at the limit of detection remain visible.")
  ) +
  theme_bw(base_size = 13) +
  theme(
    axis.text.x      = element_text(angle = 30, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(OUTPUT_DIR, "Viral_Titer_By_Group.tiff"), plot = p_full,
       width = FIG_WIDTH, height = FIG_HEIGHT, units = "in", dpi = FIG_DPI,
       compression = "lzw")
print(p_full)
```

## Combined Group Violin Plot

```{r violin-merged, message=FALSE, warning=FALSE}
p_merged <- ggplot(df_clean, aes(x = MergedGroup, y = Titer_pseudo, fill = MergedGroup)) +
  geom_violin(trim = FALSE, alpha = 0.45, color = "black", bw = VIOLIN_BW) +
  geom_jitter(width = 0.18, size = 1, alpha = 0.60) +
  scale_fill_manual(values = PALETTE_MERGED, guide = "none") +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(
    x     = "Group",
    y     = paste0("Viral Titer + ", PSEUDOCOUNT, " (pfu/mL, log scale)"),
    title = "Viral Titers by Combined Group"
  ) +
  theme_bw(base_size = 13) +
  theme(
    axis.text.x      = element_text(angle = 30, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(OUTPUT_DIR, "Viral_Titer_By_MergedGroup.tiff"), plot = p_merged,
       width = FIG_WIDTH, height = FIG_HEIGHT, units = "in", dpi = FIG_DPI,
       compression = "lzw")
print(p_merged)
```

# Significance Testing

## Assumption Checks

```{r assumptions}
# shapiro_test() errors on any group with fewer than 3 non-missing values, and
# is undefined for a group with zero variance — drop both before testing and
# report which groups were skipped instead of failing the render.
shapiro_input <- df_clean %>%
  filter(!is.na(logTiter)) %>%
  group_by(ShortGroup) %>%
  filter(dplyr::n() >= 3, sd(logTiter, na.rm = TRUE) > 0) %>%
  ungroup()

skipped_groups <- setdiff(
  as.character(unique(df_clean$ShortGroup)),
  as.character(unique(shapiro_input$ShortGroup))
)

shapiro_res <- shapiro_input %>%
  group_by(ShortGroup) %>%
  shapiro_test(logTiter)

shapiro_res %>%
  dplyr::select(Group = ShortGroup, `Statistic (W)` = statistic, `p-value` = p) %>%
  knitr::kable(caption = "Shapiro-Wilk Normality Test by Group", digits = 4)

if (length(skipped_groups) > 0) {
  cat("Groups skipped (n < 3 or zero variance on the log scale): ",
      paste(skipped_groups, collapse = ", "), "\n", sep = "")
}

lev_res <- leveneTest(logTiter ~ ShortGroup, data = df_clean)
lev_df  <- data.frame(
  `Source` = c("Group", "Residuals"),
  `Df`     = lev_res$Df,
  `F value` = c(lev_res$`F value`[1], NA),
  `Pr(>F)` = c(lev_res$`Pr(>F)`[1], NA),
  check.names = FALSE
)
knitr::kable(lev_df, caption = "Levene's Test for Homogeneity of Variance", digits = 4)
```

```{r assumption-verdict}
# Act on the two tables above rather than just printing them: state the verdict
# explicitly, and run Welch's ANOVA as a robustness check whenever the
# equal-variance assumption behind the classical aov()/Tukey below is rejected.
normality_ok  <- nrow(shapiro_res) == 0 || all(shapiro_res$p >= ALPHA)
levene_p      <- lev_res$`Pr(>F)`[1]
variances_ok  <- is.na(levene_p) || levene_p >= ALPHA

cat(sprintf(
  "Normality (Shapiro-Wilk, per group, alpha = %.2f): %s\n",
  ALPHA,
  if (normality_ok) "not rejected in any group"
  else paste0("REJECTED in: ",
              paste(shapiro_res$ShortGroup[shapiro_res$p < ALPHA], collapse = ", "))
))
cat(sprintf(
  "Homogeneity of variance (Levene, alpha = %.2f): %s (p = %s)\n",
  ALPHA,
  if (variances_ok) "not rejected" else "REJECTED",
  format.pval(levene_p, digits = 3)
))

if (!variances_ok) {
  cat("\nLevene's test is significant, so the classical ANOVA + Tukey HSD below",
      "assume more than the data support. Welch's one-way ANOVA, which does not",
      "assume equal variances, is reported here as the robustness check:\n\n")
  welch_full <- oneway.test(logTiter ~ ShortGroup, data = df_clean, var.equal = FALSE)
  print(welch_full)
  cat("\nIf the two disagree, prefer the Welch result and switch the pairwise",
      "comparisons to Games-Howell (rstatix::games_howell_test).\n")
} else {
  cat("\nEqual variances are a reasonable assumption, so the classical one-way",
      "ANOVA with Tukey HSD below is the appropriate test.\n")
}
```

## One-Way ANOVA — All Groups

```{r anova-full}
fit_full <- aov(logTiter ~ ShortGroup, data = df_clean)

tidy(fit_full) %>%
  gt() %>%
  tab_header(title = "ANOVA: log\u2081\u2080(Titer) by Group") %>%
  fmt_number(columns = c(statistic, p.value), decimals = 3) %>%
  cols_label(term = "Source", df = "df", sumsq = "Sum Sq",
             meansq = "Mean Sq", statistic = "F value", p.value = "Pr(>F)")

tukey_raw <- TukeyHSD(fit_full, conf.level = 0.95)$ShortGroup
tukey_tbl <- as.data.frame(tukey_raw)
tukey_tbl$comparison <- rownames(tukey_tbl)
rownames(tukey_tbl) <- NULL
tukey_tbl <- tukey_tbl[, c("comparison", "diff", "lwr", "upr", "p adj")]
names(tukey_tbl)[5] <- "p_adj"

tukey_tbl %>%
  gt() %>%
  tab_header(title = "Tukey HSD: Pairwise Comparisons") %>%
  fmt_number(columns = c(diff, lwr, upr, p_adj), decimals = 3) %>%
  cols_label(comparison = "Comparison", diff = "Difference",
             lwr = "Lower CI", upr = "Upper CI", p_adj = "Adjusted p")
```

## One-Way ANOVA — Combined Groups

```{r anova-merged}
fit_merged <- aov(logTiter ~ MergedGroup, data = df_clean)

tidy(fit_merged) %>%
  gt() %>%
  tab_header(title = "ANOVA: log\u2081\u2080(Titer) by Merged Group") %>%
  fmt_number(columns = c(statistic, p.value), decimals = 3) %>%
  cols_label(term = "Source", df = "df", sumsq = "Sum Sq",
             meansq = "Mean Sq", statistic = "F value", p.value = "Pr(>F)")

tukey_m_raw <- TukeyHSD(fit_merged, conf.level = 0.95)$MergedGroup
tukey_m_tbl <- as.data.frame(tukey_m_raw)
tukey_m_tbl$comparison <- rownames(tukey_m_tbl)
rownames(tukey_m_tbl) <- NULL
tukey_m_tbl <- tukey_m_tbl[, c("comparison", "diff", "lwr", "upr", "p adj")]
names(tukey_m_tbl)[5] <- "p_adj"

tukey_m_tbl %>%
  gt() %>%
  tab_header(title = "Tukey HSD: Merged Group Pairwise Comparisons") %>%
  fmt_number(columns = c(diff, lwr, upr, p_adj), decimals = 3) %>%
  cols_label(comparison = "Comparison", diff = "Difference",
             lwr = "Lower CI", upr = "Upper CI", p_adj = "Adjusted p")
```

# Violin Plot with Significance Brackets

The brackets below carry the **Tukey HSD adjusted p-values from the merged-group
ANOVA above** — the same multiplicity correction as the table, rather than a
second, uncorrected set of pairwise t-tests. Comparisons that are not
significant at `alpha` are omitted.

```{r violin-sig, fig.height=5.5, fig.width=6, message=FALSE, warning=FALSE}
df_plot <- df_clean %>% filter(!is.na(Titer) & Titer >= 0)

# ── Pull the Tukey adjusted p-value for each configured comparison ───────────
# TukeyHSD names its rows "B-A"; look the pair up in either orientation so the
# brackets can never disagree with the table printed above.
tukey_p_lookup <- setNames(tukey_m_tbl$p_adj, tukey_m_tbl$comparison)

tukey_p_for <- function(g1, g2) {
  for (key in c(paste0(g1, "-", g2), paste0(g2, "-", g1))) {
    if (key %in% names(tukey_p_lookup)) return(unname(tukey_p_lookup[[key]]))
  }
  NA_real_
}

p_signif_label <- function(p) {
  dplyr::case_when(
    is.na(p)    ~ NA_character_,
    p <= 1e-4   ~ "****",
    p <= 1e-3   ~ "***",
    p <= 1e-2   ~ "**",
    p <= 0.05   ~ "*",
    TRUE        ~ "ns"
  )
}

# Bracket heights are given on the raw (untransformed) titer scale because
# scale_y_log10() transforms them along with the data.
y_top   <- max(df_plot$Titer_pseudo, na.rm = TRUE)
sig_tbl <- do.call(rbind, lapply(COMPARISONS, function(cmp) {
  data.frame(
    group1 = cmp[1],
    group2 = cmp[2],
    p_adj  = tukey_p_for(cmp[1], cmp[2]),
    stringsAsFactors = FALSE
  )
}))
sig_tbl$label <- p_signif_label(sig_tbl$p_adj)
sig_tbl <- sig_tbl[!is.na(sig_tbl$p_adj) & sig_tbl$p_adj < ALPHA, , drop = FALSE]
if (nrow(sig_tbl) > 0) {
  sig_tbl$y.position <- y_top * 3 * (2.6 ^ (seq_len(nrow(sig_tbl)) - 1))
}

p_combined <- ggplot(df_plot, aes(x = MergedGroup, y = Titer_pseudo, fill = MergedGroup)) +
  geom_violin(trim = FALSE, alpha = 0.45, color = "black", bw = VIOLIN_BW) +
  geom_jitter(width = 0.18, size = 1, alpha = 0.60) +
  scale_fill_manual(values = PALETTE_MERGED, guide = "none") +
  scale_y_log10(labels = scales::label_scientific()) +
  expand_limits(y = y_top * if (nrow(sig_tbl) > 0) 3 * 2.6 ^ nrow(sig_tbl) else 5) +
  labs(
    x       = "Group",
    y       = paste0("Viral Titer + ", PSEUDOCOUNT, " (pfu/mL, log scale)"),
    title   = "Viral Titers by Combined Group",
    caption = paste0("Brackets show Tukey HSD adjusted p-values (alpha = ",
                     ALPHA, "); non-significant comparisons are not drawn.")
  ) +
  theme_bw(base_size = 13) +
  theme(
    axis.text.x      = element_text(angle = 30, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

if (nrow(sig_tbl) > 0) {
  p_combined <- p_combined +
    ggpubr::stat_pvalue_manual(
      sig_tbl,
      label        = "label",
      size         = 5,
      bracket.size = 0.6
    )
}

ggsave(file.path(OUTPUT_DIR, "Viral_Titer_CombinedGroups_sig.tiff"), plot = p_combined,
       width = FIG_WIDTH, height = FIG_HEIGHT + 0.5, units = "in", dpi = FIG_DPI,
       compression = "lzw")
print(p_combined)
```

```{r violin-sig-table}
sig_report <- do.call(rbind, lapply(COMPARISONS, function(cmp) {
  p <- tukey_p_for(cmp[1], cmp[2])
  data.frame(
    Comparison     = paste(cmp[1], "vs", cmp[2]),
    `Tukey adj. p` = p,
    Significance   = p_signif_label(p),
    Bracket        = if (!is.na(p) && p < ALPHA) "drawn" else "omitted (ns)",
    check.names    = FALSE,
    stringsAsFactors = FALSE
  )
}))
knitr::kable(
  sig_report,
  caption = "Configured comparisons and the Tukey HSD adjusted p-values used for the brackets",
  digits  = 4
)
```
