---
title: "Jonsson Lab Bioinformatics Templates"
subtitle: "A gallery of reusable analysis templates for virology and genomics research"
page-layout: full
toc: false
---

```{r}
#| label: setup
#| include: false
library(dplyr)
```

## About

This repository contains **8 standardized analysis templates** built from real lab analyses.
Each template:

- Ships with **synthetic example data** so it runs out of the box
- Preserves the **exact plot style** (themes, colors, dimensions) used in our publications
- Has a **`## ── USER CONFIGURATION ──`** block at the top — change those variables, nothing else
- Can be downloaded and dropped into any new project

---

## Template Gallery

```{r}
#| label: gallery-table
#| echo: false
library(DT)

templates <- data.frame(
  `#` = c("01", "02", "03", "04", "05", "06", "07", "08"),
  Template = c(
    "[Plaque Assay + Violin Plots](templates/01_plaque-assay-violin/template.qmd)",
    "[Image Infection + Dose-Response](templates/02_image-infection-dose-response/template.qmd)",
    "[Multi-Condition ANOVA](templates/03_multicondition-anova/template.qmd)",
    "[MagPix / Luminex Multiplex](templates/04_magpix-luminex/template.qmd)",
    "[Phylo-Geographic Analysis](templates/05_phylo-geographic/template.qmd)",
    "[GO Enrichment Analysis](templates/06_go-enrichment/template.qmd)",
    "[RNA-seq / DESeq2](templates/07_rnaseq-deseq2/template.qmd)",
    "[VCF Mutation Analysis](templates/08_vcf-mutation-analysis/template.qmd)"
  ),
  Category = c(
    "Infection & Virology", "Infection & Virology", "Infection & Virology",
    "Immunology", "Phylogenetics", "Genomics", "Transcriptomics", "Genomics"
  ),
  Language = c("R", "R", "R", "R", "R", "R", "R", "Python"),
  `Key Statistics` = c(
    "One-way ANOVA, Tukey HSD",
    "4PL dose-response, EC50/CC50",
    "Welch ANOVA, Holm correction",
    "Two-way ANOVA, emmeans, FDR",
    "Mantel test, geographic distance",
    "clusterProfiler, GO/KEGG",
    "DESeq2, volcano, PCA, bubble plots",
    "Entropy, Kruskal-Wallis"
  ),
  `Key Plot Types` = c(
    "Violin + jitter, log10 titer",
    "Ridge, density gate, dose-response curve",
    "Bar + error, fold-change table",
    "Dot, bar, violin, box, heatmap",
    "Scatter pie, Mantel correlogram, map",
    "Dot plot, bar, upset, Venn",
    "Volcano, bubble, PCA, triangle",
    "Mutation freq bar, entropy heatmap"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

datatable(
  templates,
  escape    = FALSE,
  rownames  = FALSE,
  options   = list(
    pageLength = 10,
    dom        = "ft",
    columnDefs = list(list(width = "20%", targets = 1))
  )
)
```

---

## How to Use a Template

1. **Browse** the gallery — click any template to see the rendered analysis with example plots.
2. **Download** the `.Rmd` or `.ipynb` file (use the *</> Code* button at the top right of each page, or download directly from GitHub).
3. **Edit the config block** at the top of the file:

```r
## ── USER CONFIGURATION ──────────────────────────────────────────────────────
DATA_FILE   <- "your_data.csv"   # path to your input file
GROUP_COL   <- "Treatment"       # column name for experimental groups
TITER_COL   <- "Titer_pfuml"    # column name for the measurement
# ... (all other parameters are documented inline)
## ────────────────────────────────────────────────────────────────────────────
```

4. **Run** — everything below the config block executes without modification.

---

## Setup

### R Templates (01–07)

```r
# Install all required packages at once
source("install_packages.R")
```

### Python Template (08)

```bash
conda env create -f environment.yml
conda activate jonsson-bioinfo
```

---

## Repository Structure

```
lab-bioinfo-templates/
├── templates/
│   ├── 01_plaque-assay-violin/
│   │   ├── template.Rmd      ← the template you copy
│   │   └── data/simulate_data.R
│   ├── 02_image-infection-dose-response/  ...
│   └── ...
├── _quarto.yml               ← Quarto site config
├── install_packages.R        ← one-shot R installer
└── environment.yml           ← Python conda env
```
