---
title: "Jonsson Lab Bioinformatics Templates"
subtitle: "Reusable analysis templates for virology and genomics research"
page-layout: full
toc: false
---

```{r}
#| label: setup
#| include: false
library(dplyr)
```

**8 standardized templates** built from real lab analyses. Each ships with synthetic example
data, preserves exact publication plot styles, and has a single `## ── USER CONFIGURATION ──`
block at the top — change those variables, run everything else unchanged.

---

::: {.gallery-section-heading}
Infection & Virology
:::

::: {.card-grid}

::: {.template-card .card-virology}
::: {.card-body}
#### 01 · Plaque Assay + Violin Plots
[R]{.badge-r}

Replicate QC (Limits of Agreement), one-way ANOVA, Tukey HSD, and violin plots for plaque assay titres.

[Open Template →](templates/01_plaque-assay-violin/template.html){.btn .btn-primary}
:::
:::

::: {.template-card .card-virology}
::: {.card-body}
#### 02 · Image Infection + Dose-Response
[R]{.badge-r}

CQ1-style gating, infection thresholding, EC50/CC50 estimation, and dose-response curves.

[Open Template →](templates/02_image-infection-dose-response/template.html){.btn .btn-primary}
:::
:::

::: {.template-card .card-virology}
::: {.card-body}
#### 03 · Multi-Condition ANOVA
[R]{.badge-r}

Welch ANOVA with planned contrasts and Holm correction across conditions or cell types.

[Open Template →](templates/03_multicondition-anova/template.html){.btn .btn-primary}
:::
:::

:::

::: {.gallery-section-heading}
Immunology
:::

::: {.card-grid}

::: {.template-card .card-immuno}
::: {.card-body}
#### 04 · MagPix / Luminex Multiplex
[R]{.badge-r}

Two-way ANOVA, emmeans contrasts, FDR correction, and cytokine visualization panels.

[Open Template →](templates/04_magpix-luminex/template.html){.btn .btn-primary}
:::
:::

:::

::: {.gallery-section-heading}
Phylogenetics
:::

::: {.card-grid}

::: {.template-card .card-phylo}
::: {.card-body}
#### 05 · Phylo-Geographic Analysis
[R]{.badge-r}

Isolation-by-distance testing, scatter pie maps, and Mantel correlograms for pathogen datasets.

[Open Template →](templates/05_phylo-geographic/template.html){.btn .btn-primary}
:::
:::

:::

::: {.gallery-section-heading}
Genomics & Transcriptomics
:::

::: {.card-grid}

::: {.template-card .card-genomics}
::: {.card-body}
#### 06 · GO Enrichment Analysis
[R]{.badge-r}

GO and KEGG enrichment across comparisons with bubble plots and export tables.

[Open Template →](templates/06_go-enrichment/template.html){.btn .btn-primary}
:::
:::

::: {.template-card .card-genomics}
::: {.card-body}
#### 07 · RNA-seq / DESeq2
[R]{.badge-r}

DESeq2 from counts or SummarizedExperiment — volcano, PCA, bubble plots, and heatmaps.

[Open Template →](templates/07_rnaseq-deseq2/template.html){.btn .btn-primary}
:::
:::

::: {.template-card .card-genomics}
::: {.card-body}
#### 08 · VCF Mutation Analysis
[Python]{.badge-python}

Mutation-frequency summaries and Shannon entropy heatmaps from per-sample VCF files.

[Open Template →](templates/08_vcf-mutation-analysis/template.html){.btn .btn-primary}
:::
:::

:::

---

## Template Catalog

```{r}
#| label: gallery-table
#| echo: false
library(DT)

templates <- data.frame(
  `#` = c("01", "02", "03", "04", "05", "06", "07", "08"),
  Template = c(
    "<a href='templates/01_plaque-assay-violin/template.html'>Plaque Assay + Violin Plots</a>",
    "<a href='templates/02_image-infection-dose-response/template.html'>Image Infection + Dose-Response</a>",
    "<a href='templates/03_multicondition-anova/template.html'>Multi-Condition ANOVA</a>",
    "<a href='templates/04_magpix-luminex/template.html'>MagPix / Luminex Multiplex</a>",
    "<a href='templates/05_phylo-geographic/template.html'>Phylo-Geographic Analysis</a>",
    "<a href='templates/06_go-enrichment/template.html'>GO Enrichment Analysis</a>",
    "<a href='templates/07_rnaseq-deseq2/template.html'>RNA-seq / DESeq2</a>",
    "<a href='templates/08_vcf-mutation-analysis/template.html'>VCF Mutation Analysis</a>"
  ),
  Category = c(
    "Infection & Virology", "Infection & Virology", "Infection & Virology",
    "Immunology", "Phylogenetics", "Genomics", "Transcriptomics", "Genomics"
  ),
  Language = c("R", "R", "R", "R", "R", "R", "R", "Python"),
  `Key Statistics` = c(
    "One-way ANOVA, Tukey HSD",
    "4PL dose-response, EC50/CC50",
    "Welch ANOVA, Holm correction",
    "Two-way ANOVA, emmeans, FDR",
    "Mantel test, geographic distance",
    "clusterProfiler, GO/KEGG",
    "DESeq2, Wald test, FDR",
    "Shannon entropy, Kruskal-Wallis"
  ),
  `Key Plots` = c(
    "Violin + jitter, log10 titer",
    "Ridge, density gate, dose-response curve",
    "Bar + error, fold-change table",
    "Dot, bar, violin, heatmap",
    "Scatter pie, Mantel correlogram, map",
    "Dot plot, bar, Venn",
    "Volcano, PCA, heatmap, bubble",
    "Mutation freq bar, entropy heatmap"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

datatable(
  templates,
  escape    = FALSE,
  rownames  = FALSE,
  options   = list(
    pageLength = 10,
    dom        = "ft",
    columnDefs = list(list(width = "20%", targets = 1))
  )
)
```

---

## How to Use

1. Click a card above to see the rendered analysis with example plots
2. Download the `.Rmd` or `.ipynb` via the *</> Code* button (top-right of each page) or from GitHub
3. Edit only the `## ── USER CONFIGURATION ──` block at the top of the file
4. Run — everything below executes without modification

**Install R packages once:**
```r
source("install_packages.R")
```
