---
title: "05 · Phylo-Geographic Analysis"
subtitle: "Isolation by Distance (Mantel Test), Scatter Pie Map, and Mantel Correlogram"
description: |
  Performs isolation-by-distance analysis using Mantel tests on patristic distances
  from a phylogenetic tree and geographic (Haversine) distances. Produces a scatter
  pie map showing clade composition per sampling location and a Mantel correlogram.
categories: [phylogenetics, geographic, mantel, virology]
---

## Overview

| Item | Details |
|------|---------|
| **Input** | Aligned FASTA (S segment required, M optional) · Newick tree · coordinates CSV · metadata CSV |
| **Key packages** | `ape`, `vegan`, `geosphere`, `scatterpie`, `sf`, `rnaturalearth`, `ggspatial` |
| **Statistics** | Mantel test (Pearson) · patristic distances · Haversine geographic distances |
| **Output** | Isolation-by-distance plot · scatter pie map · Mantel correlogram · TIFF + SVG |
| **Download** | [template.Rmd](template.Rmd) |

::: {.callout-tip}
**When to use this template:** You have a phylogenetic tree and sampling coordinates
for a pathogen, and want to test whether genetic distance correlates with geographic
distance (isolation by distance) and visualize the geographic distribution of clades.
:::

::: {.callout-note}
**Clade colors:** Default palette uses Okabe-Ito colorblind-safe colors.
Update `CLADE_COLORS` in the USER CONFIGURATION block to match your clade assignments.
The scatter pie radius (`radius` column in `pie_data`) is in degrees — adjust `mutate(radius = 0.4)`
to scale pies appropriately for your map extent.
:::

---

```{r}
#| label: simulate
#| include: false
source("data/simulate_data.R")
```

```{r child="template.Rmd"}
```
