---
title: "Image-Based Infection + Dose-Response Analysis"
author: "Jonsson Laboratory"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

<style>
.main-container table, table.table {
  border: 1.5px solid #2c3e50 !important;
  border-collapse: collapse !important;
  margin-top: 12px;
  margin-bottom: 24px;
  width: 100%;
}
.main-container th, .main-container td, table.table th, table.table td {
  border: 1px solid #bdc3c7 !important;
  padding: 8px 12px !important;
}
.main-container th, table.table th {
  background-color: #ecf0f1 !important;
  border-bottom: 2px solid #2c3e50 !important;
  color: #2c3e50;
  font-weight: bold;
}
.main-container tr:nth-child(even), table.table tr:nth-child(even) {
  background-color: #f8f9fa;
}
.main-container tr:hover, table.table tr:hover {
  background-color: #eaf2f8;
}
</style>

```{r user-config, message=FALSE, warning=FALSE}
## ── USER CONFIGURATION ──────────────────────────────────────────────────────
#
# Input files (CQ1 CSV exports per plate, nucleus and virus channels):
PLATE1_NUCLEUS_FILE <- "data/plate1_nucleus.csv"
PLATE1_VIRUS_FILE   <- "data/plate1_virus.csv"

PLATE2_NUCLEUS_FILE <- "data/plate2_nucleus.csv"
PLATE2_VIRUS_FILE   <- "data/plate2_virus.csv"

PLATE3_NUCLEUS_FILE <- "data/plate3_nucleus.csv"
PLATE3_VIRUS_FILE   <- "data/plate3_virus.csv"

# Column names in nuclei CSV:
NUCLEUS_DAPI_COL    <- "(Nucleus) MeanIntensity CH1"   # Nuclear DAPI stain
NUCLEUS_AREA_COL    <- "(Nucleus) Area"                # Nuclear area (2D)
VIRUS_INTENSITY_COL <- "(Virus) MeanIntensity CH2"     # NP infection channel

# Study labels
VIRUS_NAME    <- "ANDV"
COMPOUND_NAME <- "Compound"
DOSE_UNIT     <- "\u00b5M"

# Cell gating thresholds
AREA_MIN        <- 30     # Floor to retain pyknotic cells
INTENSITY_MIN   <- 300    # Hard DAPI intensity floor
CIRCULARITY_MIN <- 0.40   # Doublet and fragment cutoff

# Data Exclusions
DOSES_TO_EXCLUDE <- c(0.75)                      # Dose concentrations to exclude
WELLS_TO_EXCLUDE <- c("P2_C3", "P2_B4", "P2_C4") # Failing or artifact wells

# Plate Layout Maps
PLATE1_MAP <- tibble::tribble(
  ~WellName,   ~Treatment,            ~Concentration_uM,
  "P1_B2",  "Virus Only",          0,
  "P1_B3",  "Virus + Compound",    0.125,
  "P1_B4",  "Virus + Compound",    0.25,
  "P1_B5",  "Mock+Ab",             NA,
  "P1_C2",  "Virus Only",          0,
  "P1_C3",  "Virus + Compound",    0.125,
  "P1_C4",  "Virus + Compound",    0.25,
  "P1_C5",  "Mock+Ab",             NA
)

PLATE2_MAP <- tibble::tribble(
  ~WellName,   ~Treatment,            ~Concentration_uM,
  "P2_B2",  "Virus Only",          0,
  "P2_B3",  "Virus + Compound",    0.5,
  "P2_B4",  "Virus + Compound",    0.75,
  "P2_B5",  "Mock+Ab",             NA,
  "P2_C2",  "Virus Only",          0,
  "P2_C3",  "Virus + Compound",    0.5,
  "P2_C4",  "Virus + Compound",    0.75,
  "P2_C5",  "Mock+Ab",             NA
)

PLATE3_MAP <- tibble::tribble(
  ~WellName,   ~Treatment,            ~Concentration_uM,
  "P3_B2",  "Virus Only",          0,
  "P3_B3",  "Virus + Compound",    1,
  "P3_B4",  "Virus + Compound",    2,
  "P3_B5",  "Mock+Ab",             NA,
  "P3_C2",  "Virus Only",          0,
  "P3_C3",  "Virus + Compound",    1,
  "P3_C4",  "Virus + Compound",    2,
  "P3_C5",  "Mock+Ab",             NA
)

PLATE_MAP <- dplyr::bind_rows(PLATE1_MAP, PLATE2_MAP, PLATE3_MAP)
## ────────────────────────────────────────────────────────────────────────────
```

# Setup

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(scales)
library(ggridges)
library(ggpubr)
library(RColorBrewer)
library(gtools)
library(grid)
library(ragg)
library(patchwork)
library(mclust)
library(minpack.lm)
library(MASS)
```

```{r global-theme}
base_family <- "sans"

theme_set(
  theme_bw(base_size = 8, base_family = base_family) +
    theme(
      plot.title   = element_text(face = "bold", size = 10, margin = margin(b = 4)),
      axis.title   = element_text(face = "bold", size = 9),
      axis.text    = element_text(size = 8),
      legend.title = element_text(size = 8),
      legend.text  = element_text(size = 8),
      axis.line    = element_line(linewidth = 0.6),
      axis.ticks   = element_line(linewidth = 0.6),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
)

update_geom_defaults("text",     list(size = 8 / ggplot2::.pt))
update_geom_defaults("label",    list(size = 8 / ggplot2::.pt))
update_geom_defaults("point",    list(size = 1.8))
update_geom_defaults("errorbar", list(linewidth = 0.5))
update_geom_defaults("line",     list(linewidth = 0.9))

dir.create("Plots", showWarnings = FALSE)
```

# Load & Annotate Data

```{r load-data, message=FALSE}
read_plate <- function(nucleus_file, virus_file, plate_prefix) {
  df_nuclei <- read_csv(nucleus_file, show_col_types = FALSE)
  df_virus  <- read_csv(virus_file,   show_col_types = FALSE)

  dapi_col <- names(df_nuclei)[grepl("MeanIntensity.*CH1", names(df_nuclei), ignore.case = TRUE)][1]
  area_col <- names(df_nuclei)[grepl("\\(Nucleus\\)\\s*(Volume|Area)", names(df_nuclei), ignore.case = TRUE)][1]
  sphericity_col <- names(df_nuclei)[grepl("Sphericity|Circularity|Compactness", names(df_nuclei), ignore.case = TRUE)][1]
  virus_col <- names(df_virus)[grepl("MeanIntensity.*CH2", names(df_virus), ignore.case = TRUE)][1]

  nuclei_clean <- df_nuclei %>%
    dplyr::select(
      WellName, FieldIndex, ObjectNumber,
      dapi_intensity = all_of(dapi_col),
      Area           = all_of(area_col),
      Sphericity     = all_of(sphericity_col)
    ) %>%
    dplyr::mutate(integrated_intensity = Area * dapi_intensity)

  virus_clean <- df_virus %>%
    dplyr::select(
      WellName, FieldIndex, ObjectNumber,
      virus_intensity = all_of(virus_col)
    )

  nuclei_clean %>%
    dplyr::left_join(virus_clean, by = c("WellName", "FieldIndex", "ObjectNumber")) %>%
    dplyr::mutate(
      WellName = paste0(plate_prefix, "_",
                        str_replace(str_to_upper(WellName),
                                    "^([A-D])-?0*([1-9][0-9]?).*$", "\\1\\2"))
    ) %>%
    tidyr::drop_na(virus_intensity)
}

df_p1 <- read_plate(PLATE1_NUCLEUS_FILE, PLATE1_VIRUS_FILE, "P1")
df_p2 <- read_plate(PLATE2_NUCLEUS_FILE, PLATE2_VIRUS_FILE, "P2")
df_p3 <- read_plate(PLATE3_NUCLEUS_FILE, PLATE3_VIRUS_FILE, "P3")

master_df <- dplyr::bind_rows(df_p1, df_p2, df_p3)

master_df_raw_annotated <- master_df %>%
  dplyr::left_join(PLATE_MAP, by = "WellName") %>%
  dplyr::filter(!is.na(Treatment))

master_df_annotated <- master_df_raw_annotated %>%
  dplyr::filter(!WellName %in% WELLS_TO_EXCLUDE) %>%
  { if (length(DOSES_TO_EXCLUDE) > 0)
      dplyr::filter(., !Concentration_uM %in% DOSES_TO_EXCLUDE)
    else . }
```

# Cell Gating (QC)

```{r gating-plot, fig.height=4, fig.width=9, warning=FALSE, message=FALSE}
master_df_pre <- master_df_annotated %>%
  dplyr::filter(Area > AREA_MIN & dapi_intensity > INTENSITY_MIN) %>%
  dplyr::mutate(Is_Singlet = Sphericity > CIRCULARITY_MIN)

master_df_singlets <- master_df_pre %>% dplyr::filter(Is_Singlet)

vars <- c("Area", "integrated_intensity")
center <- colMeans(master_df_singlets[vars])
cov_mat <- cov(master_df_singlets[vars])
chi_cutoff <- qchisq(0.95, df = 2)

master_df_pre <- master_df_pre %>%
  dplyr::mutate(
    dist_cc = if_else(Is_Singlet, 
                      mahalanobis(cbind(Area, integrated_intensity), center = center, cov = cov_mat), 
                      NA_real_),
    Keep_Cell = Is_Singlet & (dist_cc <= chi_cutoff)
  )

master_df_gated <- master_df_pre %>% dplyr::filter(Keep_Cell)
```

# Infection Gate: Control-Based Normalization + Population GMM

```{r infection-gate, warning=FALSE}
VIRUS_LABEL_VO  <- "Virus Only"
master_df_gated <- master_df_gated %>%
  mutate(Plate = stringr::str_extract(WellName, "^P[123]"))

gmm_threshold <- function(mu, sdv, pi_) {
  o <- order(mu); mu <- mu[o]; sdv <- sdv[o]; pi_ <- pi_[o]
  sig1_sq <- sdv[1]^2; sig2_sq <- sdv[2]^2
  A <- 1/(2*sig1_sq) - 1/(2*sig2_sq)
  B <- -mu[1]/sig1_sq + mu[2]/sig2_sq
  C <- mu[1]^2/(2*sig1_sq) - mu[2]^2/(2*sig2_sq) + log(sdv[1]/sdv[2]) - log(pi_[1]/pi_[2])
  disc <- B^2 - 4*A*C
  if (disc >= 0 && abs(A) > 1e-10) {
    roots <- c((-B + sqrt(disc))/(2*A), (-B - sqrt(disc))/(2*A))
    r_valid <- roots[roots > mu[1] & roots < mu[2]]
    if (length(r_valid) > 0) return(r_valid[1])
  }
  (mu[1] + mu[2]) / 2
}

fit_high_mode <- function(xlog) {
  fit <- tryCatch(Mclust(xlog, G = 2, modelNames = "V", verbose = FALSE), error = function(e) NULL)
  if (!is.null(fit)) return(max(as.numeric(fit$parameters$mean)))
  as.numeric(quantile(xlog, 0.75, na.rm = TRUE))
}

# 1. Compute Anchors (L_p, H_p)
mock_cells <- master_df_gated %>% filter(Treatment == "Mock+Ab") %>% mutate(xlog = log10(virus_intensity + 1))
L_p_table  <- mock_cells %>% group_by(Plate, WellName) %>% summarise(well_L = median(xlog, na.rm = TRUE), .groups = "drop") %>%
  group_by(Plate) %>% summarise(L_p = mean(well_L, na.rm = TRUE), .groups = "drop")

vo_cells   <- master_df_gated %>% filter(Treatment == VIRUS_LABEL_VO) %>% mutate(xlog = log10(virus_intensity + 1))
H_p_table  <- vo_cells %>% group_by(Plate, WellName) %>% summarise(well_H = fit_high_mode(xlog), .groups = "drop") %>%
  group_by(Plate) %>% summarise(H_p = mean(well_H, na.rm = TRUE), .groups = "drop")

anchors <- left_join(L_p_table, H_p_table, by = "Plate")

# 2. Normalize Cells
master_df_norm <- master_df_gated %>%
  mutate(xlog = log10(virus_intensity + 1)) %>%
  left_join(anchors, by = "Plate") %>%
  mutate(
    span   = H_p - L_p,
    x_norm = if_else(is.na(span) | span < 1e-6, NA_real_, (xlog - L_p) / span)
  )

# 3. Fit Population 2-GMM Gate
infected_culture_norm <- master_df_norm %>% filter(Treatment %in% c(VIRUS_LABEL_VO, "Virus + Compound"))
target_n_gate <- as.integer(round(median(table(infected_culture_norm$WellName))))

set.seed(42)
gate_balanced <- infected_culture_norm %>%
  group_by(WellName) %>%
  slice_sample(n = target_n_gate) %>%
  ungroup()

x_gate <- gate_balanced$x_norm[is.finite(gate_balanced$x_norm)]
fit_gate <- Mclust(x_gate, G = 2, modelNames = "V", verbose = FALSE)

mu_gate  <- as.numeric(fit_gate$parameters$mean)
sdv_gate <- sqrt(fit_gate$parameters$variance$sigmasq)
pi_gate  <- as.numeric(fit_gate$parameters$pro)
if (length(sdv_gate) == 1L) sdv_gate <- rep(sdv_gate, 2L)

threshold_final <- gmm_threshold(mu_gate, sdv_gate, pi_gate)
master_df_final <- master_df_norm
```

# Percent Infection per Well

```{r percent-infection}
total_counts <- master_df_final %>% dplyr::count(Treatment, Concentration_uM, WellName, name = "total_cell_count")
positive_counts <- master_df_final %>% filter(x_norm > threshold_final) %>% dplyr::count(Treatment, Concentration_uM, WellName, name = "positive_cell_count")

percent_summary <- total_counts %>%
  left_join(positive_counts, by = c("Treatment", "Concentration_uM", "WellName")) %>%
  mutate(
    positive_cell_count = replace_na(positive_cell_count, 0),
    percent_positive     = (positive_cell_count / total_cell_count) * 100,
    Plate = case_when(
      grepl("^P1_", WellName) ~ "Plate 1",
      grepl("^P2_", WellName) ~ "Plate 2",
      grepl("^P3_", WellName) ~ "Plate 3"
    ),
    Precision_Flag = if_else(total_cell_count < 500, "Low Precision (n<500)", "Normal")
  )

percent_summary %>%
  mutate(
    Well_Clean = stringr::str_remove(WellName, "^P[123]_"),
    Dose_Label = case_when(
      Treatment == "Mock+Ab"    ~ "Mock",
      Treatment == "Virus Only" ~ "0 µM (Control)",
      TRUE                      ~ paste0(Concentration_uM, " µM")
    ),
    Status_Label = case_when(
      Precision_Flag == "Normal" ~ "Valid",
      TRUE                        ~ Precision_Flag
    )
  ) %>%
  dplyr::select(
    Plate,
    Well              = Well_Clean,
    Treatment,
    Dose              = Dose_Label,
    `Total Cells`     = total_cell_count,
    `Infected Cells`  = positive_cell_count,
    `% Infection`    = percent_positive,
    `QC Status`       = Status_Label
  ) %>%
  knitr::kable(
    caption     = "Well-Level Percent Infection (Global Population GMM Threshold)",
    digits      = 1,
    format.args = list(big.mark = ",")
  )
```

# Co-Primary Endpoint 1: % Infection Dose-Response (4PL)

```{r dose-response-model}
plate_anchors <- percent_summary %>%
  group_by(Plate) %>%
  summarise(
    vo_anchor   = mean(percent_positive[Treatment == "Virus Only"], na.rm = TRUE),
    mock_anchor = mean(percent_positive[Treatment == "Mock+Ab"], na.rm = TRUE),
    mock_n      = sum(total_cell_count[Treatment == "Mock+Ab"], na.rm = TRUE),
    .groups     = "drop"
  )

global_mock_fallback <- mean(percent_summary$percent_positive[percent_summary$Treatment == "Mock+Ab"], na.rm = TRUE)
if (is.nan(global_mock_fallback) || is.na(global_mock_fallback)) global_mock_fallback <- 0

summary_rel <- percent_summary %>%
  left_join(plate_anchors, by = "Plate") %>%
  mutate(
    effective_mock = if_else(!is.na(mock_n) & mock_n >= 200, mock_anchor, global_mock_fallback),
    percent_infection_rel = 100 * (percent_positive - effective_mock) / (vo_anchor - effective_mock)
  )

drc_data_rel <- summary_rel %>% filter(Treatment == "Virus + Compound")
min_pos   <- min(drc_data_rel$Concentration_uM, na.rm = TRUE)
zero_tick <- min_pos / 10

# Binomial-weighted fit dataset (0.5 µM outlier masked from fit)
fit_data_wide <- summary_rel %>%
  filter(Treatment == "Virus Only" | (Treatment == "Virus + Compound" & Concentration_uM != 0.5)) %>%
  mutate(
    dose_fit   = ifelse(Treatment == "Virus Only", zero_tick, Concentration_uM),
    prop_raw   = positive_cell_count / total_cell_count,
    prop_clamp = pmin(pmax(prop_raw, 1e-3), 1 - 1e-3),
    bin_weight = total_cell_count / (prop_clamp * (1 - prop_clamp))
  )

bootstrap_ec50_ci <- function(fit, R = 4000, seed = 42) {
  if (is.null(fit)) return(c(NA_real_, NA_real_, NA_real_))
  mu <- coef(fit)
  vc <- tryCatch(vcov(fit), error = function(e) NULL)
  if (is.null(vc)) return(c(NA_real_, NA_real_, NA_real_))
  set.seed(seed)
  sim <- tryCatch(MASS::mvrnorm(R, mu, vc), error = function(e) NULL)
  if (is.null(sim)) return(c(NA_real_, NA_real_, NA_real_))
  ec <- sim[, "ec50"]
  ec <- ec[is.finite(ec) & ec > 0]
  if (length(ec) < 10) return(c(NA_real_, NA_real_, NA_real_))
  c(lower = unname(quantile(ec, 0.025)), median = unname(quantile(ec, 0.50)), upper = unname(quantile(ec, 0.975)))
}

fit_rel <- NULL
for (ec50_try in c(0.125, 0.25, 0.5, 1, 2, 5, 0.05)) {
  fit_rel <- tryCatch(
    minpack.lm::nlsLM(
      percent_infection_rel ~ bottom + (top - bottom) / (1 + (dose_fit / ec50) ^ hill),
      data    = fit_data_wide,
      start   = list(ec50 = ec50_try, hill = 1.6, bottom = 50, top = 100),
      lower   = c(ec50 = 1e-4, hill = 0.05, bottom = -20, top = 80),
      upper   = c(ec50 = 1e3,  hill = 50, bottom = 95, top = 120),
      weights = fit_data_wide$bin_weight,
      control = minpack.lm::nls.lm.control(maxiter = 500)
    ),
    error   = function(e) NULL
  )
  if (!is.null(fit_rel)) break
}

ci_rel <- bootstrap_ec50_ci(fit_rel)
```

## Dose-Response Plot

```{r dose-response-plot, message=FALSE, warning=FALSE}
newx <- exp(seq(log(zero_tick), log(max(drc_data_rel$Concentration_uM, na.rm = TRUE)), length.out = 300))

pred_df_rel <- tibble(Concentration_uM = numeric(0), y = numeric(0))
if (!is.null(fit_rel)) {
  coef_rel   <- coef(fit_rel)
  ec50_rel   <- coef_rel["ec50"]
  hill_rel   <- coef_rel["hill"]
  bottom_rel <- coef_rel["bottom"]
  top_rel    <- coef_rel["top"]
  pred_df_rel <- tibble(
    Concentration_uM = newx,
    y = bottom_rel + (top_rel - bottom_rel) / (1 + (Concentration_uM / ec50_rel)^hill_rel)
  )
  ec50_rel_lab <- paste0("EC50 = ", signif(ec50_rel, 3), " ", DOSE_UNIT)
} else {
  ec50_rel_lab <- "Fit Failed"
}

pts_sum <- summary_rel %>%
  filter(Treatment %in% c("Virus Only", "Virus + Compound")) %>%
  mutate(x_plot = ifelse(Concentration_uM == 0, zero_tick, Concentration_uM)) %>%
  group_by(x_plot) %>%
  summarise(n  = n(),
            y  = mean(percent_infection_rel, na.rm = TRUE),
            se = ifelse(n > 1, sd(percent_infection_rel, na.rm = TRUE) / sqrt(n), 0),
            .groups = "drop") %>%
  mutate(y    = ifelse(x_plot == zero_tick, 100, y),
         ymin = pmax(0,   y - se),
         ymax = pmin(100, y + se))

dose_breaks     <- sort(unique(c(zero_tick, drc_data_rel$Concentration_uM)))
dose_labels_vec <- c("0", as.character(dose_breaks[-1]))

p_dr_rel <- ggplot() +
  geom_errorbar(data = pts_sum, aes(x = x_plot, ymin = ymin, ymax = ymax),
                width = 0.06, linewidth = 0.5, colour = "steelblue4") +
  geom_point(data = pts_sum, aes(x = x_plot, y = y),
             shape = 21, size = 2.6, stroke = 0.4,
             fill = "steelblue4", colour = "steelblue4")

if (!is.null(fit_rel)) {
  y_rel_halfway <- bottom_rel + (top_rel - bottom_rel) / 2
  p_dr_rel <- p_dr_rel + 
    geom_line(data = pred_df_rel, aes(x = Concentration_uM, y = y),
              color = "steelblue", linewidth = 0.9) +
    geom_segment(aes(x = ec50_rel, xend = ec50_rel, y = 0, yend = y_rel_halfway),
                 linetype = "dashed", color = "steelblue4", linewidth = 0.5) +
    geom_segment(aes(x = zero_tick, xend = ec50_rel, y = y_rel_halfway, yend = y_rel_halfway),
                 linetype = "dashed", color = "steelblue4", linewidth = 0.5) +
    annotate("label", x = ec50_rel, y = 25, label = ec50_rel_lab,
             fill = "steelblue4", color = "white", size = 3, label.size = 0)
}

p_dr_rel <- p_dr_rel +
  scale_x_log10(breaks = dose_breaks, labels = dose_labels_vec, minor_breaks = NULL) +
  annotation_logticks(sides = "b", short = unit(1.2, "mm"), mid = unit(1.6, "mm"), long = unit(2.2, "mm")) +
  labs(
    title = paste(COMPOUND_NAME, "relative dose\u2013response"),
    x     = paste0(COMPOUND_NAME, " (", DOSE_UNIT, ", log10)"),
    y     = "% infected cells"
  ) +
  coord_cartesian(ylim = c(0, 105), clip = "off") +
  theme_bw(base_size = 8) +
  theme(
    plot.title        = element_text(face = "bold", size = 9, margin = margin(b = 4)),
    axis.title        = element_text(face = "bold", size = 8),
    axis.text         = element_text(size = 7),
    axis.line         = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks        = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks.length = unit(2.2, "mm"),
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    legend.position   = "none",
    plot.margin       = margin(6, 10, 6, 8)
  )

print(p_dr_rel)
ggsave("Plots/compound_dose_response_relative.tiff", p_dr_rel,
       width = 4, height = 3, units = "in", dpi = 600, compression = "lzw", device = ragg::agg_tiff)
```

# Co-Primary Endpoint 2: Normalized NP Intensity Dose-Response

```{r intensity-dose-response, warning=FALSE}
intensity_well <- master_df_final %>%
  filter(Treatment %in% c("Mock+Ab", "Virus Only", "Virus + Compound")) %>%
  group_by(Plate, WellName, Treatment, Concentration_uM) %>%
  summarise(
    n_cells     = n(),
    mean_xlog   = mean(xlog, na.rm = TRUE),
    L_p = first(L_p), H_p = first(H_p),
    .groups = "drop"
  ) %>%
  mutate(intensity_norm_mean = 100 * (mean_xlog - L_p) / (H_p - L_p))

int_plate_anchors <- intensity_well %>%
  group_by(Plate) %>%
  summarise(
    vo_i   = mean(intensity_norm_mean[Treatment == "Virus Only"], na.rm = TRUE),
    mock_i = mean(intensity_norm_mean[Treatment == "Mock+Ab"], na.rm = TRUE),
    .groups = "drop"
  )

intensity_rel <- intensity_well %>%
  left_join(int_plate_anchors, by = "Plate") %>%
  mutate(intensity_rel_mean = 100 * (intensity_norm_mean - mock_i) / (vo_i - mock_i))

min_pos_i   <- min(intensity_rel$Concentration_uM[intensity_rel$Treatment == "Virus + Compound"], na.rm = TRUE)
zero_tick_i <- min_pos_i / 10

fit_i <- intensity_rel %>%
  filter(Treatment == "Virus Only" | (Treatment == "Virus + Compound" & Concentration_uM != 0.5)) %>%
  mutate(dose_fit = ifelse(Treatment == "Virus Only", zero_tick_i, Concentration_uM))

fit_int <- NULL
for (ec50_try in c(0.125, 0.25, 0.5, 1, 2, 5, 0.05)) {
  fit_int <- tryCatch(
    minpack.lm::nlsLM(
      intensity_rel_mean ~ bottom + (top - bottom) / (1 + (dose_fit / ec50) ^ hill),
      data    = fit_i,
      start   = list(ec50 = ec50_try, hill = 1.6, bottom = 50, top = 100),
      lower   = c(ec50 = 1e-4, hill = 0.05, bottom = -20, top = 80),
      upper   = c(ec50 = 1e3,  hill = 50, bottom = 95, top = 120),
      weights = fit_i$n_cells,
      control = minpack.lm::nls.lm.control(maxiter = 500)
    ),
    error = function(e) NULL
  )
  if (!is.null(fit_int)) break
}

ci_int <- bootstrap_ec50_ci(fit_int)

ec50_compare <- tibble(
  Endpoint   = c("Normalized NP intensity (mean) -- co-primary",
                 "% infected (population GMM gate) -- co-primary"),
  `EC50 (µM)` = c(
    if (!is.null(fit_int)) unname(coef(fit_int)["ec50"]) else NA_real_,
    if (!is.null(fit_rel))  unname(coef(fit_rel)["ec50"])  else NA_real_
  ),
  `95% CI`    = c(
    if (!is.null(fit_int)) paste0(signif(ci_int["lower"],3)," – ",signif(ci_int["upper"],3)) else "–",
    if (!is.null(fit_rel))  paste0(signif(ci_rel["lower"],3)," – ",signif(ci_rel["upper"],3))  else "–"
  )
)
knitr::kable(ec50_compare, caption = "EC50 comparison across co-primary endpoints.", digits = 3)
```

## Intensity Dose-Response Plot

```{r intensity-dose-response-plot, warning=FALSE, message=FALSE}
pts_i <- intensity_rel %>%
  filter(Treatment %in% c("Virus Only", "Virus + Compound")) %>%
  mutate(
    x_plot = ifelse(Treatment == "Virus Only", zero_tick_i, Concentration_uM),
    x_plot = ifelse(is.na(x_plot), NA_real_, x_plot)
  ) %>%
  filter(!is.na(x_plot)) %>%
  group_by(x_plot) %>%
  summarise(n  = n(),
            y  = mean(intensity_rel_mean, na.rm = TRUE),
            se = ifelse(n > 1, sd(intensity_rel_mean, na.rm = TRUE) / sqrt(n), 0),
            .groups = "drop") %>%
  mutate(y    = ifelse(x_plot == zero_tick_i, 100, y),
         ymin = pmax(0,   y - se),
         ymax = pmin(105, y + se))

newx_i <- exp(seq(log(zero_tick_i), log(max(intensity_rel$Concentration_uM[intensity_rel$Treatment == "Virus + Compound"], na.rm = TRUE)), length.out = 300))

pred_i <- tibble(Concentration_uM = numeric(0), y = numeric(0))
if (!is.null(fit_int)) {
  cf <- coef(fit_int)
  pred_i <- tibble(
    Concentration_uM = newx_i,
    y = cf["bottom"] + (cf["top"] - cf["bottom"]) / (1 + (Concentration_uM / cf["ec50"]) ^ cf["hill"])
  )
}

dose_breaks_i <- sort(unique(c(zero_tick_i, intensity_rel$Concentration_uM[intensity_rel$Treatment == "Virus + Compound"])))
dose_labels_i <- c("0", as.character(dose_breaks_i[-1]))

p_dr_int <- ggplot() +
  geom_errorbar(data = pts_i, aes(x = x_plot, ymin = ymin, ymax = ymax),
                width = 0.06, linewidth = 0.5, colour = "steelblue4") +
  geom_point(data = pts_i, aes(x = x_plot, y = y),
             shape = 21, size = 2.6, stroke = 0.4,
             fill = "steelblue4", colour = "steelblue4")

if (!is.null(fit_int)) {
  cf <- coef(fit_int)
  ec50_val  <- unname(cf["ec50"])
  halfway   <- unname(cf["bottom"] + (cf["top"] - cf["bottom"]) / 2)
  int_label <- paste0("EC50 = ", signif(ec50_val, 3), " ", DOSE_UNIT)
  p_dr_int  <- p_dr_int +
    geom_line(data = pred_i, aes(x = Concentration_uM, y = y),
              color = "steelblue", linewidth = 0.9) +
    geom_segment(aes(x = ec50_val, xend = ec50_val, y = 0, yend = halfway),
                 linetype = "dashed", color = "steelblue4", linewidth = 0.5) +
    geom_segment(aes(x = zero_tick_i, xend = ec50_val, y = halfway, yend = halfway),
                 linetype = "dashed", color = "steelblue4", linewidth = 0.5) +
    annotate("label", x = ec50_val, y = 25, label = int_label,
             fill = "steelblue4", color = "white", size = 3, label.size = 0)
}

p_dr_int <- p_dr_int +
  scale_x_log10(breaks = dose_breaks_i, labels = dose_labels_i, minor_breaks = NULL) +
  annotation_logticks(sides = "b", short = unit(1.2, "mm"), mid = unit(1.6, "mm"), long = unit(2.2, "mm")) +
  labs(
    title = paste(COMPOUND_NAME, "intensity dose\u2013response"),
    x     = paste0(COMPOUND_NAME, " (", DOSE_UNIT, ", log10)"),
    y     = "Normalized NP intensity (%)"
  ) +
  coord_cartesian(ylim = c(0, 105), clip = "off") +
  theme_bw(base_size = 8) +
  theme(
    plot.title        = element_text(face = "bold", size = 9, margin = margin(b = 4)),
    axis.title        = element_text(face = "bold", size = 8),
    axis.text         = element_text(size = 7),
    axis.line         = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks        = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks.length = unit(2.2, "mm"),
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    legend.position   = "none",
    plot.margin       = margin(6, 10, 6, 8)
  )

print(p_dr_int)
ggsave("Plots/compound_dose_response_intensity.tiff", p_dr_int,
       width = 4, height = 3, units = "in", dpi = 600, compression = "lzw", device = ragg::agg_tiff)
```

# Results Summary

Two co-primary endpoints are reported side-by-side:
1. **% infected** - fraction of cells above the population 2-GMM gate, fitted with binomial-weighted relative 4PL.
2. **Normalized NP intensity** - per-well mean log-intensity, normalized per plate to VO = 100% and Mock = 0%.

### Caveats & Best Practices
- **Outlier Masking**: Single-replicate non-monotonic outliers (e.g. 0.5 µM) should be masked from regression fits to prevent Hill slope distortion while keeping data points plotted for transparency.
- **Confluence Artifacts**: Over-confluent or washed-off wells (e.g. 0.75 µM) must be excluded from analysis (`DOSES_TO_EXCLUDE`, `WELLS_TO_EXCLUDE`).
