---
title: "Image-Based Infection + Dose-Response Analysis"
author: "Alejandro Ponce-Flores"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

<style>
.main-container table, table.table {
  border: 1.5px solid #2c3e50 !important;
  border-collapse: collapse !important;
  margin-top: 12px;
  margin-bottom: 24px;
  width: 100%;
}
.main-container th, .main-container td, table.table th, table.table td {
  border: 1px solid #bdc3c7 !important;
  padding: 8px 12px !important;
}
.main-container th, table.table th {
  background-color: #ecf0f1 !important;
  border-bottom: 2px solid #2c3e50 !important;
  color: #2c3e50;
  font-weight: bold;
}
.main-container tr:nth-child(even), table.table tr:nth-child(even) {
  background-color: #f8f9fa;
}
.main-container tr:hover, table.table tr:hover {
  background-color: #eaf2f8;
}
</style>

```{r user-config, message=FALSE, warning=FALSE}
## ── USER CONFIGURATION ──────────────────────────────────────────────────────
#
# Input files (CQ1 per-cell CSV exports, one per channel):
NUCLEUS_FILE <- "data/nucleus.csv"   # DAPI / nuclear channel
VIRUS_FILE   <- "data/virus.csv"     # Virus / reporter channel

# Column names in NUCLEUS_FILE for nuclear area & DAPI mean intensity
NUCLEUS_DAPI_COL    <- "(nucleus) MeanIntensity CH1"
NUCLEUS_AREA_COL    <- "(nucleus) Area"
VIRUS_INTENSITY_COL <- "(Virus) MeanIntensity CH2"

# Study labels
VIRUS_NAME    <- "Virus"
COMPOUND_NAME <- "Compound"
DOSE_UNIT     <- "\u00b5M"

# Cell gating thresholds
AREA_MIN        <- 30     # floor to retain pyknotic nuclei
AREA_MAX        <- 400    # ceiling to exclude clumps
INTENSITY_MIN   <- 300    # DAPI intensity floor
CIRCULARITY_MIN <- 0.40   # doublet / fragment cutoff (if sphericity available)

# Data Exclusions
DOSES_TO_EXCLUDE <- integer(0)   # e.g., c(0.75)
WELLS_TO_EXCLUDE <- character(0) # e.g., c("C3", "E4")

# ── Plate map ─────────────────────────────────────────────────────────────────
# Edit this tribble to match your actual well layout.
# Treatment options:
#   "Mock"             — uninfected, no drug (absolute negative control)
#   "Mock+Ab"          — uninfected + neutralizing Ab (gate reference)
#   "Virus Only"       — infected, no drug (positive control for infection)
#   "Toxicity Control" — drug only, no virus (for CC50)
#   "Virus + Compound" — infected + drug (experimental)
#
PLATE_MAP <- tibble::tribble(
  ~WellName,  ~Treatment,          ~Concentration_uM,
  "C3",  "Mock",             NA,
  "C4",  "Mock",             NA,
  "C5",  "Mock",             NA,
  "C6",  "Mock",             NA,
  "C11", "Mock+Ab",          NA,
  "C7",  "Virus Only",        0,
  "C8",  "Virus Only",        0,
  "D1",  "Toxicity Control",  0.01,
  "D2",  "Toxicity Control",  0.1,
  "D3",  "Toxicity Control",  0.5,
  "D4",  "Toxicity Control",  1,
  "D5",  "Toxicity Control",  2,
  "D6",  "Toxicity Control",  4,
  "D7",  "Toxicity Control",  8,
  "D8",  "Toxicity Control",  12,
  "D9",  "Toxicity Control",  16,
  "D10", "Toxicity Control",  20,
  "D11", "Toxicity Control",  30,
  "D12", "Toxicity Control",  50,
  "E1",  "Virus + Compound",  0.01,
  "E2",  "Virus + Compound",  0.1,
  "E3",  "Virus + Compound",  0.5,
  "E4",  "Virus + Compound",  1,
  "E5",  "Virus + Compound",  2,
  "E6",  "Virus + Compound",  4,
  "E7",  "Virus + Compound",  8,
  "E8",  "Virus + Compound",  12,
  "E9",  "Virus + Compound",  16,
  "E10", "Virus + Compound",  20,
  "E11", "Virus + Compound",  30,
  "E12", "Virus + Compound",  50,
  "F1",  "Virus + Compound",  0.01,
  "F2",  "Virus + Compound",  0.1,
  "F3",  "Virus + Compound",  0.5,
  "F4",  "Virus + Compound",  1,
  "F5",  "Virus + Compound",  2,
  "F6",  "Virus + Compound",  4,
  "F7",  "Virus + Compound",  8,
  "F8",  "Virus + Compound",  12,
  "F9",  "Virus + Compound",  16,
  "F10", "Virus + Compound",  20,
  "F11", "Virus + Compound",  30,
  "F12", "Virus + Compound",  50
)
## ────────────────────────────────────────────────────────────────────────────
```

# Setup

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(scales)
library(ggridges)
library(ggpubr)
library(RColorBrewer)
library(gtools)
library(grid)
library(ragg)
library(patchwork)
library(mclust)
library(minpack.lm)
library(MASS)
```

```{r global-theme}
base_family <- "sans"

theme_set(
  theme_bw(base_size = 8, base_family = base_family) +
    theme(
      plot.title   = element_text(face = "bold", size = 10, margin = margin(b = 4)),
      axis.title   = element_text(face = "bold", size = 9),
      axis.text    = element_text(size = 8),
      legend.title = element_text(size = 8),
      legend.text  = element_text(size = 8),
      axis.line    = element_line(linewidth = 0.6),
      axis.ticks   = element_line(linewidth = 0.6),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
)

update_geom_defaults("text",     list(size = 8 / ggplot2::.pt))
update_geom_defaults("label",    list(size = 8 / ggplot2::.pt))
update_geom_defaults("point",    list(size = 1.8))
update_geom_defaults("errorbar", list(linewidth = 0.5))
update_geom_defaults("line",     list(linewidth = 0.9))

dir.create("Plots", showWarnings = FALSE)
```

# Load & Annotate Data

```{r load-data, message=FALSE}
process_plate_files <- function(nucleus_file, virus_file) {
  df_nuclei <- read_csv(nucleus_file, show_col_types = FALSE)
  df_virus  <- read_csv(virus_file,   show_col_types = FALSE)

  dapi_col <- names(df_nuclei)[grepl("MeanIntensity.*CH1", names(df_nuclei), ignore.case = TRUE)][1]
  if (is.na(dapi_col)) dapi_col <- NUCLEUS_DAPI_COL

  area_col <- names(df_nuclei)[grepl("(Area|Volume)", names(df_nuclei), ignore.case = TRUE)][1]
  if (is.na(area_col)) area_col <- NUCLEUS_AREA_COL

  sphericity_col <- names(df_nuclei)[grepl("Sphericity|Circularity|Compactness", names(df_nuclei), ignore.case = TRUE)][1]
  
  virus_col <- names(df_virus)[grepl("MeanIntensity.*CH2", names(df_virus), ignore.case = TRUE)][1]
  if (is.na(virus_col)) virus_col <- VIRUS_INTENSITY_COL

  if (!is.na(sphericity_col)) {
    nuclei_clean <- df_nuclei %>%
      dplyr::select(
        WellName, FieldIndex, ObjectNumber,
        dapi_intensity = all_of(dapi_col),
        Area           = all_of(area_col),
        Sphericity     = all_of(sphericity_col)
      )
  } else {
    nuclei_clean <- df_nuclei %>%
      dplyr::select(
        WellName, FieldIndex, ObjectNumber,
        dapi_intensity = all_of(dapi_col),
        Area           = all_of(area_col)
      ) %>%
      dplyr::mutate(Sphericity = 1.0)
  }

  virus_clean <- df_virus %>%
    dplyr::select(
      WellName, FieldIndex, ObjectNumber,
      virus_intensity = all_of(virus_col)
    )

  nuclei_clean %>%
    dplyr::left_join(virus_clean, by = c("WellName", "FieldIndex", "ObjectNumber")) %>%
    dplyr::mutate(
      WellName_std = str_replace(str_to_upper(WellName), "^([A-H])-?0*([1-9][0-9]?).*$", "\\1\\2")
    )
}

master_df <- process_plate_files(NUCLEUS_FILE, VIRUS_FILE)

master_df_raw_annotated <- master_df %>%
  dplyr::left_join(PLATE_MAP, by = c("WellName_std" = "WellName")) %>%
  dplyr::filter(!is.na(Treatment))

master_df_annotated <- master_df_raw_annotated %>%
  dplyr::filter(!WellName_std %in% WELLS_TO_EXCLUDE) %>%
  { if (length(DOSES_TO_EXCLUDE) > 0)
      dplyr::filter(., !Concentration_uM %in% DOSES_TO_EXCLUDE)
    else . }
```

# Cell Gating (QC)

```{r gating-plot, warning=FALSE, message=FALSE}
master_df_pre <- master_df_annotated %>%
  dplyr::filter(Area > AREA_MIN & Area < AREA_MAX & dapi_intensity > INTENSITY_MIN) %>%
  dplyr::mutate(
    integrated_intensity = Area * dapi_intensity,
    Is_Singlet = Sphericity >= CIRCULARITY_MIN
  )

master_df_gated <- master_df_pre %>% dplyr::filter(Is_Singlet)
```

# Infection Gate: Control-Based Normalization + Population GMM

```{r infection-gate, warning=FALSE}
VIRUS_LABEL_VO  <- "Virus Only"
master_df_gated <- master_df_gated %>% mutate(Plate = "P1")

gmm_threshold <- function(mu, sdv, pi_) {
  o <- order(mu); mu <- mu[o]; sdv <- sdv[o]; pi_ <- pi_[o]
  sig1_sq <- sdv[1]^2; sig2_sq <- sdv[2]^2
  A <- 1/(2*sig1_sq) - 1/(2*sig2_sq)
  B <- -mu[1]/sig1_sq + mu[2]/sig2_sq
  C <- mu[1]^2/(2*sig1_sq) - mu[2]^2/(2*sig2_sq) + log(sdv[1]/sdv[2]) - log(pi_[1]/pi_[2])
  disc <- B^2 - 4*A*C
  if (disc >= 0 && abs(A) > 1e-10) {
    roots <- c((-B + sqrt(disc))/(2*A), (-B - sqrt(disc))/(2*A))
    r_valid <- roots[roots > mu[1] & roots < mu[2]]
    if (length(r_valid) > 0) return(r_valid[1])
  }
  (mu[1] + mu[2]) / 2
}

fit_high_mode <- function(xlog) {
  fit <- tryCatch(Mclust(xlog, G = 2, modelNames = "V", verbose = FALSE), error = function(e) NULL)
  if (!is.null(fit)) return(max(as.numeric(fit$parameters$mean)))
  as.numeric(quantile(xlog, 0.75, na.rm = TRUE))
}

# 1. Compute Anchors (L_p, H_p)
mock_cells <- master_df_gated %>% filter(Treatment %in% c("Mock", "Mock+Ab")) %>% mutate(xlog = log10(virus_intensity + 1))
L_p_val    <- if (nrow(mock_cells) > 0) median(mock_cells$xlog, na.rm = TRUE) else 2.0

vo_cells   <- master_df_gated %>% filter(Treatment == VIRUS_LABEL_VO) %>% mutate(xlog = log10(virus_intensity + 1))
H_p_val    <- if (nrow(vo_cells) > 0) fit_high_mode(vo_cells$xlog) else 4.0

# 2. Normalize Cells
master_df_norm <- master_df_gated %>%
  mutate(
    xlog   = log10(virus_intensity + 1),
    L_p    = L_p_val,
    H_p    = H_p_val,
    span   = H_p - L_p,
    x_norm = if_else(is.na(span) | span < 1e-6, NA_real_, (xlog - L_p) / span)
  )

# 3. Fit Population 2-GMM Gate
infected_culture_norm <- master_df_norm %>% filter(Treatment %in% c(VIRUS_LABEL_VO, "Virus + Compound"))
med_val <- suppressWarnings(median(table(infected_culture_norm$WellName_std), na.rm = TRUE))
target_n_gate <- if (is.finite(med_val) && med_val > 0) as.integer(round(med_val)) else 200L

set.seed(42)
gate_balanced <- infected_culture_norm %>%
  group_by(WellName_std) %>%
  filter(row_number() <= target_n_gate) %>%
  ungroup()

x_gate <- gate_balanced$x_norm[is.finite(gate_balanced$x_norm)]
fit_gate <- tryCatch(
  Mclust(x_gate, G = 2, modelNames = "V", verbose = FALSE),
  error = function(e) NULL
)

if (!is.null(fit_gate) && !is.null(fit_gate$parameters$mean) && length(fit_gate$parameters$mean) == 2) {
  mu_gate  <- as.numeric(fit_gate$parameters$mean)
  sdv_gate <- sqrt(fit_gate$parameters$variance$sigmasq)
  pi_gate  <- as.numeric(fit_gate$parameters$pro)
  if (length(sdv_gate) == 1L) sdv_gate <- rep(sdv_gate, 2L)
  threshold_final <- gmm_threshold(mu_gate, sdv_gate, pi_gate)
} else {
  threshold_final <- 0.5
}
master_df_final <- master_df_norm
```

# Percent Infection per Well

```{r percent-infection}
total_counts <- master_df_final %>% dplyr::count(Treatment, Concentration_uM, WellName_std, name = "total_cell_count")
positive_counts <- master_df_final %>% filter(x_norm > threshold_final) %>% dplyr::count(Treatment, Concentration_uM, WellName_std, name = "positive_cell_count")

percent_summary <- total_counts %>%
  left_join(positive_counts, by = c("Treatment", "Concentration_uM", "WellName_std")) %>%
  mutate(
    positive_cell_count = replace_na(positive_cell_count, 0),
    percent_positive     = (positive_cell_count / total_cell_count) * 100,
    Plate = "Plate 1",
    Precision_Flag = if_else(total_cell_count < 200, "Low Precision", "Normal")
  )

percent_summary %>%
  mutate(
    Dose_Label = case_when(
      Treatment %in% c("Mock", "Mock+Ab") ~ "Mock",
      Treatment == "Virus Only"           ~ "0 µM (Control)",
      TRUE                                ~ paste0(Concentration_uM, " µM")
    ),
    Status_Label = case_when(
      Precision_Flag == "Normal" ~ "Valid",
      TRUE                       ~ Precision_Flag
    )
  ) %>%
  dplyr::select(
    Well              = WellName_std,
    Treatment,
    Dose              = Dose_Label,
    `Total Cells`     = total_cell_count,
    `Infected Cells`  = positive_cell_count,
    `% Infection`    = percent_positive,
    `QC Status`       = Status_Label
  ) %>%
  knitr::kable(
    caption     = "Well-Level Percent Infection (Global Population GMM Threshold)",
    digits      = 1,
    format.args = list(big.mark = ",")
  )
```

# Co-Primary Endpoint 1: % Infection Dose-Response (4PL)

```{r dose-response-model}
vo_anchor   <- mean(percent_summary$percent_positive[percent_summary$Treatment == "Virus Only"], na.rm = TRUE)
mock_anchor <- mean(percent_summary$percent_positive[percent_summary$Treatment %in% c("Mock", "Mock+Ab")], na.rm = TRUE)
if (is.nan(mock_anchor) || is.na(mock_anchor)) mock_anchor <- 0

summary_rel <- percent_summary %>%
  mutate(
    percent_infection_rel = 100 * (percent_positive - mock_anchor) / (vo_anchor - mock_anchor)
  )

drc_data_rel <- summary_rel %>% filter(Treatment == "Virus + Compound")
pos_conc  <- drc_data_rel$Concentration_uM[drc_data_rel$Concentration_uM > 0 & is.finite(drc_data_rel$Concentration_uM)]
min_pos   <- if (length(pos_conc) > 0) min(pos_conc) else 0.01
zero_tick <- min_pos / 10

fit_data_wide <- summary_rel %>%
  filter(Treatment == "Virus Only" | Treatment == "Virus + Compound") %>%
  mutate(
    dose_fit   = ifelse(Treatment == "Virus Only", zero_tick, Concentration_uM),
    prop_raw   = positive_cell_count / total_cell_count,
    prop_clamp = pmin(pmax(prop_raw, 1e-3), 1 - 1e-3),
    bin_weight = total_cell_count / (prop_clamp * (1 - prop_clamp))
  )

bootstrap_ec50_ci <- function(fit, R = 2000, seed = 42) {
  if (is.null(fit)) return(c(NA_real_, NA_real_, NA_real_))
  mu <- coef(fit)
  vc <- tryCatch(vcov(fit), error = function(e) NULL)
  if (is.null(vc)) return(c(NA_real_, NA_real_, NA_real_))
  set.seed(seed)
  sim <- tryCatch(MASS::mvrnorm(R, mu, vc), error = function(e) NULL)
  if (is.null(sim)) return(c(NA_real_, NA_real_, NA_real_))
  ec <- sim[, "ec50"]
  ec <- ec[is.finite(ec) & ec > 0]
  if (length(ec) < 10) return(c(NA_real_, NA_real_, NA_real_))
  c(lower = unname(quantile(ec, 0.025)), median = unname(quantile(ec, 0.50)), upper = unname(quantile(ec, 0.975)))
}

fit_rel <- NULL
for (ec50_try in c(0.5, 1, 2, 5, 8, 10, 0.1)) {
  fit_rel <- tryCatch(
    minpack.lm::nlsLM(
      percent_infection_rel ~ bottom + (top - bottom) / (1 + (dose_fit / ec50) ^ hill),
      data    = fit_data_wide,
      start   = list(ec50 = ec50_try, hill = 1.0, bottom = 0, top = 100),
      lower   = c(ec50 = 1e-4, hill = 0.05, bottom = -20, top = 80),
      upper   = c(ec50 = 1e3,  hill = 50, bottom = 50, top = 120),
      weights = fit_data_wide$bin_weight,
      control = minpack.lm::nls.lm.control(maxiter = 500)
    ),
    error   = function(e) NULL
  )
  if (!is.null(fit_rel)) break
}

ci_rel <- bootstrap_ec50_ci(fit_rel)
```

## Dose-Response Plot

```{r dose-response-plot, message=FALSE, warning=FALSE}
max_pos <- if (length(pos_conc) > 0) max(pos_conc) else 100
newx    <- exp(seq(log(zero_tick), log(max_pos), length.out = 300))

pred_df_rel <- tibble(Concentration_uM = numeric(0), y = numeric(0))
if (!is.null(fit_rel)) {
  coef_rel   <- coef(fit_rel)
  ec50_rel   <- coef_rel["ec50"]
  hill_rel   <- coef_rel["hill"]
  bottom_rel <- coef_rel["bottom"]
  top_rel    <- coef_rel["top"]
  pred_df_rel <- tibble(
    Concentration_uM = newx,
    y = bottom_rel + (top_rel - bottom_rel) / (1 + (Concentration_uM / ec50_rel)^hill_rel)
  )
  ec50_rel_lab <- paste0("EC50 = ", signif(ec50_rel, 3), " ", DOSE_UNIT)
} else {
  ec50_rel_lab <- "Fit Failed"
}

pts_sum <- summary_rel %>%
  filter(Treatment %in% c("Virus Only", "Virus + Compound")) %>%
  mutate(x_plot = ifelse(Concentration_uM == 0, zero_tick, Concentration_uM)) %>%
  group_by(x_plot) %>%
  summarise(n  = n(),
            y  = mean(percent_infection_rel, na.rm = TRUE),
            se = ifelse(n > 1, sd(percent_infection_rel, na.rm = TRUE) / sqrt(n), 0),
            .groups = "drop") %>%
  mutate(y    = ifelse(x_plot == zero_tick, 100, y),
         ymin = pmax(0,   y - se),
         ymax = pmin(100, y + se))

if (length(pos_conc) > 6) {
  raw_brks <- 10^seq(floor(log10(min_pos)), ceiling(log10(max_pos)))
  dose_breaks <- sort(unique(c(zero_tick, raw_brks[raw_brks >= min_pos & raw_brks <= max_pos], max_pos)))
} else {
  dose_breaks <- sort(unique(c(zero_tick, drc_data_rel$Concentration_uM)))
}
dose_labels_vec <- c("0", as.character(dose_breaks[-1]))

p_dr_rel <- ggplot() +
  geom_errorbar(data = pts_sum, aes(x = x_plot, ymin = ymin, ymax = ymax),
                width = 0.06, linewidth = 0.5, colour = "steelblue4") +
  geom_point(data = pts_sum, aes(x = x_plot, y = y),
             shape = 21, size = 2.6, stroke = 0.4,
             fill = "steelblue4", colour = "steelblue4")

if (!is.null(fit_rel)) {
  y_rel_halfway <- bottom_rel + (top_rel - bottom_rel) / 2
  p_dr_rel <- p_dr_rel + 
    geom_line(data = pred_df_rel, aes(x = Concentration_uM, y = y),
              color = "steelblue", linewidth = 0.9) +
    geom_segment(aes(x = ec50_rel, xend = ec50_rel, y = 0, yend = y_rel_halfway),
                 linetype = "dashed", color = "steelblue4", linewidth = 0.5) +
    geom_segment(aes(x = zero_tick, xend = ec50_rel, y = y_rel_halfway, yend = y_rel_halfway),
                 linetype = "dashed", color = "steelblue4", linewidth = 0.5) +
    annotate("label", x = ec50_rel, y = 25, label = ec50_rel_lab,
             fill = "steelblue4", color = "white", size = 3, label.size = 0)
}

p_dr_rel <- p_dr_rel +
  scale_x_log10(limits = c(zero_tick * 0.8, max_pos * 1.2), breaks = dose_breaks, labels = dose_labels_vec, minor_breaks = NULL) +
  annotation_logticks(sides = "b", short = unit(1.2, "mm"), mid = unit(1.6, "mm"), long = unit(2.2, "mm")) +
  labs(
    title = paste(COMPOUND_NAME, "relative dose\u2013response"),
    x     = paste0(COMPOUND_NAME, " (", DOSE_UNIT, ", log10)"),
    y     = "% infected cells"
  ) +
  coord_cartesian(ylim = c(0, 105), clip = "off") +
  theme_bw(base_size = 8) +
  theme(
    plot.title        = element_text(face = "bold", size = 9, margin = margin(b = 4)),
    axis.title        = element_text(face = "bold", size = 8),
    axis.text         = element_text(size = 7),
    axis.line         = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks        = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks.length = unit(2.2, "mm"),
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    legend.position   = "none",
    plot.margin       = margin(6, 10, 6, 8)
  )

print(p_dr_rel)
ggsave("Plots/compound_dose_response_relative.tiff", p_dr_rel,
       width = 4, height = 3, units = "in", dpi = 600, compression = "lzw", device = ragg::agg_tiff)
```

# Co-Primary Endpoint 2: Normalized NP Intensity Dose-Response

```{r intensity-dose-response, warning=FALSE}
intensity_well <- master_df_final %>%
  filter(Treatment %in% c("Mock", "Mock+Ab", "Virus Only", "Virus + Compound")) %>%
  group_by(WellName_std, Treatment, Concentration_uM) %>%
  summarise(
    n_cells     = n(),
    mean_xlog   = mean(xlog, na.rm = TRUE),
    L_p = first(L_p), H_p = first(H_p),
    .groups = "drop"
  ) %>%
  mutate(intensity_norm_mean = 100 * (mean_xlog - L_p) / (H_p - L_p))

vo_i   <- mean(intensity_well$intensity_norm_mean[intensity_well$Treatment == "Virus Only"], na.rm = TRUE)
mock_i <- mean(intensity_well$intensity_norm_mean[intensity_well$Treatment %in% c("Mock", "Mock+Ab")], na.rm = TRUE)

intensity_rel <- intensity_well %>%
  mutate(intensity_rel_mean = 100 * (intensity_norm_mean - mock_i) / (vo_i - mock_i))

pos_conc_i  <- intensity_rel$Concentration_uM[intensity_rel$Concentration_uM > 0 & is.finite(intensity_rel$Concentration_uM)]
min_pos_i   <- if (length(pos_conc_i) > 0) min(pos_conc_i) else 0.01
zero_tick_i <- min_pos_i / 10

fit_i <- intensity_rel %>%
  filter(Treatment == "Virus Only" | Treatment == "Virus + Compound") %>%
  mutate(dose_fit = ifelse(Treatment == "Virus Only", zero_tick_i, Concentration_uM))

fit_int <- NULL
for (ec50_try in c(0.5, 1, 2, 5, 8, 10, 0.1)) {
  fit_int <- tryCatch(
    minpack.lm::nlsLM(
      intensity_rel_mean ~ bottom + (top - bottom) / (1 + (dose_fit / ec50) ^ hill),
      data    = fit_i,
      start   = list(ec50 = ec50_try, hill = 1.0, bottom = 0, top = 100),
      lower   = c(ec50 = 1e-4, hill = 0.05, bottom = -20, top = 80),
      upper   = c(ec50 = 1e3,  hill = 50, bottom = 50, top = 120),
      weights = fit_i$n_cells,
      control = minpack.lm::nls.lm.control(maxiter = 500)
    ),
    error = function(e) NULL
  )
  if (!is.null(fit_int)) break
}

ci_int <- bootstrap_ec50_ci(fit_int)

ec50_compare <- tibble(
  Endpoint   = c("Normalized NP intensity (mean) -- co-primary",
                 "% infected (population GMM gate) -- co-primary"),
  `EC50 (µM)` = c(
    if (!is.null(fit_int)) unname(coef(fit_int)["ec50"]) else NA_real_,
    if (!is.null(fit_rel))  unname(coef(fit_rel)["ec50"])  else NA_real_
  ),
  `95% CI`    = c(
    if (!is.null(fit_int)) paste0(signif(ci_int["lower"],3)," – ",signif(ci_int["upper"],3)) else "–",
    if (!is.null(fit_rel))  paste0(signif(ci_rel["lower"],3)," – ",signif(ci_rel["upper"],3))  else "–"
  )
)
knitr::kable(ec50_compare, caption = "EC50 comparison across co-primary endpoints.", digits = 3)
```

## Intensity Dose-Response Plot

```{r intensity-dose-response-plot, warning=FALSE, message=FALSE}
pts_i <- intensity_rel %>%
  filter(Treatment %in% c("Virus Only", "Virus + Compound")) %>%
  mutate(
    x_plot = ifelse(Treatment == "Virus Only", zero_tick_i, Concentration_uM),
    x_plot = ifelse(is.na(x_plot), NA_real_, x_plot)
  ) %>%
  filter(!is.na(x_plot)) %>%
  group_by(x_plot) %>%
  summarise(n  = n(),
            y  = mean(intensity_rel_mean, na.rm = TRUE),
            se = ifelse(n > 1, sd(intensity_rel_mean, na.rm = TRUE) / sqrt(n), 0),
            .groups = "drop") %>%
  mutate(y    = ifelse(x_plot == zero_tick_i, 100, y),
         ymin = pmax(0,   y - se),
         ymax = pmin(105, y + se))

max_pos_i <- if (length(pos_conc_i) > 0) max(pos_conc_i) else 100
newx_i    <- exp(seq(log(zero_tick_i), log(max_pos_i), length.out = 300))

pred_i <- tibble(Concentration_uM = numeric(0), y = numeric(0))
if (!is.null(fit_int)) {
  cf <- coef(fit_int)
  pred_i <- tibble(
    Concentration_uM = newx_i,
    y = cf["bottom"] + (cf["top"] - cf["bottom"]) / (1 + (Concentration_uM / cf["ec50"]) ^ cf["hill"])
  )
}

if (length(pos_conc_i) > 6) {
  raw_brks_i <- 10^seq(floor(log10(min_pos_i)), ceiling(log10(max_pos_i)))
  dose_breaks_i <- sort(unique(c(zero_tick_i, raw_brks_i[raw_brks_i >= min_pos_i & raw_brks_i <= max_pos_i], max_pos_i)))
} else {
  dose_breaks_i <- sort(unique(c(zero_tick_i, intensity_rel$Concentration_uM[intensity_rel$Treatment == "Virus + Compound"])))
}
dose_labels_i <- c("0", as.character(dose_breaks_i[-1]))

p_dr_int <- ggplot() +
  geom_errorbar(data = pts_i, aes(x = x_plot, ymin = ymin, ymax = ymax),
                width = 0.06, linewidth = 0.5, colour = "steelblue4") +
  geom_point(data = pts_i, aes(x = x_plot, y = y),
             shape = 21, size = 2.6, stroke = 0.4,
             fill = "steelblue4", colour = "steelblue4")

if (!is.null(fit_int)) {
  cf <- coef(fit_int)
  ec50_val  <- unname(cf["ec50"])
  halfway   <- unname(cf["bottom"] + (cf["top"] - cf["bottom"]) / 2)
  int_label <- paste0("EC50 = ", signif(ec50_val, 3), " ", DOSE_UNIT)
  p_dr_int  <- p_dr_int +
    geom_line(data = pred_i, aes(x = Concentration_uM, y = y),
              color = "steelblue", linewidth = 0.9) +
    geom_segment(aes(x = ec50_val, xend = ec50_val, y = 0, yend = halfway),
                 linetype = "dashed", color = "steelblue4", linewidth = 0.5) +
    geom_segment(aes(x = zero_tick_i, xend = ec50_val, y = halfway, yend = halfway),
                 linetype = "dashed", color = "steelblue4", linewidth = 0.5) +
    annotate("label", x = ec50_val, y = 25, label = int_label,
             fill = "steelblue4", color = "white", size = 3, label.size = 0)
}

p_dr_int <- p_dr_int +
  scale_x_log10(limits = c(zero_tick_i * 0.8, max_pos_i * 1.2), breaks = dose_breaks_i, labels = dose_labels_i, minor_breaks = NULL) +
  annotation_logticks(sides = "b", short = unit(1.2, "mm"), mid = unit(1.6, "mm"), long = unit(2.2, "mm")) +
  labs(
    title = paste(COMPOUND_NAME, "intensity dose\u2013response"),
    x     = paste0(COMPOUND_NAME, " (", DOSE_UNIT, ", log10)"),
    y     = "Normalized NP intensity (%)"
  ) +
  coord_cartesian(ylim = c(0, 105), clip = "off") +
  theme_bw(base_size = 8) +
  theme(
    plot.title        = element_text(face = "bold", size = 9, margin = margin(b = 4)),
    axis.title        = element_text(face = "bold", size = 8),
    axis.text         = element_text(size = 7),
    axis.line         = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks        = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks.length = unit(2.2, "mm"),
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    legend.position   = "none",
    plot.margin       = margin(6, 10, 6, 8)
  )

print(p_dr_int)
ggsave("Plots/compound_dose_response_intensity.tiff", p_dr_int,
       width = 4, height = 3, units = "in", dpi = 600, compression = "lzw", device = ragg::agg_tiff)
```

# Results Summary

Two co-primary endpoints are reported side-by-side:
1. **% infected** - fraction of cells above the population 2-GMM gate, fitted with binomial-weighted relative 4PL.
2. **Normalized NP intensity** - per-well mean log-intensity, normalized per plate to VO = 100% and Mock = 0%.

### Caveats & Best Practices
- **Outlier Masking**: Single-replicate non-monotonic outliers should be masked from regression fits to prevent Hill slope distortion while keeping data points plotted for transparency.
- **Confluence Artifacts**: Over-confluent or washed-off wells must be excluded from analysis (`DOSES_TO_EXCLUDE`, `WELLS_TO_EXCLUDE`).
