---
title: "Image-Based Infection + Dose-Response Analysis"
author: "Alejandro Ponce-Flores"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

<style>
.main-container table, table.table {
  border: 1.5px solid #2c3e50 !important;
  border-collapse: collapse !important;
  margin-top: 12px;
  margin-bottom: 24px;
  width: 100%;
}
.main-container th, .main-container td, table.table th, table.table td {
  border: 1px solid #bdc3c7 !important;
  padding: 8px 12px !important;
}
.main-container th, table.table th {
  background-color: #ecf0f1 !important;
  border-bottom: 2px solid #2c3e50 !important;
  color: #2c3e50;
  font-weight: bold;
}
.main-container tr:nth-child(even), table.table tr:nth-child(even) {
  background-color: #f8f9fa;
}
.main-container tr:hover, table.table tr:hover {
  background-color: #eaf2f8;
}
</style>

```{r user-config, message=FALSE, warning=FALSE}
## ── USER CONFIGURATION ──────────────────────────────────────────────────────
#
# Input files (CQ1 per-cell CSV exports, one per channel):
NUCLEUS_FILE <- "data/nucleus.csv"   # DAPI / nuclear channel
VIRUS_FILE   <- "data/virus.csv"     # Virus / reporter channel

# Column names in NUCLEUS_FILE / VIRUS_FILE.
#   These are AUTHORITATIVE: if a name below is present in the file it is used.
#   Autodetection by pattern is a fallback for when it is not, and it stops with
#   an error if the pattern is ambiguous (HCS exports routinely carry several
#   "Area" columns) rather than silently taking the first match.
NUCLEUS_DAPI_COL       <- "(nucleus) MeanIntensity CH1"
NUCLEUS_AREA_COL       <- "(nucleus) Area"
NUCLEUS_SPHERICITY_COL <- "(nucleus) Sphericity"
VIRUS_INTENSITY_COL    <- "(Virus) MeanIntensity CH2"

# REQUIRE_SPHERICITY: Sphericity drives the doublet-discrimination gate. When
#   TRUE (recommended) a missing Sphericity column is a hard error. When FALSE
#   the doublet step is SKIPPED and clearly reported as skipped \u2014 it is never
#   faked by assigning every object a perfect sphericity of 1.0, which would
#   make the gate a silent no-op while still printing the QC panel.
REQUIRE_SPHERICITY <- TRUE

# Study labels
VIRUS_NAME    <- "Virus"
COMPOUND_NAME <- "Compound"
DOSE_UNIT     <- "\u00b5M"

# Cell gating thresholds
AREA_MIN        <- 30     # floor to retain pyknotic nuclei
AREA_MAX        <- 700    # ceiling to exclude clumps (3+ merged nuclei)
INTENSITY_MIN   <- 300    # DAPI intensity floor
# CIRCULARITY_MIN: Sphericity cutoff separating single nuclei from doublets and
#   segmentation debris. Real nuclear sphericity runs ~0.7-1.0 for singlets, so
#   a cutoff far below that range retains essentially everything and turns this
#   step into a no-op. Inspect the "Doublet & Shape Discrimination" panel below
#   and put the line in the valley between the two modes.
CIRCULARITY_MIN <- 0.75

# MAHALANOBIS_LEVEL: Confidence level for the cell-cycle / DNA-content gate.
MAHALANOBIS_LEVEL <- 0.95

# GATE_REFERENCE_TREATMENTS: Which wells define the "normal" DNA-content
#   distribution the Mahalanobis gate is fitted on. Fitting on ALL wells lets
#   cytotoxic high-dose wells shift the ellipse that decides which of their own
#   cells to keep \u2014 a feedback loop. Untreated / mock wells are the right
#   reference. Set to NULL to fall back to pooling every singlet.
GATE_REFERENCE_TREATMENTS <- c("Mock", "Mock+Ab")

# MIN_GATE_REFERENCE_CELLS: If the reference population has fewer cells than
#   this, the gate falls back to all singlets (with a visible warning).
MIN_GATE_REFERENCE_CELLS <- 200

# WELL_NAME_REGEX / WELL_NAME_REPLACEMENT: Normalizes plate well names to the
#   "<row><col>" form used by PLATE_MAP (e.g. "b-03" -> "B3"). The row class
#   covers A-P and the column 1-24, so 96-, 384- and 1536-well plates all work.
#   Any well that fails to normalize is reported, not silently discarded.
WELL_NAME_REGEX       <- "^([A-P])-?0*([1-9][0-9]?).*$"
WELL_NAME_REPLACEMENT <- "\\1\\2"

# Well-level QC
# MIN_CELLS_PER_WELL: Wells with fewer recovered nuclei than this are EXCLUDED
#   from the dose-response fits (their percent-infection estimate is too noisy
#   to weight sensibly). They remain in the per-well table, marked as excluded.
MIN_CELLS_PER_WELL   <- 50
# PRECISION_WARN_CELLS: Wells below this are flagged "Low Precision" in the
#   table but still contribute to the fits.
PRECISION_WARN_CELLS <- 200

# Dose-response fitting
# ZERO_DOSE_TICK_DIVISOR: The zero-dose control has no place on a log axis; it
#   is drawn at min(positive dose) / this divisor.
ZERO_DOSE_TICK_DIVISOR <- 10
# INCLUDE_ZERO_DOSE_IN_FIT: Whether the zero-dose (Virus Only) control enters
#   the 4PL regression as if it were a real dose at that fictitious tick.
#   FALSE (default) is correct: the plotted tick position is an artifact of the
#   log axis, and feeding it to the model pulls the `top` asymptote and shifts
#   EC50. The control is still plotted, and still anchors the normalization.
INCLUDE_ZERO_DOSE_IN_FIT <- FALSE
# EC50_CI_DRAWS: Number of parametric draws used for the EC50/CC50 confidence
#   interval (see ec50_ci_parametric() below).
EC50_CI_DRAWS <- 2000
# LOG_TICK_DOSE_COUNT: Above this many DISTINCT doses, the x axis switches from
#   one break per dose to decade breaks.
LOG_TICK_DOSE_COUNT <- 6

# Reproducibility: single seed for every stochastic step in this template
# (the balanced GMM training sample and the parametric CI draws).
RANDOM_SEED <- 42

# Output
OUTPUT_DIR <- "Plots"
PLOT_DPI   <- 300

# Data Exclusions
DOSES_TO_EXCLUDE <- integer(0)   # e.g., c(0.75)
WELLS_TO_EXCLUDE <- character(0) # e.g., c("C3", "E4")

# ── Plate map ─────────────────────────────────────────────────────────────────
# Edit this tribble to match your actual well layout.
# Treatment options:
#   "Mock"             — uninfected, no drug (absolute negative control)
#   "Mock+Ab"          — uninfected + neutralizing Ab (gate reference)
#   "Virus Only"       — infected, no drug (positive control for infection)
#   "Toxicity Control" — drug only, no virus (for CC50)
#   "Virus + Compound" — infected + drug (experimental)
#
PLATE_MAP <- tibble::tribble(
  ~WellName,  ~Treatment,          ~Concentration_uM,
  "C3",  "Mock",             NA,
  "C4",  "Mock",             NA,
  "C5",  "Mock",             NA,
  "C6",  "Mock",             NA,
  "C11", "Mock+Ab",          NA,
  "C7",  "Virus Only",        0,
  "C8",  "Virus Only",        0,
  "D1",  "Toxicity Control",  0.01,
  "D2",  "Toxicity Control",  0.1,
  "D3",  "Toxicity Control",  0.5,
  "D4",  "Toxicity Control",  1,
  "D5",  "Toxicity Control",  2,
  "D6",  "Toxicity Control",  4,
  "D7",  "Toxicity Control",  8,
  "D8",  "Toxicity Control",  12,
  "D9",  "Toxicity Control",  16,
  "D10", "Toxicity Control",  20,
  "D11", "Toxicity Control",  30,
  "D12", "Toxicity Control",  50,
  "E1",  "Virus + Compound",  0.01,
  "E2",  "Virus + Compound",  0.1,
  "E3",  "Virus + Compound",  0.5,
  "E4",  "Virus + Compound",  1,
  "E5",  "Virus + Compound",  2,
  "E6",  "Virus + Compound",  4,
  "E7",  "Virus + Compound",  8,
  "E8",  "Virus + Compound",  12,
  "E9",  "Virus + Compound",  16,
  "E10", "Virus + Compound",  20,
  "E11", "Virus + Compound",  30,
  "E12", "Virus + Compound",  50,
  "F1",  "Virus + Compound",  0.01,
  "F2",  "Virus + Compound",  0.1,
  "F3",  "Virus + Compound",  0.5,
  "F4",  "Virus + Compound",  1,
  "F5",  "Virus + Compound",  2,
  "F6",  "Virus + Compound",  4,
  "F7",  "Virus + Compound",  8,
  "F8",  "Virus + Compound",  12,
  "F9",  "Virus + Compound",  16,
  "F10", "Virus + Compound",  20,
  "F11", "Virus + Compound",  30,
  "F12", "Virus + Compound",  50
)
## ────────────────────────────────────────────────────────────────────────────
```

# Setup

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(scales)
library(ggridges)
library(ggpubr)
library(RColorBrewer)
library(gtools)
library(grid)
library(ragg)
library(patchwork)
library(mclust)
library(minpack.lm)
# NOTE: MASS is deliberately NOT loaded. library(MASS) masks dplyr::select(),
# which is what forced the defensive `dplyr::` prefixes scattered through this
# file. Nothing here needs it: the EC50 confidence interval is drawn from a
# univariate normal on the log scale (see ec50_ci_parametric() below).
```

```{r global-theme}
base_family <- "sans"

theme_set(
  theme_bw(base_size = 8, base_family = base_family) +
    theme(
      plot.title   = element_text(face = "bold", size = 10, margin = margin(b = 4)),
      axis.title   = element_text(face = "bold", size = 9),
      axis.text    = element_text(size = 8),
      legend.title = element_text(size = 8),
      legend.text  = element_text(size = 8),
      axis.line    = element_line(linewidth = 0.6),
      axis.ticks   = element_line(linewidth = 0.6),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
)

update_geom_defaults("text",     list(size = 8 / ggplot2::.pt))
update_geom_defaults("label",    list(size = 8 / ggplot2::.pt))
update_geom_defaults("point",    list(size = 1.8))
update_geom_defaults("errorbar", list(linewidth = 0.5))
update_geom_defaults("line",     list(linewidth = 0.9))

dir.create(OUTPUT_DIR, showWarnings = FALSE, recursive = TRUE)
```

# Load & Annotate Data

```{r load-data, message=FALSE}
# ── Column resolution: the configured name WINS; the pattern is a fallback ────
# An ambiguous pattern is an error, not a coin flip. HCS exports routinely carry
# several columns matching "Area", and silently taking the first one produces a
# plausible-looking analysis of the wrong measurement.
resolve_column <- function(df, configured, pattern, what, file_label,
                           required = TRUE) {
  if (!is.null(configured) && configured %in% names(df)) return(configured)

  hits <- names(df)[grepl(pattern, names(df), ignore.case = TRUE)]
  if (length(hits) == 1L) {
    message("Column for ", what, " not found as ", sQuote(configured),
            " in ", file_label, "; using autodetected ", sQuote(hits), ".")
    return(hits)
  }
  if (length(hits) > 1L) {
    stop("Ambiguous autodetection for ", what, " in ", file_label, ": ",
         paste(sQuote(hits), collapse = ", "),
         ". Set the exact column name in the USER CONFIGURATION block.")
  }
  if (required) {
    stop("No column for ", what, " in ", file_label,
         ". Looked for ", sQuote(configured), " and for /", pattern, "/. ",
         "Available columns: ", paste(sQuote(names(df)), collapse = ", "))
  }
  NA_character_
}

process_plate_files <- function(nucleus_file, virus_file) {
  for (f in c(nucleus_file, virus_file)) {
    if (!file.exists(f)) {
      stop("Input file not found: ", sQuote(f),
           ". Run `Rscript data/simulate_data.R` from the template directory ",
           "to regenerate the demo data, or point the config at your own export.")
    }
  }

  df_nuclei <- read_csv(nucleus_file, show_col_types = FALSE)
  df_virus  <- read_csv(virus_file,   show_col_types = FALSE)

  key_cols <- c("WellName", "FieldIndex", "ObjectNumber")
  for (nm in c("nucleus", "virus")) {
    d <- if (nm == "nucleus") df_nuclei else df_virus
    miss <- setdiff(key_cols, names(d))
    if (length(miss) > 0) {
      stop("The ", nm, " export is missing the object key column(s): ",
           paste(sQuote(miss), collapse = ", "),
           ". These are needed to match nuclei to the reporter channel.")
    }
  }

  dapi_col <- resolve_column(df_nuclei, NUCLEUS_DAPI_COL,
                             "MeanIntensity.*CH1", "nuclear DAPI intensity",
                             "NUCLEUS_FILE")
  area_col <- resolve_column(df_nuclei, NUCLEUS_AREA_COL,
                             "(Area|Volume)", "nuclear area / volume",
                             "NUCLEUS_FILE")
  sphericity_col <- resolve_column(df_nuclei, NUCLEUS_SPHERICITY_COL,
                                   "Sphericity", "nuclear sphericity",
                                   "NUCLEUS_FILE", required = REQUIRE_SPHERICITY)
  virus_col <- resolve_column(df_virus, VIRUS_INTENSITY_COL,
                              "MeanIntensity.*CH2", "virus reporter intensity",
                              "VIRUS_FILE")

  nuclei_clean <- df_nuclei %>%
    dplyr::select(
      WellName, FieldIndex, ObjectNumber,
      dapi_intensity = all_of(dapi_col),
      Area           = all_of(area_col)
    )

  # A missing Sphericity column means the doublet gate CANNOT run. Record that
  # honestly (NA) instead of substituting 1.0, which would silently pass every
  # object through a gate the report claims was applied.
  has_sphericity <- !is.na(sphericity_col)
  nuclei_clean$Sphericity <- if (has_sphericity) {
    df_nuclei[[sphericity_col]]
  } else {
    NA_real_
  }
  attr(nuclei_clean, "has_sphericity") <- has_sphericity

  nuclei_clean <- nuclei_clean %>%
    dplyr::mutate(integrated_intensity = Area * dapi_intensity)

  virus_clean <- df_virus %>%
    dplyr::select(
      WellName, FieldIndex, ObjectNumber,
      virus_intensity = all_of(virus_col)
    )

  out <- nuclei_clean %>%
    dplyr::left_join(virus_clean, by = key_cols) %>%
    dplyr::mutate(
      WellName_std = str_replace(str_to_upper(WellName),
                                 WELL_NAME_REGEX, WELL_NAME_REPLACEMENT)
    )
  attr(out, "has_sphericity") <- has_sphericity
  out
}

master_df      <- process_plate_files(NUCLEUS_FILE, VIRUS_FILE)
HAS_SPHERICITY <- isTRUE(attr(master_df, "has_sphericity"))

# ── Well-name normalization + plate-map join diagnostics ─────────────────────
# Both of these used to fail silently. A well name the regex cannot parse comes
# through unchanged, misses the PLATE_MAP join, and is then deleted by
# filter(!is.na(Treatment)) — on a 384-well plate that was every single cell.
unparsed_wells <- master_df %>%
  dplyr::filter(!str_detect(str_to_upper(WellName), WELL_NAME_REGEX)) %>%
  dplyr::distinct(WellName) %>%
  dplyr::pull(WellName)

if (length(unparsed_wells) > 0) {
  stop(length(unparsed_wells), " well name(s) could not be normalized by ",
       "WELL_NAME_REGEX (", WELL_NAME_REGEX, "): ",
       paste(sQuote(head(unparsed_wells, 10)), collapse = ", "),
       if (length(unparsed_wells) > 10) ", ..." else "",
       ". Adjust WELL_NAME_REGEX / WELL_NAME_REPLACEMENT in the config.")
}

unmapped <- master_df %>%
  dplyr::filter(!WellName_std %in% PLATE_MAP$WellName) %>%
  dplyr::count(WellName_std, name = "cells")

if (nrow(unmapped) > 0) {
  warning(sum(unmapped$cells), " cell(s) in ", nrow(unmapped),
          " well(s) are absent from PLATE_MAP and will be dropped: ",
          paste(unmapped$WellName_std, collapse = ", "))
}

missing_from_data <- setdiff(PLATE_MAP$WellName, unique(master_df$WellName_std))
if (length(missing_from_data) > 0) {
  warning("PLATE_MAP lists well(s) with no cells in the data: ",
          paste(missing_from_data, collapse = ", "))
}
if (length(intersect(PLATE_MAP$WellName, unique(master_df$WellName_std))) == 0) {
  stop("No well in PLATE_MAP matches any normalized well name in the data. ",
       "Normalized names look like: ",
       paste(sQuote(head(unique(master_df$WellName_std), 5)), collapse = ", "),
       "; PLATE_MAP expects: ",
       paste(sQuote(head(PLATE_MAP$WellName, 5)), collapse = ", "))
}
stopifnot(!any(duplicated(PLATE_MAP$WellName)))

master_df_raw_annotated <- master_df %>%
  dplyr::left_join(PLATE_MAP, by = c("WellName_std" = "WellName")) %>%
  dplyr::filter(!is.na(Treatment))

master_df_annotated <- master_df_raw_annotated %>%
  dplyr::filter(!WellName_std %in% WELLS_TO_EXCLUDE) %>%
  { if (length(DOSES_TO_EXCLUDE) > 0)
      dplyr::filter(., !Concentration_uM %in% DOSES_TO_EXCLUDE)
    else . }

tibble::tibble(
  Stage = c("Objects in export",
            "Mapped to PLATE_MAP",
            "After WELLS_TO_EXCLUDE / DOSES_TO_EXCLUDE"),
  Cells = c(nrow(master_df), nrow(master_df_raw_annotated), nrow(master_df_annotated))
) %>%
  knitr::kable(caption = "Cell accounting from raw export to annotated dataset",
               format.args = list(big.mark = ","))
```

# Cell Gating (3-Step Cytometric QC Pipeline)

```{r gating-plot, fig.height=4, fig.width=9, warning=FALSE, message=FALSE}
# --- Step 1: Size / intensity pre-gate --------------------------------------
# AREA_MIN keeps pyknotic nuclei; AREA_MAX removes clumps of 3+ merged nuclei
# (previously documented but never actually applied).
master_df_pre <- master_df_annotated %>%
  dplyr::filter(Area > AREA_MIN, Area < AREA_MAX, dapi_intensity > INTENSITY_MIN)

if (nrow(master_df_pre) == 0) {
  stop("The Step 1 pre-gate removed every cell. Check AREA_MIN (", AREA_MIN,
       "), AREA_MAX (", AREA_MAX, ") and INTENSITY_MIN (", INTENSITY_MIN,
       ") against the units of your export.")
}

# --- Step 2: Sphericity-based doublet discrimination -------------------------
sphericity_cutoff <- CIRCULARITY_MIN

if (HAS_SPHERICITY) {
  master_df_pre <- master_df_pre %>%
    dplyr::mutate(Is_Singlet = Sphericity > sphericity_cutoff)

  p_doublets <- ggplot(master_df_pre, aes(x = Area, y = Sphericity)) +
    geom_bin2d(bins = 200) +
    scale_fill_gradientn(colors = c("navy", "deepskyblue", "green", "yellow", "red"), trans = "log10", name = "Cells") +
    geom_hline(yintercept = sphericity_cutoff, color = "black", linetype = "dashed", linewidth = 1) +
    labs(
      title = "Doublet & Shape Discrimination",
      subtitle = paste0("Single cells sit above the Sphericity cut (",
                        sphericity_cutoff, ", dashed line)"),
      x = "Nuclear Volume / Area (pixels²)", y = "Sphericity"
    ) +
    theme_bw(base_size = 8) +
    theme(legend.position = "none")
} else {
  # No Sphericity column: the step is skipped, and the report says so instead
  # of drawing a panel implying a gate that never ran.
  master_df_pre <- master_df_pre %>% dplyr::mutate(Is_Singlet = TRUE)

  p_doublets <- ggplot() +
    annotate("text", x = 0, y = 0, size = 3, lineheight = 1.2,
             label = paste("Step 2 SKIPPED\n\nNo Sphericity column in the export,",
                           "\nso doublets could not be discriminated.",
                           "\nEvery object is treated as a singlet.")) +
    theme_void()
}

# --- Step 3: DNA-content (cell cycle) gate -----------------------------------
master_df_singlets <- master_df_pre %>% dplyr::filter(Is_Singlet)

cc_vars <- c("Area", "integrated_intensity")

# Fit the ellipse on UNTREATED reference wells only. Pooling every singlet lets
# cytotoxic high-dose wells shift the very ellipse that decides which of their
# own cells to keep.
gate_ref <- if (!is.null(GATE_REFERENCE_TREATMENTS)) {
  master_df_singlets %>% dplyr::filter(Treatment %in% GATE_REFERENCE_TREATMENTS)
} else {
  master_df_singlets[0, ]
}

if (nrow(gate_ref) < MIN_GATE_REFERENCE_CELLS) {
  warning("Only ", nrow(gate_ref), " cell(s) in GATE_REFERENCE_TREATMENTS (",
          paste(GATE_REFERENCE_TREATMENTS, collapse = ", "),
          "); need at least ", MIN_GATE_REFERENCE_CELLS,
          ". Falling back to pooling all singlets, which lets treated wells ",
          "influence their own gate.")
  gate_ref <- master_df_singlets
  gate_ref_label <- "all singlets (fallback)"
} else {
  gate_ref_label <- paste(GATE_REFERENCE_TREATMENTS, collapse = " + ")
}

center     <- colMeans(as.data.frame(gate_ref)[, cc_vars])
cov_mat    <- cov(as.data.frame(gate_ref)[, cc_vars])
chi_cutoff <- qchisq(MAHALANOBIS_LEVEL, df = length(cc_vars))

master_df_pre <- master_df_pre %>%
  dplyr::mutate(
    dist_cc = if_else(Is_Singlet,
                      mahalanobis(cbind(Area, integrated_intensity), center = center, cov = cov_mat),
                      NA_real_),
    Keep_Cell = Is_Singlet & (dist_cc <= chi_cutoff)
  )

# The ellipse is drawn from the reference population, so it is the same one the
# gate uses. Note that a bivariate-normal contour on a bimodal DNA distribution
# clips part of the G2/M mode — that is a known limitation of this gate, not an
# accident, so the subtitle says so rather than claiming G2/M is retained.
ellipse_pts <- {
  ang  <- seq(0, 2 * pi, length.out = 200)
  chol_c <- chol(cov_mat)
  circ <- cbind(cos(ang), sin(ang)) * sqrt(chi_cutoff)
  pts  <- circ %*% chol_c
  tibble::tibble(Area = pts[, 1] + center[["Area"]],
                 integrated_intensity = pts[, 2] + center[["integrated_intensity"]])
}

p_cycle <- ggplot(master_df_singlets, aes(x = Area, y = integrated_intensity)) +
  geom_bin2d(bins = 200) +
  scale_fill_gradientn(colors = c("navy", "deepskyblue", "green", "yellow", "red"), trans = "log10", name = "Cells") +
  geom_path(data = ellipse_pts, aes(x = Area, y = integrated_intensity),
            inherit.aes = FALSE, color = "black", linetype = "dashed", linewidth = 1) +
  scale_y_continuous(labels = scales::label_scientific()) +
  labs(
    title = "Cell Cycle & Quality Control",
    subtitle = paste0(MAHALANOBIS_LEVEL * 100, "% Mahalanobis contour fitted on ",
                      gate_ref_label,
                      "\n(a normal contour on a bimodal DNA distribution clips part of the G2/M mode)"),
    x = "Nuclear Volume / Area (pixels²)", y = "Integrated Intensity (Total DNA, a.u.)"
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "right")

# Display side-by-side using Patchwork
qc_plates_plot <- p_doublets + p_cycle
print(qc_plates_plot)

ggsave(file.path(OUTPUT_DIR, "nuclear_gating_qc.tiff"), qc_plates_plot,
       width = 9, height = 4, units = "in", dpi = PLOT_DPI,
       compression = "lzw", device = ragg::agg_tiff)

master_df_gated <- master_df_pre %>% dplyr::filter(Keep_Cell)

if (nrow(master_df_gated) == 0) {
  stop("No cells survived the 3-step gate. Review CIRCULARITY_MIN (",
       CIRCULARITY_MIN, ") and MAHALANOBIS_LEVEL (", MAHALANOBIS_LEVEL,
       ") against the QC panels above.")
}
```

```{r apply-gate}
step_counts <- tibble::tibble(
  Step  = c("0. Annotated cells",
            paste0("1. Area in (", AREA_MIN, ", ", AREA_MAX, ") & DAPI > ", INTENSITY_MIN),
            if (HAS_SPHERICITY) paste0("2. Sphericity > ", CIRCULARITY_MIN)
            else "2. Sphericity gate SKIPPED (column absent)",
            paste0("3. Within ", MAHALANOBIS_LEVEL * 100, "% DNA-content contour")),
  Cells = c(nrow(master_df_annotated), nrow(master_df_pre),
            nrow(master_df_singlets), nrow(master_df_gated))
) %>%
  dplyr::mutate(`% of annotated` = round(100 * Cells / nrow(master_df_annotated), 1))

knitr::kable(step_counts,
             caption = "Cells retained after each gating step",
             format.args = list(big.mark = ","))

counts_before <- dplyr::count(master_df_annotated, Treatment, name = "cells_before")
counts_before %>%
  left_join(dplyr::count(master_df_gated, Treatment, name = "cells_after"), by = "Treatment") %>%
  mutate(cells_after       = replace_na(cells_after, 0),
         percent_remaining = round((cells_after / cells_before) * 100, 1)) %>%
  knitr::kable(caption = "Automated 3-Step Cytometric Gating Summary")
```

# Single-Cell NP Intensity Distributions (Violin Plot)

```{r gating-post-violin, fig.height=5, fig.width=8}
violin_data <- master_df_gated %>%
  filter(Treatment %in% c("Mock", "Mock+Ab", "Virus Only", "Virus + Compound")) %>%
  mutate(Condition = case_when(
    Treatment %in% c("Mock", "Mock+Ab") ~ "Mock",
    Treatment == "Virus Only"           ~ "Virus Only",
    TRUE                                ~ paste0(Concentration_uM, " ", DOSE_UNIT)
  ))

bar_levels <- c("Mock", "Virus Only", mixedsort(unique(violin_data$Condition[grepl(DOSE_UNIT, violin_data$Condition)]), decreasing = TRUE))
violin_data <- violin_data %>% mutate(Condition = factor(Condition, levels = bar_levels))

n_doses  <- sum(grepl(DOSE_UNIT, bar_levels))
fill_pal <- c(
  "Mock"       = "grey80",
  "Virus Only" = "#E41A1C",
  setNames(colorRampPalette(c("#C6DBEF", "#08306B"))(n_doses),
           bar_levels[grepl(DOSE_UNIT, bar_levels)])
)

p_violin <- ggplot(violin_data, aes(x = Condition, y = virus_intensity, fill = Condition)) +
  geom_violin(trim = FALSE, alpha = 0.7, color = "black", bw = 0.15) +
  geom_boxplot(width = 0.1, color = "black", fill = "white", outlier.shape = NA, alpha = 0.8) +
  scale_fill_manual(values = fill_pal, guide = "none") +
  scale_y_log10(labels = scales::label_scientific()) +
  labs(
    title   = "Single-Cell NP Intensity Distributions per Treatment",
    x       = "Treatment Condition",
    y       = "NP Intensity (a.u., log10 scale)",
    caption = "Internal white boxes indicate median and interquartile range (IQR)"
  ) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(p_violin)
```

# Infection Gate: Control-Based Normalization + Population GMM

```{r infection-gate, warning=FALSE, fig.width=6, fig.height=4}
VIRUS_LABEL_VO  <- "Virus Only"
master_df_gated <- master_df_gated %>% mutate(Plate = "P1")

gmm_threshold <- function(mu, sdv, pi_) {
  o <- order(mu); mu <- mu[o]; sdv <- sdv[o]; pi_ <- pi_[o]
  sig1_sq <- sdv[1]^2; sig2_sq <- sdv[2]^2
  A <- 1/(2*sig1_sq) - 1/(2*sig2_sq)
  B <- -mu[1]/sig1_sq + mu[2]/sig2_sq
  C <- mu[1]^2/(2*sig1_sq) - mu[2]^2/(2*sig2_sq) + log(sdv[1]/sdv[2]) - log(pi_[1]/pi_[2])
  disc <- B^2 - 4*A*C
  if (disc >= 0 && abs(A) > 1e-10) {
    roots <- c((-B + sqrt(disc))/(2*A), (-B - sqrt(disc))/(2*A))
    r_valid <- roots[roots > mu[1] & roots < mu[2]]
    if (length(r_valid) > 0) return(r_valid[1])
  }
  (mu[1] + mu[2]) / 2
}

fit_high_mode <- function(xlog) {
  fit <- tryCatch(Mclust(xlog, G = 2, modelNames = "V", verbose = FALSE), error = function(e) NULL)
  if (!is.null(fit)) return(max(as.numeric(fit$parameters$mean)))
  as.numeric(quantile(xlog, 0.75, na.rm = TRUE))
}

# 1. Compute Anchors (L_p, H_p)
mock_cells <- master_df_gated %>% filter(Treatment %in% c("Mock", "Mock+Ab")) %>% mutate(xlog = log10(virus_intensity + 1))
L_p_val    <- if (nrow(mock_cells) > 0) median(mock_cells$xlog, na.rm = TRUE) else 2.0

vo_cells   <- master_df_gated %>% filter(Treatment == VIRUS_LABEL_VO) %>% mutate(xlog = log10(virus_intensity + 1))
H_p_val    <- if (nrow(vo_cells) > 0) fit_high_mode(vo_cells$xlog) else 4.0

# 2. Normalize Cells
master_df_norm <- master_df_gated %>%
  mutate(
    xlog   = log10(virus_intensity + 1),
    L_p    = L_p_val,
    H_p    = H_p_val,
    span   = H_p - L_p,
    x_norm = if_else(is.na(span) | span < 1e-6, NA_real_, (xlog - L_p) / span)
  )

# 3. Fit Population 2-GMM Gate
infected_culture_norm <- master_df_norm %>% filter(Treatment %in% c(VIRUS_LABEL_VO, "Virus + Compound"))
med_val <- suppressWarnings(median(table(infected_culture_norm$WellName_std), na.rm = TRUE))
target_n_gate <- if (is.finite(med_val) && med_val > 0) as.integer(round(med_val)) else 200L

# Down-sample each well to the same size so no single well dominates the GMM.
# slice_sample() takes a RANDOM subset; filter(row_number() <= n) took the first
# n cells in acquisition order, which is a systematically biased "balanced"
# sample that the set.seed() call made look reproducible rather than correct.
set.seed(RANDOM_SEED)
gate_balanced <- infected_culture_norm %>%
  group_by(WellName_std) %>%
  slice_sample(n = target_n_gate, replace = FALSE) %>%
  ungroup()

x_gate <- gate_balanced$x_norm[is.finite(gate_balanced$x_norm)]
fit_gate <- tryCatch(
  Mclust(x_gate, G = 2, modelNames = "V", verbose = FALSE),
  error = function(e) NULL
)

# GATE_FALLBACK_THRESHOLD is only reached when the 2-component mixture cannot be
# fitted. Every percent-infection number and both EC50s depend on this value, so
# falling back is reported loudly instead of silently.
GATE_FALLBACK_THRESHOLD <- 0.5

gate_is_fitted <- !is.null(fit_gate) &&
  !is.null(fit_gate$parameters$mean) &&
  length(fit_gate$parameters$mean) == 2

if (gate_is_fitted) {
  mu_gate  <- as.numeric(fit_gate$parameters$mean)
  sdv_gate <- sqrt(fit_gate$parameters$variance$sigmasq)
  pi_gate  <- as.numeric(fit_gate$parameters$pro)
  if (length(sdv_gate) == 1L) sdv_gate <- rep(sdv_gate, 2L)
  threshold_final <- gmm_threshold(mu_gate, sdv_gate, pi_gate)
} else {
  threshold_final <- GATE_FALLBACK_THRESHOLD
  warning("The 2-component GMM infection gate could not be fitted. ",
          "Falling back to a FIXED threshold of ", GATE_FALLBACK_THRESHOLD,
          " on the normalized scale. Every percent-infection value and both ",
          "EC50 estimates below are conditional on this arbitrary cut — treat ",
          "them as indicative only, and inspect the density plot.")
}
master_df_final <- master_df_norm

# Display GMM Infection Gate Density Plot
ctrl_dens_df <- master_df_final %>%
  filter(Treatment %in% c("Mock", "Mock+Ab", VIRUS_LABEL_VO)) %>%
  mutate(Group = if_else(Treatment %in% c("Mock", "Mock+Ab"), "Mock (Negative Control)", "Virus Only (Positive Control)"))

gate_label <- if (gate_is_fitted) {
  paste0("GMM Infection Gate = ", round(threshold_final, 3))
} else {
  paste0("FALLBACK Gate = ", round(threshold_final, 3), " (GMM fit failed)")
}

p_gmm_density <- ggplot(ctrl_dens_df, aes(x = x_norm, fill = Group)) +
  geom_density(alpha = 0.55, color = "black", linewidth = 0.3) +
  geom_vline(xintercept = threshold_final, linetype = "dashed", color = "firebrick3", linewidth = 0.9) +
  annotate("label", x = threshold_final, y = 1.1,
           label = gate_label,
           fill = "firebrick3", color = "white", fontface = "bold", size = 3, linewidth = 0) +
  scale_fill_manual(values = c("Mock (Negative Control)" = "grey60", "Virus Only (Positive Control)" = "#E41A1C")) +
  labs(
    title = "Control-Based GMM Population Infection Gate",
    subtitle = "Normalized virus intensity distribution (0 = Mock median, 1 = Virus Only high mode)",
    x = "Normalized Virus Intensity (log10 scale)",
    y = "Density",
    fill = "Control Group"
  ) +
  theme_bw(base_size = 9) +
  theme(legend.position = "top", plot.title = element_text(face = "bold"))

print(p_gmm_density)
ggsave(file.path(OUTPUT_DIR, "infection_gate_gmm_density.tiff"), p_gmm_density,
       width = 5.5, height = 3.8, units = "in", dpi = PLOT_DPI)
```

# Single-Cell Population Shift Across Doses (Ridgeline Plot)

```{r ridgeline-plot, warning=FALSE, message=FALSE, fig.width=6, fig.height=5.5}
treatment_order <- c("Mock+Ab", "Mock", "Virus Only", "Virus + Compound")

bdgr_levels <- master_df_final %>%
  filter(Treatment == "Virus + Compound") %>%
  distinct(Concentration_uM) %>%
  arrange(Concentration_uM) %>%
  pull(Concentration_uM)

plot_df <- master_df_final %>%
  filter(Treatment %in% treatment_order) %>%
  mutate(
    group_label = case_when(
      Treatment %in% c("Mock", "Mock+Ab") ~ "Mock",
      Treatment == "Virus Only"           ~ "Virus Only",
      TRUE ~ paste0(Concentration_uM, " ", DOSE_UNIT, " ", COMPOUND_NAME)
    ),
    group_label = fct_inorder(group_label)
  ) %>%
  mutate(
    fill_key = case_when(
      Treatment %in% c("Mock", "Mock+Ab") ~ "Mock",
      Treatment == "Virus Only"           ~ "Virus Only",
      TRUE ~ paste0(COMPOUND_NAME, "_", Concentration_uM)
    ),
    fill_key = factor(fill_key)
  )

bdgr_fill_keys <- paste0(COMPOUND_NAME, "_", bdgr_levels)
n_bdgr         <- length(bdgr_fill_keys)
bdgr_cols      <- colorRampPalette(RColorBrewer::brewer.pal(9, "Blues"))(n_bdgr)

fill_values <- c(
  setNames("grey80", "Mock"),
  setNames("#D55E00", "Virus Only"),
  setNames(bdgr_cols, bdgr_fill_keys)
)

# Infection threshold back-transformed to the raw fluorescence scale.
# The forward transform is xlog = log10(intensity + 1), so the inverse must
# subtract that same 1 — otherwise the dashed line is drawn at the wrong place.
raw_threshold <- 10^(L_p_val + threshold_final * (H_p_val - L_p_val)) - 1

p_ridgeline <- ggplot(plot_df, aes(x = virus_intensity, y = group_label, fill = fill_key)) +
  geom_density_ridges(alpha = 0.8, scale = 3, rel_min_height = 0.01) +
  geom_vline(xintercept = raw_threshold, color = "black", linetype = "dashed", linewidth = 1) +
  scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x)), minor_breaks = NULL) +
  annotation_logticks(sides = "b", short = unit(1.2, "mm"), mid = unit(1.6, "mm"), long = unit(2.2, "mm")) +
  scale_fill_manual(values = fill_values, breaks = names(fill_values)) +
  labs(
    title = paste("Effect of", COMPOUND_NAME, "on", VIRUS_NAME, "NP signal"),
    subtitle = "Single-cell fluorescence distributions; dashed line indicates GMM infection gate",
    x = "NP channel fluorescence (a.u., log10 scale)",
    y = NULL
  ) +
  theme_bw(base_size = 8) +
  theme(
    legend.position   = "none",
    axis.ticks.length = unit(2.2, "mm"),
    panel.grid.minor  = element_blank(),
    plot.title        = element_text(face = "bold", size = 10)
  ) +
  coord_cartesian(clip = "off")

print(p_ridgeline)
ggsave(file.path(OUTPUT_DIR, "compound_ridgeline.tiff"), p_ridgeline,
       width = 5.5, height = 4.5, units = "in", dpi = PLOT_DPI,
       compression = "lzw", device = ragg::agg_tiff)
```


# Well-Level Summary

```{r well-summary}
total_counts <- master_df_final %>% dplyr::count(Treatment, Concentration_uM, WellName_std, name = "total_cell_count")
positive_counts <- master_df_final %>% filter(x_norm > threshold_final) %>% dplyr::count(Treatment, Concentration_uM, WellName_std, name = "positive_cell_count")

percent_summary <- total_counts %>%
  left_join(positive_counts, by = c("Treatment", "Concentration_uM", "WellName_std")) %>%
  mutate(
    positive_cell_count = replace_na(positive_cell_count, 0),
    percent_positive     = (positive_cell_count / total_cell_count) * 100,
    Plate = "Plate 1",
    Precision_Flag = if_else(total_cell_count < PRECISION_WARN_CELLS,
                             "Low Precision", "Normal"),
    # Precision_Flag used to be displayed and then ignored. It now drives a real
    # decision: wells below MIN_CELLS_PER_WELL are excluded from every 4PL fit,
    # while wells between that and PRECISION_WARN_CELLS are kept but flagged.
    Include_In_Fit = total_cell_count >= MIN_CELLS_PER_WELL
  )

excluded_wells <- percent_summary %>% filter(!Include_In_Fit)
if (nrow(excluded_wells) > 0) {
  message(nrow(excluded_wells), " well(s) below MIN_CELLS_PER_WELL (",
          MIN_CELLS_PER_WELL, ") excluded from the dose-response fits: ",
          paste(excluded_wells$WellName_std, collapse = ", "))
}
```

```{r fit-helpers}
## ── Shared 4PL machinery ────────────────────────────────────────────────────
## One fitter, used for all three curves in this document: relative %
## infection, normalized NP intensity, and host-cell viability (CC50).
## The inflection parameter is always named `ec50` internally; it is CC50 when
## the response is viability.
fit_4pl <- function(data, response, dose = "dose_fit", weights = NULL,
                    starts = c(0.5, 1, 2, 5, 8, 10, 20, 40, 0.1),
                    lower = c(ec50 = 1e-4, hill = 0.05, bottom = -20, top = 80),
                    upper = c(ec50 = 1e3,  hill = 50,   bottom = 50,  top = 120)) {
  if (nrow(data) < 4) return(NULL)
  form <- stats::as.formula(
    paste0(response, " ~ bottom + (top - bottom) / (1 + (", dose, " / ec50)^hill)")
  )
  for (ec50_try in starts) {
    args <- list(
      formula = form,
      data    = data,
      start   = list(ec50 = ec50_try, hill = 1.0, bottom = 0, top = 100),
      lower   = lower,
      upper   = upper,
      control = minpack.lm::nls.lm.control(maxiter = 500)
    )
    # `weights` must be OMITTED, not passed as NULL: nlsLM() treats an explicit
    # NULL as a zero-length weight vector and aborts with
    # "evaluation of fn function returns non-sensible value!".
    if (!is.null(weights)) args$weights <- weights

    fit <- tryCatch(do.call(minpack.lm::nlsLM, args), error = function(e) NULL)
    if (!is.null(fit)) return(fit)
  }
  NULL
}

## ── Parametric confidence interval for the inflection point ─────────────────
## This is NOT a bootstrap: it does not resample the data. It draws parameter
## vectors from the multivariate normal implied by the fit's coefficient
## covariance (the usual asymptotic/delta-method approximation) and reads the
## quantiles of the induced EC50/CC50 distribution.
## Sampling is done on the LOG scale because EC50 is strictly positive and
## right-skewed — a normal approximation on the raw scale puts mass on
## impossible negative concentrations and produces asymmetric coverage.
ec50_ci_parametric <- function(fit, draws = EC50_CI_DRAWS, seed = RANDOM_SEED,
                               level = 0.95) {
  na3 <- c(lower = NA_real_, median = NA_real_, upper = NA_real_)
  if (is.null(fit)) return(na3)

  mu <- coef(fit)
  vc <- tryCatch(vcov(fit), error = function(e) NULL)
  if (is.null(vc) || !("ec50" %in% names(mu))) return(na3)

  # Reparameterize to log(ec50) via the delta method, then draw there.
  ec50_hat <- unname(mu[["ec50"]])
  se_ec50  <- suppressWarnings(sqrt(vc["ec50", "ec50"]))
  if (!is.finite(ec50_hat) || ec50_hat <= 0 ||
      !is.finite(se_ec50) || se_ec50 <= 0) return(na3)

  se_log <- se_ec50 / ec50_hat            # delta method: SE[log X] ~ SE[X] / X
  set.seed(seed)
  log_draws <- stats::rnorm(draws, mean = log(ec50_hat), sd = se_log)
  ec <- exp(log_draws)
  ec <- ec[is.finite(ec)]
  if (length(ec) < 10) return(na3)

  a <- (1 - level) / 2
  c(lower  = unname(stats::quantile(ec, a)),
    median = unname(stats::quantile(ec, 0.5)),
    upper  = unname(stats::quantile(ec, 1 - a)))
}
```

# Host Cell Viability

```{r toxicity-plot, warning=FALSE, message=FALSE, fig.width=6, fig.height=4}
viab_summary <- percent_summary %>%
  filter(Treatment %in% c("Mock", "Virus Only", "Virus + Compound", "Toxicity Control")) %>%
  mutate(Dose_Val = case_when(
    Treatment == "Mock"       ~ 0,
    Treatment == "Virus Only" ~ 0,
    TRUE                      ~ Concentration_uM
  )) %>%
  group_by(Treatment, Dose_Val) %>%
  summarise(
    n_wells    = n(),
    mean_cells = mean(total_cell_count, na.rm = TRUE),
    sd_cells   = if_else(n_wells > 1, sd(total_cell_count, na.rm = TRUE), 0),
    se_cells   = if_else(n_wells > 1, sd_cells / sqrt(n_wells), 0),
    .groups    = "drop"
  )

p_toxicity <- ggplot(viab_summary, aes(x = factor(Dose_Val), y = mean_cells, fill = Treatment)) +
  geom_col(color = "black", alpha = 0.85, width = 0.65) +
  geom_errorbar(aes(ymin = pmax(0, mean_cells - se_cells), ymax = mean_cells + se_cells), width = 0.25, linewidth = 0.5) +
  scale_fill_manual(values = c(
    "Mock"             = "grey70",
    "Virus Only"       = "#E41A1C",
    "Virus + Compound" = "#2B5C8F",
    "Toxicity Control" = "#7570B3"
  )) +
  labs(
    title = "Host Cell Viability Across Concentrations",
    subtitle = "Well cell counts (mean ± SE across replicates)",
    x = paste0("Concentration (", DOSE_UNIT, ")"),
    y = "Mean Recovered Nuclei / Well",
    fill = "Group"
  ) +
  theme_bw(base_size = 9) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "top", plot.title = element_text(face = "bold"))

print(p_toxicity)
ggsave(file.path(OUTPUT_DIR, "host_cell_viability.tiff"), p_toxicity,
       width = 5.5, height = 3.8, units = "in", dpi = PLOT_DPI)
```

# Percent Infection per Well Table

```{r percent-infection-table}
percent_summary %>%
  mutate(
    Dose_Label = case_when(
      Treatment %in% c("Mock", "Mock+Ab") ~ "Mock",
      Treatment == "Virus Only"           ~ "0 µM (Control)",
      TRUE                                ~ paste0(Concentration_uM, " µM")
    ),
    Status_Label = case_when(
      !Include_In_Fit             ~ paste0("Excluded from fits (< ",
                                           MIN_CELLS_PER_WELL, " cells)"),
      Precision_Flag == "Normal"  ~ "Valid",
      TRUE                        ~ Precision_Flag
    )
  ) %>%
  dplyr::select(
    Well              = WellName_std,
    Treatment,
    Dose              = Dose_Label,
    `Total Cells`     = total_cell_count,
    `Infected Cells`  = positive_cell_count,
    `% Infection`    = percent_positive,
    `QC Status`       = Status_Label
  ) %>%
  knitr::kable(
    caption     = "Well-Level Percent Infection (Global Population GMM Threshold)",
    digits      = 1,
    format.args = list(big.mark = ",")
  )
```

# Co-Primary Endpoint 1: % Infection Dose-Response (4PL Curve)

```{r dose-response-model}
vo_anchor   <- mean(percent_summary$percent_positive[percent_summary$Treatment == "Virus Only"], na.rm = TRUE)
mock_anchor <- mean(percent_summary$percent_positive[percent_summary$Treatment %in% c("Mock", "Mock+Ab")], na.rm = TRUE)
if (is.nan(mock_anchor) || is.na(mock_anchor)) mock_anchor <- 0

# The normalization denominator is the assay window. If the infection failed
# (or the controls are missing) it collapses to ~0 and every normalized value
# becomes +/-Inf, which then silently feeds the 4PL. Refuse instead.
assay_window <- vo_anchor - mock_anchor
if (!is.finite(assay_window) || abs(assay_window) < 1) {
  stop("Assay window (Virus Only - Mock percent infection) is ",
       signif(assay_window, 3), " percentage points. ",
       "Relative normalization is not meaningful: check that the Virus Only ",
       "wells are actually infected and that the infection gate is sensible.")
}

summary_rel <- percent_summary %>%
  mutate(
    percent_infection_rel = 100 * (percent_positive - mock_anchor) / assay_window
  )

drc_data_rel <- summary_rel %>% filter(Treatment == "Virus + Compound")
pos_conc  <- drc_data_rel$Concentration_uM[drc_data_rel$Concentration_uM > 0 & is.finite(drc_data_rel$Concentration_uM)]
min_pos   <- if (length(pos_conc) > 0) min(pos_conc) else 0.01
zero_tick <- min_pos / ZERO_DOSE_TICK_DIVISOR

# INCLUDE_ZERO_DOSE_IN_FIT = FALSE (default): the Virus Only control is plotted
# at `zero_tick` because a log axis has no zero, but that tick is an artifact of
# the axis, not a measured concentration. Feeding it to the regression as a real
# dose pulls the `top` asymptote and shifts EC50.
fit_treatments_rel <- if (INCLUDE_ZERO_DOSE_IN_FIT) {
  c("Virus Only", "Virus + Compound")
} else {
  "Virus + Compound"
}

fit_data_wide <- summary_rel %>%
  filter(Treatment %in% fit_treatments_rel, Include_In_Fit) %>%
  mutate(
    dose_fit   = ifelse(Treatment == "Virus Only", zero_tick, Concentration_uM),
    prop_raw   = positive_cell_count / total_cell_count,
    prop_clamp = pmin(pmax(prop_raw, 1e-3), 1 - 1e-3),
    bin_weight = total_cell_count / (prop_clamp * (1 - prop_clamp))
  )

fit_rel <- fit_4pl(fit_data_wide, "percent_infection_rel",
                   weights = fit_data_wide$bin_weight)
ci_rel  <- ec50_ci_parametric(fit_rel)
```

## Relative % Infection 4PL Dose-Response Plot

```{r dose-response-plot, message=FALSE, warning=FALSE, fig.width=6, fig.height=4.5}
max_pos <- if (length(pos_conc) > 0) max(pos_conc) else 100
newx    <- exp(seq(log(zero_tick), log(max_pos), length.out = 300))

pred_df_rel <- tibble(Concentration_uM = numeric(0), y = numeric(0))
if (!is.null(fit_rel)) {
  coef_rel   <- coef(fit_rel)
  ec50_rel   <- coef_rel["ec50"]
  hill_rel   <- coef_rel["hill"]
  bottom_rel <- coef_rel["bottom"]
  top_rel    <- coef_rel["top"]
  pred_df_rel <- tibble(
    Concentration_uM = newx,
    y = bottom_rel + (top_rel - bottom_rel) / (1 + (Concentration_uM / ec50_rel)^hill_rel)
  )
  ec50_rel_lab <- paste0("EC50 = ", signif(ec50_rel, 3), " ", DOSE_UNIT)
} else {
  ec50_rel_lab <- "Fit Failed"
}

pts_sum <- summary_rel %>%
  filter(Treatment %in% c("Virus Only", "Virus + Compound")) %>%
  mutate(x_plot = ifelse(Concentration_uM == 0, zero_tick, Concentration_uM)) %>%
  group_by(x_plot) %>%
  summarise(n  = n(),
            y  = mean(percent_infection_rel, na.rm = TRUE),
            se = ifelse(n > 1, sd(percent_infection_rel, na.rm = TRUE) / sqrt(n), 0),
            .groups = "drop") %>%
  mutate(y    = ifelse(x_plot == zero_tick, 100, y),
         ymin = pmax(0,   y - se),
         ymax = pmin(105, y + se))

# Count DISTINCT doses, not wells: with replicate wells the old `length(pos_conc)`
# was a well count, so a 3-dose experiment in triplicate took the decade-break
# branch meant for dense dose series.
if (length(unique(pos_conc)) > LOG_TICK_DOSE_COUNT) {
  raw_brks <- 10^seq(floor(log10(min_pos)), ceiling(log10(max_pos)))
  dose_breaks <- sort(unique(c(zero_tick, raw_brks[raw_brks >= min_pos & raw_brks <= max_pos], max_pos)))
} else {
  dose_breaks <- sort(unique(c(zero_tick, drc_data_rel$Concentration_uM)))
}
dose_labels_vec <- c("0", as.character(dose_breaks[-1]))

p_dr_rel <- ggplot() +
  geom_errorbar(data = pts_sum, aes(x = x_plot, ymin = ymin, ymax = ymax),
                width = 0.07, linewidth = 0.6, colour = "#2B5C8F") +
  geom_point(data = pts_sum, aes(x = x_plot, y = y),
             shape = 21, size = 2.8, stroke = 0.5,
             fill = "#2B5C8F", colour = "black")

if (!is.null(fit_rel)) {
  y_rel_halfway <- bottom_rel + (top_rel - bottom_rel) / 2
  p_dr_rel <- p_dr_rel +
    geom_line(data = pred_df_rel, aes(x = Concentration_uM, y = y),
              color = "#2B5C8F", linewidth = 1.0) +
    geom_segment(aes(x = ec50_rel, xend = ec50_rel, y = 0, yend = y_rel_halfway),
                 linetype = "dashed", color = "firebrick3", linewidth = 0.6) +
    geom_segment(aes(x = zero_tick, xend = ec50_rel, y = y_rel_halfway, yend = y_rel_halfway),
                 linetype = "dashed", color = "firebrick3", linewidth = 0.6) +
    annotate("label", x = ec50_rel, y = 25, label = ec50_rel_lab,
             fill = "#2B5C8F", color = "white", fontface = "bold", size = 3, linewidth = 0)
}

p_dr_rel <- p_dr_rel +
  scale_x_log10(limits = c(zero_tick * 0.8, max_pos * 1.2), breaks = dose_breaks, labels = dose_labels_vec, minor_breaks = NULL) +
  annotation_logticks(sides = "b", short = unit(1.2, "mm"), mid = unit(1.6, "mm"), long = unit(2.2, "mm")) +
  labs(
    title = paste(COMPOUND_NAME, "relative % infection dose–response"),
    subtitle = "Points show mean ± SE across replicate wells; curve fitted via 4PL regression",
    x     = paste0(COMPOUND_NAME, " (", DOSE_UNIT, ", log10 scale)"),
    y     = "% Infected Cells (Relative to Control)"
  ) +
  coord_cartesian(ylim = c(0, 110), clip = "off") +
  theme_bw(base_size = 9) +
  theme(
    plot.title        = element_text(face = "bold", size = 10, margin = margin(b = 4)),
    axis.title        = element_text(face = "bold", size = 9),
    axis.text         = element_text(size = 8),
    axis.line         = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks        = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks.length = unit(2.2, "mm"),
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    legend.position   = "none",
    plot.margin       = margin(6, 10, 6, 8)
  )

print(p_dr_rel)
ggsave(file.path(OUTPUT_DIR, "compound_dose_response_relative.tiff"), p_dr_rel,
       width = 5.5, height = 4.0, units = "in", dpi = PLOT_DPI)
```

# Cytotoxicity: CC50 and Selectivity Index

Host-cell viability is expressed as the number of recovered nuclei per well
relative to the untreated mock wells, and fitted with the **same 4PL machinery**
used for the antiviral endpoints. The concentration at which viability is
half-maximal is the **CC50**. The **selectivity index**, `SI = CC50 / EC50`,
is the margin between the antiviral and cytotoxic concentrations — a compound
whose EC50 is only achievable at cytotoxic doses has no therapeutic window,
however good its EC50 looks in isolation.

```{r cc50-model, warning=FALSE, message=FALSE}
mock_wells <- percent_summary %>% filter(Treatment == "Mock")
mock_mean_cells <- mean(mock_wells$total_cell_count, na.rm = TRUE)

viab_ok <- is.finite(mock_mean_cells) && mock_mean_cells > 0 &&
           any(percent_summary$Treatment == "Toxicity Control")

if (!viab_ok) {
  warning("Cannot compute CC50: need Mock wells (viability reference) and ",
          "'Toxicity Control' wells (compound without virus) in PLATE_MAP.")
}

cc50_data <- percent_summary %>%
  filter(Treatment == "Toxicity Control",
         is.finite(Concentration_uM), Concentration_uM > 0) %>%
  mutate(
    dose_fit      = Concentration_uM,
    viability_pct = 100 * total_cell_count / mock_mean_cells
  )

# Viability is bounded at 100% by construction, so allow `top` to sit near 100
# and `bottom` to fall to 0 — the antiviral fit's tighter bounds do not apply.
fit_cc50 <- if (viab_ok) {
  fit_4pl(
    cc50_data, "viability_pct",
    starts = c(20, 40, 10, 80, 5, 2, 1),
    # Viability is a percentage of the mock control: it cannot go below 0, and
    # its upper plateau should sit near 100.
    lower  = c(ec50 = 1e-4, hill = 0.05, bottom = 0,  top = 50),
    upper  = c(ec50 = 1e4,  hill = 50,   bottom = 50, top = 150)
  )
} else NULL

ci_cc50  <- ec50_ci_parametric(fit_cc50)
cc50_val <- if (!is.null(fit_cc50)) unname(coef(fit_cc50)["ec50"]) else NA_real_
ec50_val <- if (!is.null(fit_rel))  unname(coef(fit_rel)["ec50"])  else NA_real_

# A CC50 outside the tested dose range is an extrapolation, not a measurement.
tested_max <- max(cc50_data$Concentration_uM, default = NA_real_)
min_viab   <- suppressWarnings(min(cc50_data$viability_pct, na.rm = TRUE))
cc50_extrapolated <- is.finite(cc50_val) && is.finite(tested_max) &&
                     (cc50_val > tested_max || min_viab > 50)

selectivity_index <- if (is.finite(cc50_val) && is.finite(ec50_val) && ec50_val > 0) {
  cc50_val / ec50_val
} else NA_real_

fmt_ci <- function(ci) {
  if (any(!is.finite(ci[c("lower", "upper")]))) return("–")
  paste0(signif(ci[["lower"]], 3), " – ", signif(ci[["upper"]], 3))
}

tibble::tibble(
  Parameter = c(paste0("EC50 (% infection endpoint, ", DOSE_UNIT, ")"),
                paste0("CC50 (host cell viability, ", DOSE_UNIT, ")"),
                "Selectivity Index (CC50 / EC50)"),
  Estimate  = c(ec50_val, cc50_val, selectivity_index),
  `95% CI`  = c(fmt_ci(ci_rel), fmt_ci(ci_cc50), "–")
) %>%
  knitr::kable(
    caption = paste0("Antiviral potency, cytotoxicity and selectivity. ",
                     "Confidence intervals are parametric (log-scale) ",
                     "approximations from the fit covariance."),
    digits  = 3
  )

if (cc50_extrapolated) {
  cat("\n**Warning:** viability never falls below 50% within the tested range",
      "(lowest observed:", round(min_viab, 1), "% at", tested_max, DOSE_UNIT,
      "). The CC50 above is an EXTRAPOLATION beyond the highest tested dose and",
      "the selectivity index is therefore a lower bound only. Extend the dose",
      "series to bracket 50% viability before quoting a CC50.\n")
} else if (is.null(fit_cc50)) {
  cat("\n**The CC50 curve could not be fitted.** No selectivity index is reported.\n")
}
```

```{r cc50-plot, warning=FALSE, message=FALSE, fig.width=6, fig.height=4.5}
cc50_pts <- cc50_data %>%
  group_by(Concentration_uM) %>%
  summarise(n  = dplyr::n(),
            y  = mean(viability_pct, na.rm = TRUE),
            se = if_else(dplyr::n() > 1, sd(viability_pct, na.rm = TRUE) / sqrt(dplyr::n()), 0),
            .groups = "drop")

p_cc50 <- ggplot() +
  geom_hline(yintercept = 50, linetype = "dotted", colour = "grey50", linewidth = 0.5) +
  geom_errorbar(data = cc50_pts,
                aes(x = Concentration_uM, ymin = pmax(0, y - se), ymax = y + se),
                width = 0.07, linewidth = 0.6, colour = "#7570B3") +
  geom_point(data = cc50_pts, aes(x = Concentration_uM, y = y),
             shape = 21, size = 2.8, stroke = 0.5, fill = "#7570B3", colour = "black")

if (!is.null(fit_cc50)) {
  cf_t     <- coef(fit_cc50)
  tox_newx <- exp(seq(log(min(cc50_data$Concentration_uM)),
                      log(max(cc50_data$Concentration_uM)), length.out = 300))
  pred_cc50 <- tibble::tibble(
    Concentration_uM = tox_newx,
    y = cf_t[["bottom"]] + (cf_t[["top"]] - cf_t[["bottom"]]) /
        (1 + (tox_newx / cf_t[["ec50"]])^cf_t[["hill"]])
  )
  cc50_lab <- paste0("CC50 = ", signif(cc50_val, 3), " ", DOSE_UNIT,
                     if (is.finite(selectivity_index))
                       paste0("\nSI = ", signif(selectivity_index, 3)) else "")
  p_cc50 <- p_cc50 +
    geom_line(data = pred_cc50, aes(x = Concentration_uM, y = y),
              colour = "#7570B3", linewidth = 1.0) +
    geom_vline(xintercept = cc50_val, linetype = "dashed",
               colour = "firebrick3", linewidth = 0.6) +
    annotate("label", x = cc50_val, y = 25, label = cc50_lab,
             fill = "#7570B3", colour = "white", fontface = "bold",
             size = 3, linewidth = 0)
}

if (!is.null(fit_rel)) {
  p_cc50 <- p_cc50 +
    geom_vline(xintercept = ec50_val, linetype = "dashed",
               colour = "#2B5C8F", linewidth = 0.6) +
    annotate("text", x = ec50_val, y = 103, label = "EC50",
             colour = "#2B5C8F", fontface = "bold", size = 3)
}

p_cc50 <- p_cc50 +
  scale_x_log10(minor_breaks = NULL) +
  annotation_logticks(sides = "b", short = unit(1.2, "mm"),
                      mid = unit(1.6, "mm"), long = unit(2.2, "mm")) +
  coord_cartesian(ylim = c(0, 110), clip = "off") +
  labs(
    title    = paste(COMPOUND_NAME, "cytotoxicity (CC50)"),
    subtitle = "Recovered nuclei per well relative to untreated Mock; 4PL fit on the drug-only arm",
    x        = paste0(COMPOUND_NAME, " (", DOSE_UNIT, ", log10 scale)"),
    y        = "Host Cell Viability (% of Mock)"
  ) +
  theme_bw(base_size = 9) +
  theme(
    plot.title       = element_text(face = "bold", size = 10),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

print(p_cc50)
ggsave(file.path(OUTPUT_DIR, "host_cell_viability_cc50.tiff"), p_cc50,
       width = 5.5, height = 4.0, units = "in", dpi = PLOT_DPI)
```

# Co-Primary Endpoint 2: Normalized NP Intensity Dose-Response

```{r intensity-dose-response, warning=FALSE}
intensity_well <- master_df_final %>%
  filter(Treatment %in% c("Mock", "Mock+Ab", "Virus Only", "Virus + Compound")) %>%
  group_by(WellName_std, Treatment, Concentration_uM) %>%
  summarise(
    n_cells     = n(),
    mean_xlog   = mean(xlog, na.rm = TRUE),
    L_p = first(L_p), H_p = first(H_p),
    .groups = "drop"
  ) %>%
  mutate(intensity_norm_mean = 100 * (mean_xlog - L_p) / (H_p - L_p))

vo_i   <- mean(intensity_well$intensity_norm_mean[intensity_well$Treatment == "Virus Only"], na.rm = TRUE)
mock_i <- mean(intensity_well$intensity_norm_mean[intensity_well$Treatment %in% c("Mock", "Mock+Ab")], na.rm = TRUE)

# Same guard as the % infection endpoint: an assay window near zero turns every
# normalized value into +/-Inf and silently feeds it to the 4PL.
window_i <- vo_i - mock_i
if (!is.finite(window_i) || abs(window_i) < 1) {
  stop("Intensity assay window (Virus Only - Mock normalized intensity) is ",
       signif(window_i, 3), " — too small to normalize against. Check the ",
       "control wells.")
}

intensity_rel <- intensity_well %>%
  mutate(intensity_rel_mean = 100 * (intensity_norm_mean - mock_i) / window_i)

pos_conc_i  <- intensity_rel$Concentration_uM[intensity_rel$Concentration_uM > 0 & is.finite(intensity_rel$Concentration_uM)]
min_pos_i   <- if (length(pos_conc_i) > 0) min(pos_conc_i) else 0.01
zero_tick_i <- min_pos_i / ZERO_DOSE_TICK_DIVISOR

fit_treatments_i <- if (INCLUDE_ZERO_DOSE_IN_FIT) {
  c("Virus Only", "Virus + Compound")
} else {
  "Virus + Compound"
}

fit_i <- intensity_rel %>%
  filter(Treatment %in% fit_treatments_i, n_cells >= MIN_CELLS_PER_WELL) %>%
  mutate(dose_fit = ifelse(Treatment == "Virus Only", zero_tick_i, Concentration_uM))

fit_int <- fit_4pl(fit_i, "intensity_rel_mean", weights = fit_i$n_cells)
ci_int  <- ec50_ci_parametric(fit_int)

ec50_compare <- tibble(
  Endpoint   = c("Normalized NP intensity (mean) -- co-primary",
                 "% infected (population GMM gate) -- co-primary"),
  `EC50 (µM)` = c(
    if (!is.null(fit_int)) unname(coef(fit_int)["ec50"]) else NA_real_,
    if (!is.null(fit_rel))  unname(coef(fit_rel)["ec50"])  else NA_real_
  ),
  `95% CI`    = c(fmt_ci(ci_int), fmt_ci(ci_rel))
)
knitr::kable(ec50_compare, caption = "EC50 comparison across co-primary endpoints.", digits = 3)
```

## Intensity Dose-Response Plot

```{r intensity-dose-response-plot, warning=FALSE, message=FALSE, fig.width=6, fig.height=4.5}
pts_i <- intensity_rel %>%
  filter(Treatment %in% c("Virus Only", "Virus + Compound")) %>%
  mutate(
    x_plot = ifelse(Treatment == "Virus Only", zero_tick_i, Concentration_uM),
    x_plot = ifelse(is.na(x_plot), NA_real_, x_plot)
  ) %>%
  filter(!is.na(x_plot)) %>%
  group_by(x_plot) %>%
  summarise(n  = n(),
            y  = mean(intensity_rel_mean, na.rm = TRUE),
            se = ifelse(n > 1, sd(intensity_rel_mean, na.rm = TRUE) / sqrt(n), 0),
            .groups = "drop") %>%
  mutate(y    = ifelse(x_plot == zero_tick_i, 100, y),
         ymin = pmax(0,   y - se),
         ymax = pmin(105, y + se))

max_pos_i <- if (length(pos_conc_i) > 0) max(pos_conc_i) else 100
newx_i    <- exp(seq(log(zero_tick_i), log(max_pos_i), length.out = 300))

pred_i <- tibble(Concentration_uM = numeric(0), y = numeric(0))
if (!is.null(fit_int)) {
  cf <- coef(fit_int)
  pred_i <- tibble(
    Concentration_uM = newx_i,
    y = cf["bottom"] + (cf["top"] - cf["bottom"]) / (1 + (Concentration_uM / cf["ec50"]) ^ cf["hill"])
  )
}

if (length(unique(pos_conc_i)) > LOG_TICK_DOSE_COUNT) {
  raw_brks_i <- 10^seq(floor(log10(min_pos_i)), ceiling(log10(max_pos_i)))
  dose_breaks_i <- sort(unique(c(zero_tick_i, raw_brks_i[raw_brks_i >= min_pos_i & raw_brks_i <= max_pos_i], max_pos_i)))
} else {
  dose_breaks_i <- sort(unique(c(zero_tick_i, intensity_rel$Concentration_uM[intensity_rel$Treatment == "Virus + Compound"])))
}
dose_labels_i <- c("0", as.character(dose_breaks_i[-1]))

p_dr_int <- ggplot() +
  geom_errorbar(data = pts_i, aes(x = x_plot, ymin = ymin, ymax = ymax),
                width = 0.07, linewidth = 0.6, colour = "#2B5C8F") +
  geom_point(data = pts_i, aes(x = x_plot, y = y),
             shape = 21, size = 2.8, stroke = 0.5,
             fill = "#2B5C8F", colour = "black")

if (!is.null(fit_int)) {
  cf <- coef(fit_int)
  ec50_val  <- unname(cf["ec50"])
  halfway   <- unname(cf["bottom"] + (cf["top"] - cf["bottom"]) / 2)
  int_label <- paste0("EC50 = ", signif(ec50_val, 3), " ", DOSE_UNIT)
  p_dr_int  <- p_dr_int +
    geom_line(data = pred_i, aes(x = Concentration_uM, y = y),
              color = "#2B5C8F", linewidth = 1.0) +
    geom_segment(aes(x = ec50_val, xend = ec50_val, y = 0, yend = halfway),
                 linetype = "dashed", color = "firebrick3", linewidth = 0.6) +
    geom_segment(aes(x = zero_tick_i, xend = ec50_val, y = halfway, yend = halfway),
                 linetype = "dashed", color = "firebrick3", linewidth = 0.6) +
    annotate("label", x = ec50_val, y = 25, label = int_label,
             fill = "#2B5C8F", color = "white", fontface = "bold", size = 3, linewidth = 0)
}

p_dr_int <- p_dr_int +
  scale_x_log10(limits = c(zero_tick_i * 0.8, max_pos_i * 1.2), breaks = dose_breaks_i, labels = dose_labels_i, minor_breaks = NULL) +
  annotation_logticks(sides = "b", short = unit(1.2, "mm"), mid = unit(1.6, "mm"), long = unit(2.2, "mm")) +
  labs(
    title = paste(COMPOUND_NAME, "intensity dose–response"),
    subtitle = "Points show mean ± SE across replicate wells; curve fitted via 4PL regression",
    x     = paste0(COMPOUND_NAME, " (", DOSE_UNIT, ", log10 scale)"),
    y     = "Normalized NP Intensity (%)"
  ) +
  coord_cartesian(ylim = c(0, 110), clip = "off") +
  theme_bw(base_size = 9) +
  theme(
    plot.title        = element_text(face = "bold", size = 10, margin = margin(b = 4)),
    axis.title        = element_text(face = "bold", size = 9),
    axis.text         = element_text(size = 8),
    axis.line         = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks        = element_line(linewidth = 0.6, colour = "black"),
    axis.ticks.length = unit(2.2, "mm"),
    panel.grid.major  = element_blank(),
    panel.grid.minor  = element_blank(),
    legend.position   = "none",
    plot.margin       = margin(6, 10, 6, 8)
  )

print(p_dr_int)
ggsave(file.path(OUTPUT_DIR, "compound_dose_response_intensity.tiff"), p_dr_int,
       width = 5.5, height = 4.0, units = "in", dpi = PLOT_DPI)
```

# Results Summary

```{r results-summary}
tibble::tibble(
  Endpoint = c("EC50 — % infected (population GMM gate)",
               "EC50 — normalized NP intensity",
               "CC50 — host cell viability",
               "Selectivity Index (CC50 / EC50, % infection endpoint)"),
  Value = c(
    if (!is.null(fit_rel)) unname(coef(fit_rel)["ec50"]) else NA_real_,
    if (!is.null(fit_int)) unname(coef(fit_int)["ec50"]) else NA_real_,
    cc50_val,
    selectivity_index
  ),
  Unit = c(DOSE_UNIT, DOSE_UNIT, DOSE_UNIT, "ratio")
) %>%
  knitr::kable(caption = "Headline parameters", digits = 3)
```

Two co-primary antiviral endpoints are reported side-by-side, plus cytotoxicity:

1. **% infected** — fraction of cells above the population 2-GMM gate, fitted with binomial-weighted relative 4PL.
2. **Normalized NP intensity** — per-well mean log-intensity, normalized per plate to VO = 100% and Mock = 0%.
3. **CC50 / SI** — host-cell viability on the drug-only arm, fitted with the same 4PL, divided by the antiviral EC50.

### Caveats & Best Practices
- **Outlier Masking**: this template does **not** mask outliers automatically. Single-replicate non-monotonic outliers should be identified by eye and named in `WELLS_TO_EXCLUDE` / `DOSES_TO_EXCLUDE`, which removes them from the fits while the underlying cells remain in the QC panels.
- **Confluence Artifacts**: Over-confluent or washed-off wells must be excluded from analysis (`DOSES_TO_EXCLUDE`, `WELLS_TO_EXCLUDE`).
- **Selectivity**: an EC50 is only meaningful alongside a CC50. Quote the SI, and treat a CC50 that lies beyond the highest tested dose as a lower bound rather than a measurement.
- **Confidence intervals** on EC50/CC50 here are parametric (asymptotic, log-scale) approximations from the fit covariance, not resampling bootstraps. They are narrower than a true bootstrap when the fit is poorly identified.
