---
title: "Multi-Condition ANOVA — Welch's Test with Planned Contrasts"
author: "Alejandro Ponce-Flores"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width  = 7,
  fig.height = 4.5,
  dpi        = 300,
  fig.align  = "center",
  message    = FALSE,
  warning    = FALSE
)
```

::: {.callout-note}
**Analytical Purpose:** Performs multi-condition Welch's One-Way ANOVA with planned contrasts (reference group vs. all experimental conditions) and Holm-Bonferroni FWER correction. Designed for viral titres and NGS read count comparisons where heteroscedasticity is common.
:::

```{r user-config, message=FALSE, warning=FALSE}
## ── USER CONFIGURATION ──────────────────────────────────────────────────────
#
# CONDITION_FILES: Named character vector of CSV files to load.
#   Names become the "condition" label in all outputs.
#   Each CSV must be wide-format: one column per cell type / sample group,
#   with replicate runs stored as "CellType.1", "CellType.2" … suffix columns.
#
CONDITION_FILES <- c(
  "Condition_A" = "data/Condition_A.csv",
  "Condition_B" = "data/Condition_B.csv",
  "Condition_C" = "data/Condition_C.csv"
)

# FOCAL_CONDITION: Which condition to run the planned-contrast analysis on.
#   Must match one of the names() of CONDITION_FILES exactly.
FOCAL_CONDITION <- "Condition_A"

# REFERENCE_GROUP: The cell type / sample group used as the reference in
#   planned comparisons (i.e., all other groups are tested against this one).
REFERENCE_GROUP <- "CellType_A"

# LOG_TRANSFORM: Apply log10 before analysis? Almost always TRUE for viral
#   titres and NGS read-count data.
LOG_TRANSFORM <- TRUE

# ALPHA: Significance threshold for all tests.
ALPHA <- 0.05

## ────────────────────────────────────────────────────────────────────────────
```

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(car)
library(broom)
library(knitr)
```

# Load Data

```{r load-data}
clean_one <- function(path, condition_label) {
  read_csv(path, show_col_types = FALSE) %>%
    select(-matches("^Unnamed")) %>%                       # drop blank index cols
    pivot_longer(
      everything(),
      names_to  = "cell_type_raw",
      values_to = "value"
    ) %>%
    filter(!is.na(value)) %>%
    mutate(
      cell_type = str_remove(cell_type_raw, "\\.[0-9]+$") %>%   # strip ".1"
                  str_remove("\\.+$"),                           # strip trailing dots
      replicate = if_else(
        str_detect(cell_type_raw, "\\.[0-9]+$"),
        as.integer(str_extract(cell_type_raw, "\\d+$")) + 1L,
        1L
      ),
      condition   = condition_label,
      log_value   = if (LOG_TRANSFORM) log10(value) else value
    ) %>%
    select(condition, cell_type, replicate, value, log_value)
}

all_data <- imap_dfr(CONDITION_FILES, ~ clean_one(.x, .y))

all_data %>%
  head(15) %>%
  knitr::kable(
    caption     = "Preview of Processed Condition Data (First 15 Rows)",
    col.names   = c("Condition", "Cell Type", "Replicate", "Raw Value", "Log10 Value"),
    digits      = 3,
    format.args = list(big.mark = ",")
  )
```

# Assumption Checks

```{r diagnostics}
diagnostics <- all_data %>%
  group_by(condition) %>%
  summarise(
    shapiro_raw  = shapiro.test(value)$p.value,
    shapiro_log  = shapiro.test(log_value)$p.value,
    bartlett_log = bartlett.test(log_value ~ cell_type)$p.value,
    levene_log   = car::leveneTest(
      log_value ~ factor(cell_type),
      center = median
    )[["Pr(>F)"]][1],
    .groups = "drop"
  )

diagnostics %>%
  knitr::kable(
    caption     = "Normality (Shapiro-Wilk) and Homogeneity of Variance (Bartlett & Levene) Diagnostics",
    col.names   = c("Condition", "Shapiro (Raw) p", "Shapiro (Log10) p", "Bartlett (Log10) p", "Levene (Log10) p"),
    digits      = 4
  )
```

## Interpretation Guide

- **Raw data** from viral infections are typically **log-normally distributed** — expect low Shapiro-Wilk p-values on raw values.
- **Log₁₀ transform** symmetrises the distribution and stabilises variance. Check `shapiro_log` > `r ALPHA`.
- **Bartlett / Levene tests** flag heteroscedasticity. If these are significant for your focal condition, Welch's ANOVA (used below) is the correct choice.

### Why Welch's One-Way ANOVA on log₁₀ data?

1. **Handles heteroscedasticity** — group-specific variances are allowed.
2. **Tolerates moderate non-normality** — log transform already reduces most skew.
3. **More powerful than Kruskal-Wallis** at typical n = 3–6 per group.
4. Pairs with **Holm correction** for planned comparisons (controls family-wise error rate without being overly conservative).

# Focal Condition Analysis

```{r welch-anova}
focal <- all_data %>%
  filter(condition == FOCAL_CONDITION, !is.na(log_value)) %>%
  mutate(cell_type = factor(cell_type))

welch_anova <- oneway.test(log_value ~ cell_type,
                           data      = focal,
                           var.equal = FALSE)

welch_tbl <- tibble(
  Condition      = FOCAL_CONDITION,
  `F Statistic`  = unname(welch_anova$statistic),
  `Num df`       = unname(welch_anova$parameter["num df"]),
  `Denom df`     = unname(welch_anova$parameter["denom df"]),
  `p-value`      = unname(welch_anova$p.value)
)

welch_tbl %>%
  knitr::kable(
    caption = paste("Welch's One-Way ANOVA Test Result —", FOCAL_CONDITION),
    digits  = 4
  )
```

# Planned Contrasts: Reference Group vs All Others

```{r contrasts}
other_groups <- levels(focal$cell_type)[levels(focal$cell_type) != REFERENCE_GROUP]

contrasts <- map_df(other_groups, function(grp) {
  t_out <- t.test(
    focal$log_value[focal$cell_type == REFERENCE_GROUP],
    focal$log_value[focal$cell_type == grp],
    var.equal = FALSE
  )
  tibble(
    Comparison  = paste(REFERENCE_GROUP, "vs", grp),
    Log_Diff    = unname(t_out$estimate[1] - t_out$estimate[2]),
    p_raw       = t_out$p.value
  )
}) %>%
  mutate(
    p_adj       = p.adjust(p_raw, method = "holm"),
    fold_change = round(10^(-Log_Diff), 2),
    significant = ifelse(p_adj < ALPHA, "Yes", "No")
  ) %>%
  arrange(p_adj)

contrasts %>%
  knitr::kable(
    caption     = paste0("Planned Contrasts vs Reference Group (", REFERENCE_GROUP, ") with Holm-Bonferroni FWER Correction"),
    col.names   = c("Comparison", "Log10 Diff", "Raw p-value", "Holm Adj. p", "Fold Change", "Significant"),
    digits      = 4
  )
```

# Visualization

```{r summary-plot, fig.width=6.5, fig.height=4.5}
focal_summary <- focal %>%
  group_by(cell_type) %>%
  summarise(
    mean_val = mean(value, na.rm = TRUE),
    sd_val   = sd(value, na.rm = TRUE),
    n        = n(),
    se_val   = sd_val / sqrt(n),
    .groups  = "drop"
  )

p_focal <- ggplot(focal_summary, aes(x = cell_type, y = mean_val, fill = cell_type)) +
  geom_col(width = 0.65, color = "black", linewidth = 0.5, alpha = 0.85) +
  geom_errorbar(aes(ymin = pmax(0, mean_val - se_val), ymax = mean_val + se_val),
                width = 0.2, linewidth = 0.6) +
  geom_jitter(data = focal, aes(x = cell_type, y = value),
              width = 0.15, size = 2, alpha = 0.7, color = "grey20", inherit.aes = FALSE) +
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(
    title    = paste("Cell Type Comparison —", FOCAL_CONDITION),
    subtitle = paste("Reference:", REFERENCE_GROUP, "| Error bars represent Mean ± 1 SE"),
    x        = "Cell Type / Treatment Group",
    y        = ifelse(LOG_TRANSFORM, "Titer / Expression (linear scale)", "Response Value")
  ) +
  theme_bw(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 12),
    plot.subtitle    = element_text(size = 10, color = "grey30"),
    axis.title       = element_text(face = "bold"),
    axis.text.x      = element_text(angle = 25, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

print(p_focal)
```

## Summary Note

```{r summary-note}
cat(sprintf(
  "Focal condition: %s\nReference group: %s\nSignificant contrasts (Holm-adjusted p < %.2f): %d / %d\n",
  FOCAL_CONDITION,
  REFERENCE_GROUP,
  ALPHA,
  sum(contrasts$significant == "Yes"),
  nrow(contrasts)
))
```
