---
title: "07 · RNA-seq / DESeq2"
subtitle: "Differential expression, volcano plots, PCA, bubble plots, and triangle plots"
description: |
  Analyzes RNA-seq data using DESeq2. Loads a pre-computed DESeqDataSet, extracts
  results for user-defined contrasts, and produces publication-quality volcano plots,
  a PCA plot, bubble plots (viridis plasma, italic gene names), and directional triangle
  plots for user-specified gene sets.
categories: [transcriptomics, RNA-seq, DESeq2, differential-expression]
---

## Overview

| Item | Details |
|------|---------|
| **Input** | `dds_template.RData` — a saved `DESeqDataSet` processed with `DESeq()` |
| **Key packages** | `DESeq2`, `ggplot2`, `viridis`, `patchwork`, `svglite` |
| **Statistics** | DESeq2 Wald test · FDR-adjusted p-values (BH) |
| **Output** | Volcano plots · PCA · Bubble plots · Triangle plots · DEG CSVs |
| **Download** | [template.Rmd](template.Rmd) |

::: {.callout-tip}
**When to use this template:** You have a pre-processed DESeqDataSet and want to
visualize differential expression across multiple comparisons alongside curated gene sets.
:::

::: {.callout-note}
**Plot styles:** Volcano (red/blue) · Bubble (viridis plasma, italic y-axis) ·
Triangle (up/down triangles, viridis fill) · PCA (theme_classic, Arial 10pt)
:::

---

```{r}
#| label: simulate
#| include: false
source("data/simulate_data.R")
```

```{r child="template.Rmd"}
```
