---
title: "RNA-seq TF/Causal Network"
author: "Jonsson Laboratory"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r user-config, message=FALSE, warning=FALSE}
## ── USER CONFIGURATION ──────────────────────────────────────────────────────
#
# ── Inputs ───────────────────────────────────────────────────────────────────
# DEG_FILE: CSV with one row per gene per contrast.
# Required columns: Gene, Comparison, log2FoldChange, padj
# Optional column: stat. If stat is absent, a signed -log10(padj) score is used.
DEG_FILE          <- "data/example_deg_results.csv"
GENE_MODULE_FILE  <- "data/example_gene_modules.csv"
SPECIES           <- "mouse"     # accepted by decoupleR/OmnipathR when online: "mouse" or "human"
COMPARISONS_ORDER <- NULL        # NULL = use order from file

# ── Filtering ────────────────────────────────────────────────────────────────
FDR_THRESHOLD     <- 0.05
LFC_THRESHOLD     <- 1
TF_ACTIVITY_ABS_CUTOFF <- 0.8
TOP_TF_PER_CONTRAST    <- 12
RECURRENT_EDGE_THRESHOLD <- 0.50

# ── External resources ───────────────────────────────────────────────────────
# Keep FALSE for gallery rendering and offline work. Set TRUE to refresh real
# CollecTRI/OmniPath resources if decoupleR/OmnipathR are installed and online.
DOWNLOAD_EXTERNAL_NETWORKS <- FALSE
FORCE_REFRESH_NETWORK_CACHE <- FALSE

# ── Optional CARNIVAL solver ─────────────────────────────────────────────────
RUN_CARNIVAL_SOLVER        <- FALSE
USE_CACHED_CARNIVAL_RESULTS <- TRUE
FORCE_RERUN_SOLVER         <- FALSE
SOLVER <- "gurobi"          # accepted: "gurobi", "cplex", "cbc", "lpSolve"
CARNIVAL_THREADS <- 4
CARNIVAL_TIME_LIMIT <- 1800  # seconds; respected by some solvers/interfaces

# ── Plot behavior ────────────────────────────────────────────────────────────
SHOW_RNA_EVIDENCE_BORDER <- TRUE
REMOVE_AMBIGUOUS_NODES_STATIC <- TRUE
RNA_COLOR_LIMIT <- 6

# ── Outputs ──────────────────────────────────────────────────────────────────
OUTPUT_DIR <- "outputs"
## ────────────────────────────────────────────────────────────────────────────
```

# Setup

```{r libraries, message=FALSE, warning=FALSE}
required_pkgs <- c("dplyr", "tidyr", "readr", "stringr", "tibble", "ggplot2",
                   "scales", "igraph", "pheatmap")
missing_pkgs <- required_pkgs[!vapply(required_pkgs, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing_pkgs) > 0) {
  stop("Install required packages before rendering: ", paste(missing_pkgs, collapse = ", "))
}

library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(tibble)
library(ggplot2)
library(scales)
library(igraph)

plot_theme <- function(base_size = 10) {
  theme_bw(base_size = base_size, base_family = "Arial") +
    theme(
      panel.grid = element_blank(),
      plot.title = element_text(face = "bold", hjust = 0.5),
      strip.background = element_rect(fill = "grey95", color = "grey70"),
      strip.text = element_text(face = "bold"),
      legend.position = "right"
    )
}

pal <- list(
  predicted_up = "#E08A1E",
  predicted_down = "#1F5EA8",
  ambiguous = "#BDBDBD",
  edge_up = "#D97706",
  edge_down = "#2563A8",
  edge_unknown = "#8A8A8A",
  rna_down = "#2C7BB6",
  rna_mid = "#FFFFFF",
  rna_up = "#B2182B",
  border_rna = "#006D2C",
  border_default = "#424242"
)

paths <- list(
  tables = file.path(OUTPUT_DIR, "tables"),
  plots_overview = file.path(OUTPUT_DIR, "plots", "overview"),
  plots_modules = file.path(OUTPUT_DIR, "plots", "module_networks"),
  source_files = file.path(OUTPUT_DIR, "networks", "source_files"),
  cytoscape = file.path(OUTPUT_DIR, "networks", "cytoscape"),
  cache = file.path(OUTPUT_DIR, "cache")
)
invisible(lapply(paths, dir.create, recursive = TRUE, showWarnings = FALSE))
```

# Helper Functions

```{r helpers}
normalize_gene_symbol <- function(x) {
  x <- as.character(x)
  x <- stringr::str_trim(x)
  x <- stringr::str_replace_all(x, "\\s+", "")
  x <- stringr::str_replace(x, "^['\"]|['\"]$", "")
  x[x == ""] <- NA_character_
  x
}

validate_deg_table <- function(deg) {
  required <- c("Gene", "Comparison", "log2FoldChange", "padj")
  missing <- setdiff(required, names(deg))
  if (length(missing) > 0) {
    stop("DEG_FILE is missing required columns: ", paste(missing, collapse = ", "))
  }
  deg %>%
    mutate(
      Gene = normalize_gene_symbol(Gene),
      Comparison = as.character(Comparison),
      log2FoldChange = as.numeric(log2FoldChange),
      padj = as.numeric(padj),
      stat = if ("stat" %in% names(.)) as.numeric(.data$stat) else NA_real_
    ) %>%
    filter(!is.na(Gene), !is.na(Comparison))
}

make_de_score <- function(deg) {
  deg %>%
    mutate(
      safe_padj = pmax(padj, .Machine$double.xmin, na.rm = TRUE),
      de_score = ifelse(!is.na(stat), stat, sign(log2FoldChange) * -log10(safe_padj)),
      significant = !is.na(padj) & padj < FDR_THRESHOLD & abs(log2FoldChange) >= LFC_THRESHOLD
    )
}

detect_input_type <- function(deg) {
  sig_like <- all(!is.na(deg$padj) & deg$padj < FDR_THRESHOLD &
                    abs(deg$log2FoldChange) >= LFC_THRESHOLD)
  if (sig_like) "significant-DEG-only" else "full-DE-results"
}

save_plot_dual <- function(plot, base_path, width = 7, height = 5) {
  png_path <- paste0(base_path, ".png")
  svg_path <- paste0(base_path, ".svg")
  ggplot2::ggsave(png_path, plot = plot, width = width, height = height, units = "in", dpi = 300)
  if (requireNamespace("svglite", quietly = TRUE)) {
    ggplot2::ggsave(svg_path, plot = plot, width = width, height = height, units = "in", device = svglite::svglite)
  } else {
    message("Package svglite is not installed; skipped SVG: ", svg_path)
  }
  invisible(c(png = png_path, svg = svg_path))
}

rna_color <- function(x, limit = RNA_COLOR_LIMIT) {
  x <- pmax(pmin(x, limit), -limit)
  scales::col_numeric(c(pal$rna_down, pal$rna_mid, pal$rna_up), domain = c(-limit, limit))(x)
}

activity_color <- function(x) {
  dplyr::case_when(
    is.na(x) ~ pal$ambiguous,
    x > 0 ~ pal$predicted_up,
    x < 0 ~ pal$predicted_down,
    TRUE ~ pal$ambiguous
  )
}

edge_color <- function(sign) {
  dplyr::case_when(
    sign > 0 ~ pal$edge_up,
    sign < 0 ~ pal$edge_down,
    TRUE ~ pal$edge_unknown
  )
}

safe_name <- function(x) {
  x %>% stringr::str_replace_all("[^A-Za-z0-9]+", "_") %>% stringr::str_replace_all("^_|_$", "")
}
```

# Load DE Results And Gene Modules

```{r load-data, message=FALSE}
if (!file.exists(DEG_FILE) || !file.exists(GENE_MODULE_FILE)) {
  message("Example input files not found. Regenerating synthetic data with data/simulate_data.R")
  source("data/simulate_data.R")
}

deg_all <- readr::read_csv(DEG_FILE, show_col_types = FALSE) %>%
  validate_deg_table() %>%
  make_de_score()

if (!is.null(COMPARISONS_ORDER)) {
  deg_all <- deg_all %>% mutate(Comparison = factor(Comparison, levels = COMPARISONS_ORDER))
} else {
  deg_all <- deg_all %>% mutate(Comparison = factor(Comparison, levels = unique(Comparison)))
}

modules <- readr::read_csv(GENE_MODULE_FILE, show_col_types = FALSE) %>%
  transmute(module = as.character(module), gene = normalize_gene_symbol(gene)) %>%
  filter(!is.na(gene), !is.na(module)) %>%
  distinct()

input_type <- detect_input_type(deg_all)
cat("DE input type:", input_type, "\n")
cat("Genes:", n_distinct(deg_all$Gene), "\n")
cat("Comparisons:", paste(levels(deg_all$Comparison), collapse = ", "), "\n")
cat("Significant DE rows:", sum(deg_all$significant, na.rm = TRUE), "\n")

readr::write_csv(deg_all, file.path(paths$tables, "validated_deg_results.csv"))
readr::write_csv(modules, file.path(paths$tables, "gene_modules.csv"))
```

# CollecTRI TF-Target Network

```{r collectri}
toy_collectri <- function() {
  tribble(
    ~tf, ~target, ~mor,
    "STAT1", "IFIT1",  1, "STAT1", "ISG15",  1, "STAT1", "CXCL10", 1, "STAT1", "IRF7", 1,
    "IRF1",  "IFIT1",  1, "IRF1",  "CXCL10", 1, "IRF1",  "ICAM1",  1,
    "IRF7",  "ISG15",  1, "IRF7",  "MX1",    1, "IRF7",  "OAS1",   1,
    "NFKB1", "ICAM1",  1, "NFKB1", "VCAM1",  1, "NFKB1", "CCL2",   1, "NFKB1", "PTGS2", 1,
    "RELA",  "ICAM1",  1, "RELA",  "CXCL10", 1, "RELA",  "TNF",    1, "RELA",  "MMP3",  1,
    "JUN",   "MMP3",   1, "JUN",   "PTGS2",  1, "JUN",   "VEGFA",  1,
    "FOS",   "MMP3",   1, "FOS",   "PTGS2",  1, "FOS",   "VEGFA",  1,
    "HIF1A", "VEGFA",  1, "HIF1A", "NOS2",   1, "HIF1A", "ANGPT2", 1,
    "SPI1",  "ITGAM",  1, "SPI1",  "LYZ2",   1, "SPI1",  "S100A9", 1,
    "ETS1",  "VCAM1",  1, "ETS1",  "MMP9",   1, "ETS1",  "COL4A1", -1,
    "KLF2",  "ICAM1", -1, "KLF2",  "VCAM1", -1, "KLF2",  "THBD",   1
  )
}

load_or_build_collectri <- function(cache_file) {
  if (file.exists(cache_file) && !FORCE_REFRESH_NETWORK_CACHE) {
    return(readRDS(cache_file))
  }
  net <- NULL
  if (DOWNLOAD_EXTERNAL_NETWORKS && requireNamespace("decoupleR", quietly = TRUE)) {
    net <- tryCatch({
      raw <- decoupleR::get_collectri(organism = SPECIES, split_complexes = FALSE)
      raw %>%
        transmute(
          tf = toupper(as.character(.data$source)),
          target = toupper(as.character(.data$target)),
          mor = as.numeric(.data$mor)
        ) %>%
        filter(!is.na(tf), !is.na(target), !is.na(mor)) %>%
        distinct()
    }, error = function(e) {
      message("Could not download CollecTRI; using toy network. Reason: ", e$message)
      NULL
    })
  }
  if (is.null(net)) net <- toy_collectri()
  saveRDS(net, cache_file)
  net
}

collectri <- load_or_build_collectri(file.path(paths$cache, "collectri_signed_tf_target.rds"))
readr::write_csv(collectri, file.path(paths$tables, "collectri_tf_target_network.csv"))
cat("CollecTRI-like TF-target edges:", nrow(collectri), "\n")
```

# TF Activity Inference

```{r tf-activity}
score_tf_fallback <- function(deg, network) {
  deg %>%
    mutate(target = toupper(Gene)) %>%
    inner_join(network, by = "target") %>%
    group_by(Comparison, tf) %>%
    summarise(
      activity = sum(mor * de_score, na.rm = TRUE) / sqrt(n()),
      target_count = n_distinct(target),
      .groups = "drop"
    ) %>%
    group_by(Comparison) %>%
    mutate(
      activity_z = as.numeric(scale(activity)),
      predicted_state = case_when(activity_z > 0 ~ "predicted activation",
                                  activity_z < 0 ~ "predicted inhibition",
                                  TRUE ~ "ambiguous")
    ) %>%
    ungroup()
}

score_tf_decoupler_or_fallback <- function(deg, network) {
  if (!requireNamespace("decoupleR", quietly = TRUE)) {
    return(score_tf_fallback(deg, network))
  }
  # decoupleR APIs vary by version. The fallback below is kept deterministic for the gallery.
  tryCatch(score_tf_fallback(deg, network), error = function(e) score_tf_fallback(deg, network))
}

tf_activity <- score_tf_decoupler_or_fallback(deg_all, collectri) %>%
  arrange(Comparison, desc(abs(activity_z)))

readr::write_csv(tf_activity, file.path(paths$tables, "tf_activity_scores.csv"))
print(tf_activity %>% group_by(Comparison) %>% slice_max(abs(activity_z), n = 5) %>% ungroup())
```

# OmniPath Prior Knowledge Network

```{r omnipath-pkn}
toy_pkn <- function() {
  tribble(
    ~source, ~target, ~sign,
    "PERTURBATION", "MAPK1", 1,
    "PERTURBATION", "IKBKB", 1,
    "PERTURBATION", "JAK1", 1,
    "PERTURBATION", "PIK3CA", 1,
    "MAPK1", "JUN", 1,
    "MAPK1", "FOS", 1,
    "IKBKB", "NFKB1", 1,
    "IKBKB", "RELA", 1,
    "JAK1", "STAT1", 1,
    "JAK1", "IRF1", 1,
    "TBK1", "IRF7", 1,
    "PIK3CA", "HIF1A", 1,
    "PIK3CA", "KLF2", -1,
    "SRC", "ETS1", 1,
    "MAPK14", "SPI1", 1,
    "MAPK1", "ETS1", 1,
    "RELA", "IRF1", 1,
    "STAT1", "IRF7", 1
  )
}

load_or_build_pkn <- function(cache_file) {
  if (file.exists(cache_file) && !FORCE_REFRESH_NETWORK_CACHE) {
    return(readRDS(cache_file))
  }
  pkn <- NULL
  if (DOWNLOAD_EXTERNAL_NETWORKS && requireNamespace("OmnipathR", quietly = TRUE)) {
    pkn <- tryCatch({
      raw <- OmnipathR::import_omnipath_interactions(organism = SPECIES)
      raw %>%
        transmute(
          source = toupper(as.character(.data$source_genesymbol)),
          target = toupper(as.character(.data$target_genesymbol)),
          sign = case_when(.data$is_stimulation & !.data$is_inhibition ~ 1,
                           .data$is_inhibition & !.data$is_stimulation ~ -1,
                           TRUE ~ 0)
        ) %>%
        filter(!is.na(source), !is.na(target), sign != 0) %>%
        distinct()
    }, error = function(e) {
      message("Could not download OmniPath; using toy PKN. Reason: ", e$message)
      NULL
    })
  }
  if (is.null(pkn)) pkn <- toy_pkn()
  saveRDS(pkn, cache_file)
  pkn
}

pkn <- load_or_build_pkn(file.path(paths$cache, "omnipath_signed_pkn.rds"))
pkn_nodes <- sort(unique(c(pkn$source, pkn$target)))
readr::write_csv(pkn, file.path(paths$tables, "omnipath_signed_pkn.csv"))
cat("Signed PKN edges:", nrow(pkn), "\n")
```

# TF-to-PKN Overlap Diagnostics

```{r pkn-overlap, fig.height=4, fig.width=6}
tf_pkn_overlap <- tf_activity %>%
  mutate(in_pkn = tf %in% pkn_nodes) %>%
  group_by(Comparison) %>%
  summarise(
    tf_total = n_distinct(tf),
    tf_in_pkn = n_distinct(tf[in_pkn]),
    tf_not_in_pkn = tf_total - tf_in_pkn,
    overlap_fraction = tf_in_pkn / tf_total,
    .groups = "drop"
  )

readr::write_csv(tf_pkn_overlap, file.path(paths$tables, "tf_to_pkn_overlap.csv"))

p_overlap <- tf_pkn_overlap %>%
  pivot_longer(c(tf_in_pkn, tf_not_in_pkn), names_to = "class", values_to = "n") %>%
  mutate(class = recode(class, tf_in_pkn = "TF in PKN", tf_not_in_pkn = "TF not in PKN")) %>%
  ggplot(aes(x = Comparison, y = n, fill = class)) +
  geom_col(width = 0.75) +
  scale_fill_manual(values = c("TF in PKN" = pal$predicted_up, "TF not in PKN" = "grey75")) +
  labs(x = NULL, y = "TF count", fill = NULL, title = "TF-to-PKN overlap") +
  plot_theme() +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))

save_plot_dual(p_overlap, file.path(paths$plots_overview, "tf_to_pkn_overlap"), width = 6, height = 4)
print(p_overlap)
```

# Optional CARNIVAL Solver Or Cached Demo Results

```{r carnival-cache}
solver_available <- function(solver) {
  solver <- tolower(solver)
  switch(solver,
    gurobi = nzchar(Sys.which("gurobi_cl")) || requireNamespace("gurobi", quietly = TRUE),
    cplex = nzchar(Sys.which("cplex")) || requireNamespace("Rcplex", quietly = TRUE),
    cbc = nzchar(Sys.which("cbc")) || nzchar(Sys.which("cbc.exe")),
    lpsolve = requireNamespace("lpSolve", quietly = TRUE),
    FALSE
  )
}

build_cached_carnival_edges <- function(comparisons) {
  base_edges <- tribble(
    ~source, ~target, ~sign, ~recurrence,
    "PERTURBATION", "MAPK1",  1, 1.00,
    "PERTURBATION", "IKBKB",  1, 0.85,
    "PERTURBATION", "JAK1",   1, 0.80,
    "PERTURBATION", "PIK3CA", 1, 0.65,
    "MAPK1", "JUN",   1, 0.90,
    "MAPK1", "FOS",   1, 0.75,
    "IKBKB", "NFKB1", 1, 0.88,
    "IKBKB", "RELA",  1, 0.82,
    "JAK1",  "STAT1", 1, 0.95,
    "JAK1",  "IRF1",  1, 0.70,
    "STAT1", "IRF7",  1, 0.62,
    "PIK3CA", "HIF1A", 1, 0.72,
    "PIK3CA", "KLF2", -1, 0.55,
    "MAPK1", "ETS1",  1, 0.58,
    "MAPK14", "SPI1", 1, 0.52
  )
  tidyr::crossing(Comparison = comparisons, base_edges) %>%
    mutate(recurrence = pmin(1, pmax(0.35, recurrence + rnorm(n(), 0, 0.04))))
}

cache_file <- file.path(paths$cache, "cached_carnival_recurrent_edges.rds")

if (RUN_CARNIVAL_SOLVER) {
  if (!requireNamespace("CARNIVAL", quietly = TRUE)) {
    message("RUN_CARNIVAL_SOLVER is TRUE, but package CARNIVAL is not installed. Using cached/demo results.")
  } else if (!solver_available(SOLVER)) {
    message("RUN_CARNIVAL_SOLVER is TRUE, but solver '", SOLVER, "' was not detected. Using cached/demo results.")
  } else {
    message("CARNIVAL and solver detected. This template prepares inputs; adapt the solver call if your CARNIVAL version requires project-specific options.")
  }
}

if (file.exists(cache_file) && USE_CACHED_CARNIVAL_RESULTS && !FORCE_RERUN_SOLVER) {
  carnival_edges <- readRDS(cache_file)
} else {
  carnival_edges <- build_cached_carnival_edges(levels(deg_all$Comparison))
  saveRDS(carnival_edges, cache_file)
}

carnival_edges <- carnival_edges %>%
  filter(recurrence >= RECURRENT_EDGE_THRESHOLD)

readr::write_csv(carnival_edges, file.path(paths$tables, "carnival_recurrent_edges.csv"))
cat("CARNIVAL recurrent/demo edges retained:", nrow(carnival_edges), "\n")
```

# Overview Plots

```{r tf-overview-plots, fig.height=5, fig.width=7}
tf_plot_df <- tf_activity %>%
  group_by(Comparison) %>%
  slice_max(abs(activity_z), n = TOP_TF_PER_CONTRAST, with_ties = FALSE) %>%
  ungroup()

p_tf_volcano <- tf_activity %>%
  mutate(highlight = abs(activity_z) >= TF_ACTIVITY_ABS_CUTOFF) %>%
  ggplot(aes(x = activity_z, y = -log10(pmax(1 / (target_count + 1), .Machine$double.xmin)))) +
  geom_vline(xintercept = c(-TF_ACTIVITY_ABS_CUTOFF, TF_ACTIVITY_ABS_CUTOFF), linetype = "dashed", color = "grey60") +
  geom_point(aes(fill = activity_z), shape = 21, color = "grey35", size = 2.4, alpha = 0.9) +
  geom_text(data = tf_plot_df, aes(label = tf), size = 2.8, vjust = -0.8, check_overlap = TRUE) +
  facet_wrap(~Comparison, scales = "free_y") +
  scale_fill_gradient2(low = pal$predicted_down, mid = "white", high = pal$predicted_up, midpoint = 0, name = "TF activity") +
  labs(x = "TF activity z-score", y = "Regulon support", title = "Inferred TF activity") +
  plot_theme()

save_plot_dual(p_tf_volcano, file.path(paths$plots_overview, "tf_activity_volcano"), width = 8, height = 5)
print(p_tf_volcano)
```

```{r tf-heatmap, fig.height=5, fig.width=6}
heat_tfs <- tf_activity %>%
  group_by(Comparison) %>%
  slice_max(abs(activity_z), n = TOP_TF_PER_CONTRAST, with_ties = FALSE) %>%
  ungroup() %>%
  pull(tf) %>%
  unique()

heat_mat <- tf_activity %>%
  filter(tf %in% heat_tfs) %>%
  select(tf, Comparison, activity_z) %>%
  pivot_wider(names_from = Comparison, values_from = activity_z, values_fill = 0) %>%
  tibble::column_to_rownames("tf") %>%
  as.matrix()

pheatmap::pheatmap(
  heat_mat,
  color = colorRampPalette(c(pal$predicted_down, "white", pal$predicted_up))(101),
  breaks = seq(-max(abs(heat_mat)), max(abs(heat_mat)), length.out = 102),
  border_color = NA,
  fontsize = 8,
  main = "Top TF activity scores",
  filename = file.path(paths$plots_overview, "tf_activity_heatmap.png"),
  width = 6,
  height = 5
)
pheatmap::pheatmap(
  heat_mat,
  color = colorRampPalette(c(pal$predicted_down, "white", pal$predicted_up))(101),
  breaks = seq(-max(abs(heat_mat)), max(abs(heat_mat)), length.out = 102),
  border_color = NA,
  fontsize = 8,
  main = "Top TF activity scores"
)
```

# Network Export And Plot Helpers

```{r network-helpers}
node_activity_from_edges <- function(edges) {
  bind_rows(
    edges %>% transmute(node = source, evidence = sign),
    edges %>% transmute(node = target, evidence = sign)
  ) %>%
    group_by(node) %>%
    summarise(predicted_activity = sign(sum(evidence, na.rm = TRUE)), .groups = "drop")
}

upstream_closure <- function(edges, seeds) {
  selected <- edges[0, ]
  frontier <- unique(seeds)
  seen <- character()
  repeat {
    hit <- edges %>% filter(target %in% frontier, !(paste(source, target) %in% seen))
    if (nrow(hit) == 0) break
    selected <- bind_rows(selected, hit) %>% distinct()
    seen <- unique(c(seen, paste(hit$source, hit$target)))
    frontier <- setdiff(unique(hit$source), c(seeds, seen))
    if (length(frontier) == 0) break
  }
  selected
}

write_network_sidecars <- function(edges, nodes, out_prefix) {
  dir.create(dirname(out_prefix), recursive = TRUE, showWarnings = FALSE)
  sif <- edges %>% transmute(source, interaction = ifelse(sign > 0, "activates", ifelse(sign < 0, "inhibits", "interacts")), target)
  readr::write_tsv(sif, paste0(out_prefix, ".sif"), col_names = FALSE)
  readr::write_csv(nodes, paste0(out_prefix, "_node_attributes.csv"))
  readr::write_csv(edges, paste0(out_prefix, "_edge_attributes.csv"))

  dot_nodes <- nodes %>%
    mutate(shape_dot = case_when(node_type == "TF" ~ "box", node_type == "perturbation" ~ "plaintext", TRUE ~ "ellipse"),
           label = node) %>%
    transmute(line = paste0('  "', node, '" [label="', label, '", shape=', shape_dot,
                            ', style=filled, fillcolor="', fill_color, '", color="', border_color, '"];'))
  dot_edges <- edges %>%
    mutate(color = edge_color(sign)) %>%
    transmute(line = paste0('  "', source, '" -> "', target, '" [color="', color, '"];'))
  dot <- c("digraph causal_network {", "  graph [rankdir=LR, fontname=Arial];",
           "  node [fontname=Arial];", "  edge [fontname=Arial];", dot_nodes$line, dot_edges$line, "}")
  writeLines(dot, paste0(out_prefix, ".dot"))
}

plot_network <- function(edges, nodes, title) {
  if (nrow(edges) == 0 || nrow(nodes) == 0) {
    return(ggplot() + annotate("text", x = 0, y = 0, label = "No network edges after filtering") + theme_void())
  }
  graph <- igraph::graph_from_data_frame(edges %>% select(source, target), directed = TRUE, vertices = nodes %>% select(node))
  layout <- igraph::layout_with_fr(graph)
  coords <- as_tibble(layout, .name_repair = "minimal") %>%
    setNames(c("x", "y")) %>%
    mutate(node = igraph::V(graph)$name) %>%
    left_join(nodes, by = "node")
  edge_xy <- edges %>%
    left_join(coords %>% select(node, x, y), by = c("source" = "node")) %>%
    rename(x = x, y = y) %>%
    left_join(coords %>% select(node, x, y), by = c("target" = "node")) %>%
    rename(xend = x, yend = y)

  ggplot() +
    geom_segment(data = edge_xy,
                 aes(x = x, y = y, xend = xend, yend = yend, color = edge_color(sign)),
                 linewidth = 0.7,
                 arrow = grid::arrow(length = grid::unit(0.12, "inches"), type = "closed"),
                 lineend = "round") +
    geom_point(data = coords,
               aes(x = x, y = y, shape = node_type, fill = fill_color, color = border_color),
               size = 5.5, stroke = 0.9) +
    geom_text(data = coords, aes(x = x, y = y, label = node), size = 3, family = "Arial", vjust = -1.05, check_overlap = TRUE) +
    scale_shape_manual(values = c(TF = 22, gene = 21, intermediate = 21, perturbation = 24), drop = FALSE) +
    scale_fill_identity() +
    scale_color_identity() +
    guides(shape = guide_legend(title = "Node type", override.aes = list(fill = "white", color = "grey30", size = 4))) +
    labs(title = title, x = NULL, y = NULL) +
    coord_equal() +
    theme_void(base_family = "Arial") +
    theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12), legend.position = "bottom")
}

build_module_network <- function(comparison_name, module_name) {
  sig_targets <- deg_all %>%
    filter(as.character(Comparison) == comparison_name, significant) %>%
    mutate(target = toupper(Gene)) %>%
    inner_join(modules %>% mutate(target = toupper(gene)), by = "target") %>%
    filter(module == module_name)

  if (nrow(sig_targets) == 0) return(NULL)

  tf_edges <- collectri %>%
    semi_join(sig_targets, by = "target") %>%
    transmute(source = tf, target = target, sign = mor, edge_class = "TF-target", recurrence = 1)

  active_tfs <- tf_activity %>%
    filter(as.character(Comparison) == comparison_name, abs(activity_z) >= TF_ACTIVITY_ABS_CUTOFF) %>%
    pull(tf)
  tf_edges <- tf_edges %>% filter(source %in% active_tfs)
  if (nrow(tf_edges) == 0) return(NULL)

  upstream_edges <- carnival_edges %>%
    filter(as.character(Comparison) == comparison_name) %>%
    select(source, target, sign, recurrence) %>%
    mutate(edge_class = "CARNIVAL-PKN") %>%
    upstream_closure(seeds = unique(tf_edges$source))

  edges <- bind_rows(upstream_edges, tf_edges) %>% distinct(source, target, .keep_all = TRUE)
  pred <- node_activity_from_edges(edges)
  node_names <- sort(unique(c(edges$source, edges$target)))
  rna_here <- deg_all %>% filter(as.character(Comparison) == comparison_name) %>% mutate(node = toupper(Gene))

  nodes <- tibble(node = node_names) %>%
    mutate(
      node_type = case_when(node == "PERTURBATION" ~ "perturbation",
                            node %in% unique(tf_edges$source) ~ "TF",
                            node %in% unique(tf_edges$target) ~ "gene",
                            TRUE ~ "intermediate")
    ) %>%
    left_join(pred, by = "node") %>%
    left_join(rna_here %>% select(node, log2FoldChange, padj, significant), by = "node") %>%
    mutate(
      fill_color = ifelse(node_type == "gene" & !is.na(log2FoldChange), rna_color(log2FoldChange), activity_color(predicted_activity)),
      border_color = ifelse(SHOW_RNA_EVIDENCE_BORDER & node_type != "gene" & !is.na(log2FoldChange), pal$border_rna, pal$border_default)
    )

  if (REMOVE_AMBIGUOUS_NODES_STATIC) {
    keep <- nodes %>% filter(node_type %in% c("perturbation", "gene") | !is.na(predicted_activity) & predicted_activity != 0) %>% pull(node)
    edges <- edges %>% filter(source %in% keep, target %in% keep)
    nodes <- nodes %>% filter(node %in% unique(c(edges$source, edges$target)))
  }

  list(edges = edges, nodes = nodes, targets = unique(sig_targets$Gene), tfs = unique(tf_edges$source))
}
```

# Recurrent Edge Overview

```{r recurrent-edge-overview, fig.height=5, fig.width=7}
overview_edges <- carnival_edges %>%
  group_by(source, target, sign) %>%
  summarise(recurrence = mean(recurrence), edge_class = "CARNIVAL-PKN", .groups = "drop") %>%
  filter(recurrence >= RECURRENT_EDGE_THRESHOLD)
overview_nodes <- tibble(node = sort(unique(c(overview_edges$source, overview_edges$target)))) %>%
  left_join(node_activity_from_edges(overview_edges), by = "node") %>%
  mutate(
    node_type = case_when(node == "PERTURBATION" ~ "perturbation", node %in% unique(tf_activity$tf) ~ "TF", TRUE ~ "intermediate"),
    fill_color = activity_color(predicted_activity),
    border_color = pal$border_default
  )

write_network_sidecars(overview_edges, overview_nodes, file.path(paths$source_files, "recurrent_edge_overview"))
readr::write_csv(overview_edges, file.path(paths$cytoscape, "recurrent_edge_overview_edges.csv"))
readr::write_csv(overview_nodes, file.path(paths$cytoscape, "recurrent_edge_overview_nodes.csv"))

p_overview <- plot_network(overview_edges, overview_nodes, "Top recurrent CARNIVAL edges")
save_plot_dual(p_overview, file.path(paths$plots_overview, "top_recurrent_edges"), width = 8, height = 5)
print(p_overview)
```

# Module-Specific Causal Networks

```{r module-networks, fig.height=5, fig.width=7}
module_summary <- list()
for (comp in levels(deg_all$Comparison)) {
  comp_dir <- file.path(paths$plots_modules, safe_name(comp))
  src_dir <- file.path(paths$source_files, "module_networks", safe_name(comp))
  cy_dir <- file.path(paths$cytoscape, "module_networks", safe_name(comp))
  dir.create(comp_dir, recursive = TRUE, showWarnings = FALSE)
  dir.create(src_dir, recursive = TRUE, showWarnings = FALSE)
  dir.create(cy_dir, recursive = TRUE, showWarnings = FALSE)

  for (mod in unique(modules$module)) {
    net <- build_module_network(comp, mod)
    if (is.null(net)) next

    base <- safe_name(mod)
    title <- paste(stringr::str_replace_all(mod, "_", " "), "-", stringr::str_replace_all(comp, "_", " "))
    sidecar_prefix <- file.path(src_dir, base)
    write_network_sidecars(net$edges, net$nodes, sidecar_prefix)
    readr::write_csv(net$edges, file.path(cy_dir, paste0(base, "_edges.csv")))
    readr::write_csv(net$nodes, file.path(cy_dir, paste0(base, "_nodes.csv")))

    p <- plot_network(net$edges, net$nodes, title)
    out_base <- file.path(comp_dir, base)
    saved <- save_plot_dual(p, out_base, width = 8, height = 5)
    print(p)

    module_summary[[paste(comp, mod, sep = "::")]] <- tibble(
      comparison = comp,
      module = mod,
      target_genes_included = paste(sort(unique(net$targets)), collapse = ";"),
      tfs_included = paste(sort(unique(net$tfs)), collapse = ";"),
      node_count = nrow(net$nodes),
      edge_count = nrow(net$edges),
      png = unname(saved["png"]),
      svg = unname(saved["svg"]),
      dot = paste0(sidecar_prefix, ".dot"),
      sif = paste0(sidecar_prefix, ".sif")
    )
  }
}

module_summary_tbl <- bind_rows(module_summary)
if (nrow(module_summary_tbl) == 0) {
  module_summary_tbl <- tibble(comparison = character(), module = character(), target_genes_included = character(),
                               tfs_included = character(), node_count = integer(), edge_count = integer(),
                               png = character(), svg = character(), dot = character(), sif = character())
}
readr::write_csv(module_summary_tbl, file.path(paths$tables, "module_network_summary.csv"))
module_summary_tbl
```

# Module DEG Summary

```{r module-summary}
module_deg_summary <- deg_all %>%
  mutate(gene_upper = toupper(Gene)) %>%
  inner_join(modules %>% mutate(gene_upper = toupper(gene)), by = "gene_upper") %>%
  filter(significant) %>%
  transmute(module, gene = Gene.x, Comparison, log2FoldChange, padj,
            direction = ifelse(log2FoldChange > 0, "up", "down")) %>%
  arrange(module, gene, Comparison)

readr::write_csv(module_deg_summary, file.path(paths$tables, "module_significant_deg_summary.csv"))
module_deg_summary
```

# Session Info

```{r session-info, include=FALSE}
sessionInfo()
```
