---
title: "MagPix / Luminex Multiplex Cytokine Analysis"
author: "Alejandro Ponce-Flores"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r user-config, message=FALSE, warning=FALSE}
## ── USER CONFIGURATION ──────────────────────────────────────────────────────
#
# DATA_FILE: Path to your MagPix/Luminex export CSV.
#   Required columns (in addition to analyte columns):
#   - "Location"     : well identifier (e.g., "A1")
#   - "Sample"       : sample name — used to derive Factor1 and Factor2
#   - "Original_ID"  : original sample ID (can be same as Sample)
#   - "Total Events" : bead event count (for QC)
#
DATA_FILE <- "data/luminex_data.csv"

# NON_ANALYTE_COLS: Columns that are NOT cytokine measurements.
#   All other columns will be treated as analytes.
NON_ANALYTE_COLS <- c("Location", "Sample", "Original_ID", "Total Events")

# FACTOR1_LEVELS / FACTOR2_LEVELS: Level labels for the two experimental
#   factors. Each label is matched as a *literal substring* (not a regex)
#   against the Sample column, so labels containing "+", "(" or "." are safe.
#   Element [1] is the reference / control level of that factor.
#
#   Example: Sample = "Disease_Drug_1"
#     "Disease" is found in Sample → Factor1 = "Disease"
#     "Drug"    is found in Sample → Factor2 = "Drug"
#
#   A sample matching *no* level of a factor is assigned NA and dropped
#   (with a count reported) rather than being folded into the control.
#
FACTOR1_LEVELS <- c("PBS", "Disease")     # [1] = reference / control level
FACTOR2_LEVELS <- c("Vehicle", "Drug")    # [1] = reference / control level

# NON_SAMPLE_PATTERNS: Literal substrings identifying non-biological wells that
#   Luminex exports routinely carry (standards, blanks, background). Rows whose
#   Sample matches any of these are removed before modelling.
NON_SAMPLE_PATTERNS <- c("Standard", "Blank", "Background", "Control Bead")

# EXCLUDE_OUTLIERS: Tukey 3xIQR outliers are always *flagged* and tabulated.
#   Set to TRUE to also drop them from the models; FALSE keeps them (default —
#   deleting Luminex wells on a distributional rule alone is rarely justified).
EXCLUDE_OUTLIERS <- FALSE

# OKABE_ITO: Colorblind-safe palette keyed by group label.
#   Names follow the "<Factor1>+<Factor2>" convention. Groups not listed here
#   are filled automatically from the Okabe-Ito sequence, so a design larger
#   than 2x2 still plots.
OKABE_ITO <- c(
  "PBS+Vehicle"      = "#0072B2",   # Blue
  "PBS+Drug"         = "#E69F00",   # Orange
  "Disease+Vehicle"  = "#009E73",   # Bluish green
  "Disease+Drug"     = "#D55E00"    # Vermilion
)

# CONTRAST_ORDER / CONTRAST_LABELS: Define which emmeans contrasts appear in
#   the heatmap (order matters for the y-axis) and their display labels.
#   Adjust if your Factor levels produce different contrast names.
CONTRAST_ORDER <- c(
  paste0(FACTOR2_LEVELS[2], " - ", FACTOR2_LEVELS[1], ", ", FACTOR1_LEVELS[1]),
  paste0(FACTOR2_LEVELS[2], " - ", FACTOR2_LEVELS[1], ", ", FACTOR1_LEVELS[2]),
  paste0(FACTOR1_LEVELS[2], " - ", FACTOR1_LEVELS[1], ", ", FACTOR2_LEVELS[1]),
  paste0(FACTOR1_LEVELS[2], " - ", FACTOR1_LEVELS[1], ", ", FACTOR2_LEVELS[2])
)
CONTRAST_LABELS <- c(
  paste0("Drug (", FACTOR1_LEVELS[1], ")"),
  paste0("Drug (", FACTOR1_LEVELS[2], ")"),
  paste0(FACTOR1_LEVELS[2], " (", FACTOR2_LEVELS[1], ")"),
  paste0(FACTOR1_LEVELS[2], " (", FACTOR2_LEVELS[2], ")")
)

# OUTPUT_DIR: Folder where per-analyte plots are saved.
OUTPUT_DIR <- "Plots"

## ────────────────────────────────────────────────────────────────────────────
```

# MagPix / Luminex Best Practices

- **Bead counts:** ≥35 per analyte (≥50 ideal). Re-export with bead counts if missing.
- **Left-censoring ("< LoD"):** Do not treat as zero. Replace with **LoD/√2**.
- **Right-censoring ("> ULoQ"):** Equally important and usually ignored. `parse_number("> 10000")`
  returns `10000`, which enters the model as an exact measurement. Here such values are substituted
  at the **ULoQ itself** and flagged; the resulting effect estimates are *attenuated* (conservative),
  which is stated wherever the censoring rate is non-trivial.
- **QC/outliers:** Flag wells using **Tukey 3×IQR** on the log10 scale. Flagging is not the same as
  excluding — this template flags and tabulates them, and only drops them if `EXCLUDE_OUTLIERS`
  is set to `TRUE`.
- **Non-sample wells:** Standards, blanks and background rows must be removed *before* modelling,
  not silently folded into the control group.
- **Design:** factorial (2×2 by default) — model on **log10(concentration)**, test main effects and
  interaction. Level counts are derived from the config, so a 2×3 design works unchanged.
- **Multiple analytes:** Correct FDR **within each model term**, across analytes (BH). The intercept
  is not a hypothesis of interest and is excluded from the correction family. Report fold-changes
  as 10^β.

# Setup

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr)
library(broom)
library(emmeans)
library(ggpubr)
library(ggbeeswarm)
library(scales)
library(forcats)
```

```{r make-output-dirs, include=FALSE}
# Every figure produced by this template is written to OUTPUT_DIR/Heatmaps.
dir.create(file.path(OUTPUT_DIR, "Heatmaps"), recursive = TRUE, showWarnings = FALSE)
```

# Load Data

```{r load-data, message=FALSE}
raw <- read_csv(DATA_FILE, show_col_types = FALSE)

analyte_cols <- setdiff(names(raw), NON_ANALYTE_COLS)
raw <- raw %>% mutate(across(all_of(analyte_cols), as.character))

n_wells_in <- nrow(raw)

# ── 1. Remove non-biological wells (standards, blanks, background) ────────────
# Matched as literal substrings, case-insensitively, so a pattern containing
# regex metacharacters is safe.
if (length(NON_SAMPLE_PATTERNS) > 0) {
  is_non_sample <- Reduce(
    `|`,
    lapply(NON_SAMPLE_PATTERNS,
           function(p) str_detect(raw$Sample, fixed(p, ignore_case = TRUE)))
  )
} else {
  is_non_sample <- rep(FALSE, nrow(raw))
}
n_non_sample <- sum(is_non_sample)
raw <- raw[!is_non_sample, , drop = FALSE]

# ── 2. Long format, censoring handling, factor assignment ────────────────────
# Left-censored  ("< x") → substitute LoD/sqrt(2).
# Right-censored ("> x") → substitute the ULoQ itself; the estimate is then a
#   conservative lower bound rather than a point measurement.
# A sample matching no level of a factor gets NA (never the control level).
# For a factor with more than two levels, add one `str_detect(...) ~ level`
# line per extra level, most specific first.
dat_long <- raw %>%
  pivot_longer(all_of(analyte_cols), names_to = "Analyte", values_to = "raw_value") %>%
  mutate(
    below_lod   = str_detect(raw_value, "^\\s*<"),
    above_uloq  = str_detect(raw_value, "^\\s*>"),
    censored    = below_lod | above_uloq,
    lod         = if_else(below_lod,  parse_number(raw_value), NA_real_),
    uloq        = if_else(above_uloq, parse_number(raw_value), NA_real_),
    value       = case_when(
      below_lod  ~ lod / sqrt(2),
      above_uloq ~ uloq,
      TRUE       ~ parse_number(raw_value)
    ),
    log10_value = log10(value),
    Factor1     = case_when(
      str_detect(Sample, fixed(FACTOR1_LEVELS[2])) ~ FACTOR1_LEVELS[2],
      str_detect(Sample, fixed(FACTOR1_LEVELS[1])) ~ FACTOR1_LEVELS[1],
      TRUE                                         ~ NA_character_
    ),
    Factor2     = case_when(
      str_detect(Sample, fixed(FACTOR2_LEVELS[2])) ~ FACTOR2_LEVELS[2],
      str_detect(Sample, fixed(FACTOR2_LEVELS[1])) ~ FACTOR2_LEVELS[1],
      TRUE                                         ~ NA_character_
    )
  )

# ── 3. Drop rows whose design assignment is unknown, and say how many ────────
unassigned <- dat_long %>% filter(is.na(Factor1) | is.na(Factor2))
n_unassigned_rows    <- nrow(unassigned)
unassigned_samples   <- sort(unique(unassigned$Sample))

dat_long <- dat_long %>%
  filter(!is.na(Factor1), !is.na(Factor2)) %>%
  mutate(
    Factor1 = factor(Factor1, levels = FACTOR1_LEVELS),
    Factor2 = factor(Factor2, levels = FACTOR2_LEVELS),
    Group   = factor(
      paste0(Factor1, "+", Factor2),
      # Derived from the configured level counts — works for any n x m design.
      levels = paste0(
        rep(FACTOR1_LEVELS, each  = length(FACTOR2_LEVELS)),
        "+",
        rep(FACTOR2_LEVELS, times = length(FACTOR1_LEVELS))
      )
    )
  )

stopifnot(nrow(dat_long) > 0)

cat(sprintf(
  paste0("Wells read: %d\n",
         "  removed as non-sample (standard/blank/background): %d\n",
         "Measurement rows after reshaping: %d\n",
         "  removed — Sample matched no level of Factor1/Factor2: %d (%d sample%s)\n",
         "Rows retained for analysis: %d\n"),
  n_wells_in, n_non_sample,
  n_unassigned_rows + nrow(dat_long),
  n_unassigned_rows, length(unassigned_samples),
  if (length(unassigned_samples) == 1) "" else "s",
  nrow(dat_long)
))
if (length(unassigned_samples) > 0) {
  cat("Unassigned samples: ", paste(unassigned_samples, collapse = ", "), "\n", sep = "")
}
```

# QC

## Censoring Rate per Analyte

```{r qc-censoring}
qc_censor <- dat_long %>%
  group_by(Analyte) %>%
  summarise(
    n            = n(),
    below_n      = sum(below_lod),
    below_pct    = 100 * below_n / n,
    above_n      = sum(above_uloq),
    above_pct    = 100 * above_n / n,
    total_pct    = 100 * sum(censored) / n,
    .groups      = "drop"
  ) %>%
  arrange(desc(total_pct))

qc_censor %>%
  knitr::kable(
    caption   = "Censoring Rate per Analyte (Below LoD and Above ULoQ)",
    col.names = c("Analyte", "Total Samples",
                  "< LoD Count", "< LoD %",
                  "> ULoQ Count", "> ULoQ %",
                  "Any Censoring %"),
    digits    = 1
  )
```

> **Guidance:** Analytes with >50% censoring warrant Tobit regression as a sensitivity
> check. The standard substitution (LoD/√2) used here is appropriate for <30% left-censoring.
> Right-censored values are substituted at the ULoQ, which **attenuates** any effect involving
> them — treat those estimates as conservative lower bounds, and re-run the affected analytes
> at a higher dilution rather than reporting the substituted value as a measurement.

```{r qc-censoring-note, results='asis', echo=FALSE}
worst_above <- qc_censor %>% filter(above_pct > 0) %>% arrange(desc(above_pct))
if (nrow(worst_above) > 0) {
  cat(sprintf(
    "> **Note:** %d analyte(s) contain values above the upper limit of quantification (max %.1f%%: %s). Their effect estimates are attenuated.\n",
    nrow(worst_above), worst_above$above_pct[1], worst_above$Analyte[1]
  ))
}
```

## Outliers (Tukey 3×IQR on log10 scale)

```{r qc-outliers}
fences <- dat_long %>%
  group_by(Analyte) %>%
  summarise(
    q1    = quantile(log10_value, 0.25, na.rm = TRUE),
    q3    = quantile(log10_value, 0.75, na.rm = TRUE),
    iqr   = IQR(log10_value, na.rm = TRUE),
    lower = q1 - 3 * iqr,
    upper = q3 + 3 * iqr,
    .groups = "drop"
  )

dat_long <- dat_long %>%
  left_join(fences, by = "Analyte") %>%
  mutate(
    Outlier = if_else(
      !is.na(log10_value) & (log10_value < lower | log10_value > upper),
      "Yes", "No"
    )
  ) %>%
  select(-q1, -q3, -iqr, -lower, -upper)

outlier_tbl <- dat_long %>%
  filter(Outlier == "Yes") %>%
  count(Analyte, name = "outlier_n") %>%
  arrange(desc(outlier_n))

outlier_tbl %>%
  knitr::kable(
    caption   = "Wells Flagged as Outliers (Tukey 3xIQR on log10 scale)",
    col.names = c("Analyte", "Flagged Wells")
  )

# Flagging and excluding are separate decisions — EXCLUDE_OUTLIERS controls the
# second one, so the code and the text above cannot drift apart.
n_outliers <- sum(dat_long$Outlier == "Yes")
if (isTRUE(EXCLUDE_OUTLIERS)) {
  dat_long <- dat_long %>% filter(Outlier != "Yes")
  cat(sprintf("EXCLUDE_OUTLIERS = TRUE: %d flagged well-measurements removed from all models.\n",
              n_outliers))
} else {
  cat(sprintf("EXCLUDE_OUTLIERS = FALSE: %d flagged well-measurements are reported above but RETAINED in all models.\n",
              n_outliers))
}
```

# Two-Way ANOVA (per Analyte)

```{r anova}
effect_map <- c(
  setNames(
    paste0(FACTOR1_LEVELS[2], " vs ", FACTOR1_LEVELS[1]),
    paste0("Factor1", FACTOR1_LEVELS[2])
  ),
  setNames(
    paste0(FACTOR2_LEVELS[2], " vs ", FACTOR2_LEVELS[1]),
    paste0("Factor2", FACTOR2_LEVELS[2])
  ),
  setNames(
    "Interaction",
    paste0("Factor1", FACTOR1_LEVELS[2], ":Factor2", FACTOR2_LEVELS[2])
  )
)

# The correction family is "this model term, across all analytes". The
# intercept tests "mean log10 concentration = 0", which is never a hypothesis of
# interest, is always astronomically significant, and would both inflate m and
# occupy the top ranks — so it is dropped before correcting.
res_lm <- dat_long %>%
  group_by(Analyte) %>%
  group_modify(~ {
    m <- lm(log10_value ~ Factor1 * Factor2, data = .x)
    broom::tidy(m)
  }) %>%
  ungroup() %>%
  filter(term != "(Intercept)") %>%
  group_by(term) %>%
  mutate(adj.p = p.adjust(p.value, method = "BH")) %>%
  ungroup() %>%
  mutate(
    effect      = coalesce(unname(effect_map[term]), term),
    fold_change = 10^estimate
  ) %>%
  arrange(Analyte, effect)

res_lm %>%
  head(20) %>%
  select(Analyte, term, estimate, std.error, statistic, p.value, effect, adj.p, fold_change) %>%
  knitr::kable(
    caption     = "Two-Way ANOVA Model Summary (First 20 Effects; intercepts excluded, BH within each term)",
    col.names   = c("Analyte", "Term", "Estimate", "Std Error", "Statistic", "Raw p", "Effect", "FDR adj. p", "Fold Change"),
    digits      = 4
  )
readr::write_csv(res_lm, file.path(OUTPUT_DIR, "model_LM_results.csv"))
```

# Post-Hoc Contrasts (emmeans)

```{r emmeans}
group_levels <- levels(dat_long$Group)

emm_contrasts <- dat_long %>%
  group_by(Analyte) %>%
  group_modify(~ {
    m   <- lm(log10_value ~ Factor1 * Factor2, data = .x)
    emm <- emmeans(m, ~ Factor1 * Factor2)

    drug_effects    <- contrast(emm, method = "revpairwise", by = "Factor1", infer = TRUE)
    disease_effects <- contrast(emm, method = "revpairwise", by = "Factor2", infer = TRUE)
    interaction_eff <- contrast(emm, interaction = c("revpairwise", "revpairwise"), infer = TRUE)

    s <- bind_rows(
      as.data.frame(drug_effects),
      as.data.frame(disease_effects),
      as.data.frame(interaction_eff)
    )
    s_tidy <- s %>%
      mutate(
        unique_contrast = case_when(
          !is.na(Factor1) ~ paste(contrast, Factor1, sep = ", "),
          !is.na(Factor2) ~ paste(contrast, Factor2, sep = ", "),
          TRUE            ~ contrast
        )
      )

    tibble(
      contrast       = s_tidy$unique_contrast,
      estimate_log10 = s_tidy$estimate,
      SE             = s_tidy$SE,
      df             = s_tidy$df,
      t              = s_tidy$t.ratio,
      p              = s_tidy$p.value,
      fold_change    = 10^s_tidy$estimate,
      lower_FC       = 10^s_tidy$lower.CL,
      upper_FC       = 10^s_tidy$upper.CL
    )
  }) %>%
  ungroup() %>%
  # Same family definition as the model table above: correct across analytes
  # within each contrast, not across every analyte-by-contrast cell at once.
  group_by(contrast) %>%
  mutate(p_adj = p.adjust(p, method = "BH")) %>%
  ungroup() %>%
  arrange(contrast, p_adj)

emm_contrasts %>%
  head(20) %>%
  knitr::kable(
    caption     = "emmeans Post-Hoc Contrasts & Fold Changes (First 20 Comparisons)",
    col.names   = c("Analyte", "Contrast", "Log10 Est.", "SE", "df", "t ratio", "Raw p", "Fold Change", "Lower FC", "Upper FC", "FDR adj. p"),
    digits      = 4
  )
readr::write_csv(emm_contrasts, file.path(OUTPUT_DIR, "emmeans_contrasts_foldchanges.csv"))
```

# Visualization

## Effect Size Heatmap

```{r heatmap, fig.width=6, fig.height=4}
heatmap_data <- emm_contrasts %>%
  filter(!is.na(contrast)) %>%
  mutate(
    log2_FC      = log2(fold_change),
    log2_FC_plot = if_else(p_adj < 0.05, log2_FC, 0),
    Analyte      = factor(Analyte, levels = sort(unique(Analyte))),
    contrast_disp = factor(contrast,
                           levels = CONTRAST_ORDER,
                           labels = CONTRAST_LABELS)
  ) %>%
  filter(!is.na(contrast_disp))

lim <- max(abs(heatmap_data$log2_FC_plot), na.rm = TRUE)
lim <- max(lim, 1)

p_heat <- ggplot(heatmap_data,
                 aes(x = contrast_disp, y = Analyte, fill = log2_FC_plot)) +
  geom_tile(color = "grey85", linewidth = 0.2) +
  scale_fill_gradient2(
    low      = "#3B4CC0",
    mid      = "white",
    high     = "#D55E00",
    midpoint = 0,
    limits   = c(-lim, lim),
    breaks   = pretty_breaks(n = 5),
    labels   = label_number(accuracy = 0.1)
  ) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.05))) +
  labs(
    x       = NULL,
    y       = NULL,
    fill    = "Log\u2082 Fold Change",
    title   = "Analyte-wise Effects of Treatments",
    caption = "Non-significant tiles masked to 0 (FDR < 0.05)"
  ) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    legend.position  = "right",
    legend.direction = "vertical",
    legend.title     = element_text(size = 11),
    legend.key.width = unit(12, "pt"),
    axis.text.x      = element_text(angle = 30, hjust = 1, face = "bold"),
    axis.text.y      = element_text(size = 11, lineheight = 1.3, face = "bold"),
    axis.ticks       = element_blank(),
    axis.line        = element_blank(),
    plot.title       = element_text(size = 14, face = "bold", margin = margin(b = 6)),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave(file.path(OUTPUT_DIR, "Heatmaps", "heatmap.tiff"),
       plot = p_heat, width = 6, height = 4, dpi = 600,
       device = "tiff", compression = "lzw")
print(p_heat)
```

# Per-Analyte Visualization Gallery {.tabset .tabset-fade}

```{r plot-helpers-faceted, message=FALSE, warning=FALSE}
# The gallery figures below show the data; the significance testing lives in the
# emmeans table above. No per-panel brackets are drawn, so no per-analyte
# contrasts are computed here.

# Complete the palette for any group the config does not name, so a design
# other than 2x2 does not fail on a missing colour.
OKABE_ITO_SEQ <- c("#0072B2", "#E69F00", "#009E73", "#D55E00",
                   "#CC79A7", "#56B4E9", "#F0E442", "#999999")
group_levels_all <- levels(dat_long$Group)
GROUP_COLORS <- setNames(OKABE_ITO[group_levels_all], group_levels_all)
if (any(is.na(GROUP_COLORS))) {
  GROUP_COLORS[is.na(GROUP_COLORS)] <-
    rep(OKABE_ITO_SEQ, length.out = sum(is.na(GROUP_COLORS)))
}

make_faceted_plot <- function(plot_dat, plot_type = c("dot", "bar", "violin", "box")) {
  plot_type <- match.arg(plot_type)
  
  p <- ggplot(plot_dat, aes(x = Group, y = log10_value)) +
    scale_color_manual(values = GROUP_COLORS) +
    scale_y_continuous(breaks = pretty_breaks(5), labels = math_format(10^.x)) +
    labs(y = "log\u2081\u2080 concentration (pg/mL)", x = NULL) +
    theme_bw(base_size = 11, base_family = "sans") +
    theme(
      legend.position = "none",
      axis.line       = element_line(linewidth = 0.6),
      axis.ticks      = element_line(linewidth = 0.6),
      axis.text.x = element_text(angle = 30, hjust = 1, size = 9),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      strip.text = element_text(face = "bold", size = 10),
      strip.background = element_rect(fill = "grey95", color = "grey70")
    )
  
  # `fill` is only mapped by the bar/violin/box variants; adding the fill scale
  # unconditionally makes ggplot warn about a scale with no matching data.
  if (plot_type != "dot") {
    p <- p + scale_fill_manual(values = GROUP_COLORS)
  }

  if (plot_type == "dot") {
    p <- p +
      geom_quasirandom(aes(color = Group), width = 0.15, alpha = 0.8, size = 1.5) +
      stat_summary(fun.data = mean_cl_normal, geom = "errorbar",
                   width = 0.18, linewidth = 0.4)
  } else if (plot_type == "bar") {
    p <- p +
      stat_summary(aes(fill = Group), fun = mean, geom = "col",
                   width = 0.6, alpha = 0.45, color = NA) +
      geom_quasirandom(aes(color = Group), width = 0.15, size = 1.5, alpha = 0.9) +
      stat_summary(fun.data = mean_cl_normal, geom = "errorbar",
                   width = 0.18, linewidth = 0.5)
  } else if (plot_type == "violin") {
    p <- p +
      geom_violin(aes(fill = Group, color = Group),
                  width = 0.8, alpha = 0.15, trim = FALSE, linewidth = 0.6) +
      geom_quasirandom(aes(color = Group), width = 0.15, size = 1.5, alpha = 0.9)
  } else if (plot_type == "box") {
    p <- p +
      geom_boxplot(aes(fill = Group, color = Group),
                   width = 0.6, alpha = 0.15, outlier.shape = NA, linewidth = 0.6) +
      geom_quasirandom(aes(color = Group), width = 0.15, size = 1.5, alpha = 0.9)
  }
  
  p + facet_wrap(~Analyte, scales = "free_y", ncol = 4)
}

write_plot <- function(p, name) {
  ggsave(file.path(OUTPUT_DIR, name),
         plot = p, width = 14, height = 8, dpi = 600,
         device = "tiff", compression = "lzw")
}
```

## Dot Plots

```{r dot-gal, fig.width=14, fig.height=8, message=FALSE, warning=FALSE}
p_dot <- make_faceted_plot(dat_long, "dot")
write_plot(p_dot, "Heatmaps/gallery_dotplots.tiff")
print(p_dot)
```

## Bar Plots

```{r bar-gal, fig.width=14, fig.height=8, message=FALSE, warning=FALSE}
p_bar <- make_faceted_plot(dat_long, "bar")
write_plot(p_bar, "Heatmaps/gallery_barplots.tiff")
print(p_bar)
```

## Violin Plots

```{r viol-gal, fig.width=14, fig.height=8, message=FALSE, warning=FALSE}
p_viol <- make_faceted_plot(dat_long, "violin")
write_plot(p_viol, "Heatmaps/gallery_violinplots.tiff")
print(p_viol)
```

## Box Plots

```{r box-gal, fig.width=14, fig.height=8, message=FALSE, warning=FALSE}
p_box <- make_faceted_plot(dat_long, "box")
write_plot(p_box, "Heatmaps/gallery_boxplots.tiff")
print(p_box)
```
