---
title: "03 · Multi-Condition ANOVA"
subtitle: "Welch's one-way ANOVA with planned contrasts and Holm correction"
description: |
  Compares multiple cell types or treatment conditions across one or more
  viral/experimental conditions. Uses Welch's ANOVA (robust to unequal
  variances) on log₁₀-transformed data, with planned pairwise contrasts
  against a reference group and Holm multiple-testing correction.
categories: [infection, statistics, ANOVA]
---

## Overview

| Item | Details |
|------|---------|
| **Input** | Wide-format CSV per condition (columns = cell types, rows = replicates) |
| **Key packages** | `tidyverse`, `car`, `broom` |
| **Statistics** | Shapiro-Wilk, Bartlett, Levene, Welch one-way ANOVA, Welch t-tests |
| **Correction** | Holm (FWER) |
| **Output** | Diagnostic table · ANOVA result · Contrast table with fold-changes |
| **Download** | [template.Rmd](template.Rmd) |

::: {.callout-tip}
**When to use this template:** You have measurements (e.g., viral titres, protein levels)
across several cell types or treatment groups, and you need to test whether a reference
group differs from the others, accounting for potential heteroscedasticity.
:::

---

```{r}
#| label: run-simulate
#| include: false
# Generate synthetic example data before rendering the template
source("data/simulate_data.R")
```

```{r child="template.Rmd"}
```
